### 2024-03-06 presentation/presentation.tex

\documentclass{beamer}
\setbeamertemplate{caption}[numbered]
\usepackage{tikz}
\usepackage{graphicx}
\usetheme{AnnArbor}
%\usecolortheme{wolverine}
\title[AD for POMPs]{Automatic Differentiation Accelerates Inference for Partially Observed Stochastic Processes}
\subtitle{\small{Inference for Partially Observed Structured Dynamic Systems}}
\author[Kevin Tan]{Kevin Tan \\\vspace{0.5cm} \small{Joint work with Giles Hooker and Edward Ionides}}
\date[\today]{\today}

\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{mathtools}
\usepackage{bbold}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{pifont}
%\usepackage[left=1in,right=1in,]{geometry}
\usepackage{multicol,lipsum,xcolor}
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{amsmath,amsthm,amssymb,bm}
\usepackage{bbold}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{enumerate}
%\usepackage{enumitem}
\usepackage{natbib}
\usepackage{parskip}
\usepackage{comment}
\usepackage{microtype}
\usepackage{diagbox}
\usepackage{tikz}
\usepackage{fancyhdr}

\newcommand{\Prob}{\displaystyle \mathbb {P}}
\newcommand{\jac}{\bigg| \frac{\dif x}{\dif y} \bigg|}
\newcommand{\inftyint}{\int_{-\infty}^{\infty}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\Y}{\mathcal{Y}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\C}{\mathbb{C}}

% best package ever
%\usepackage{realhats}

\def\mathfamilydefault{\rmdefault}

\usepackage{color}
\usepackage{xcolor}
\definecolor{hlpink}{rgb}{0.96, 0.76, 0.76}
\definecolor{hlblue}{rgb}{0.54, 0.81, 0.94}
\definecolor{hlyellow}{rgb}{0.98, 0.91, 0.71}
\definecolor{hlgreen}{rgb}{0.7, 0.93, 0.36}
\definecolor{hlpurple}{rgb}{0.8, 0.8, 1.0}

\newcommand{\highlight}[2][yellow]{\mathchoice%
  {\colorbox{#1}{$\displaystyle#2$}}%
  {\colorbox{#1}{$\textstyle#2$}}%
  {\colorbox{#1}{$\scriptstyle#2$}}%
  {\colorbox{#1}{$\scriptscriptstyle#2$}}}%

 \AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

\include{mop/macros}

\begin{document}

\maketitle

\section{Background}

\begin{frame}{Partially Observed Markov Processes}

\end{frame}

\section{Introduction}

\begin{frame}{Maximum Likelihood Inference is Hard in POMPs}

\begin{enumerate}
    \item \textbf{Intractable likelihood functions}! Bypassed with e.g. simulated likelihood, likelihood-free inference, particle filters, etc. 
    \item Want to perform inference without having access to the transition density of the latent Markov process, using only  a \textbf{simulator} (\cite{welch2009abc, wood2010sl, doucet2010pmcmc, ionides08, ionides15}). 
    \item \textbf{Significant Monte-Carlo noise in the likelihood estimate} makes accurate parameter estimation difficult.
\end{enumerate}
\end{frame}

\begin{frame}{When Does IF2 Struggle?}

\begin{itemize}
    \item IF2 is great at approaching the MLE quickly! 
    \item Performs parameter updates at every timestep, 
\end{itemize}

\begin{figure}
    \centering
    \includegraphics[scale=0.5]{imgs/095/optim.png}
    \label{fig:enter-label}
\end{figure}

\end{frame}

\bibliographystyle{apalike}
\bibliography{mop/bib-ifad, mop/bib-ref}

\end{document}