### 2024-05-10 mop/draft.tex
Updates from Overleaf

\documentclass{article}
\usepackage[margin=2cm]{geometry}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}

\setlist{nosep} % to reduce space in lists

\setlength{\parskip}{0.5em}
\setlength{\parindent}{0in}

\input{macros}

\newcommand{\off}{\operatorname{off}}
\newcommand{\on}{\operatorname{on}}

\title{Accelerated Inference for Partially Observed Stochastic Processes with Automatic Differentiation}
\author{Kevin Tan, Giles J. Hooker, Edward L. Ionides}
\date{}

\comments=1

\begin{document}
\maketitle
\begin{abstract}

Automatic differentiation (AD) has driven recent advances in machine learning, including deep neural networks and Hamiltonian Markov Chain Monte Carlo methods. Partially observed nonlinear stochastic dynamical systems have proved resistant to AD techniques due to (1) the requirement for simulation-based inference that does not require access to the system's transition probabilities and (2) the issue that widely used particle filter algorithms yield a likelihood function that is discontinuous as a function of the model parameters. We propose a new representation that embeds existing methods in a theoretical framework that readily permits extension to a new class of algorithms, providing opportunities for optimizing a bias/variance tradeoff. Further, we develop optimization algorithms suited to the Monte Carlo properties of the derivative estimate, including a hybrid algorithm that requires only a differentiable simulator for maximum likelihood estimation. Promising numerical results indicate that a hybrid algorithm that uses AD to refine a coarse solution from iterated filtering can beat current state-of-the-art methods on a challenging scientific benchmark problem.
\end{abstract}

\section{Introduction}

Maximum-likelihood estimation for inference in partially observed stochastic processes, also known as, continuous-state continuous-time hidden Markov models (HMMs), partially-observed Markov processes (POMPs), or partially-observed stochastic nonlinear dynamical systems, is a challenging problem that faces several challenges: 
\begin{enumerate}
    \item Intractable likelihood functions, bypassed with the approaches of simulated likelihood and likelihood-free inference. 
    \item The desire for simulation-based inference without access to the transition density of the latent Markov process, with various methods created by \cite{welch2009abc, wood2010sl, doucet2010pmcmc, ionides08, ionides15}. 
    \item Accurate parameter estimation in the presence of significant Monte-Carlo noise in the likelihood estimate.
\end{enumerate}

Many approaches to inference in these scenarios (such as the EM algorithm, the various Kalman filter variants, and MCMC) either struggle with intractable likelihood functions when the models are complex, or assume access to the probability density of next states given the current state. This is a problem in some critical applications, such as disease modeling, where the models are complex enough such that even obtaining the transition density is an intractable problem. However, the particle filter with the bootstrap proposal, a popular method for solving the filtering problem in partially-observed dynamical systems, provides an unbiased estimate of the likelihood (\cite{delmoral2004feynman}) without requiring evaluation of the transition density of the latent Markov process, enabling an arbitrary model simulator to be plugged into the algorithm.

This is the backbone behind the only simulation-based full-information maximum likelihood method available in the literature thus far, the improved iterated filtering algorithm of \citet{ionides15}. \citet{ionides15} utilize the bootstrap filter for perform maximum likelihood estimation with an iterated perturbed Bayes map, successfully tackling challenging problems in epidemiology that available Bayesian methodology fails to solve.

Still, maximum likelihood parameter estimation can be challenging, especially when the Monte Carlo variance of the evaluation is high and the number of parameters is not small. For example, though in practice the algorithm of \citet{ionides15} converges quickly to a neighborhood of the MLE, it often struggles to successfully optimize the last few-units of the log-likelihood. \citet{Ionides_mcap} attempt to address the third problem, albeit with a computationally-expensive approach.

\subsection{Automatic Differentiation for Particle Filters}

Recent advances in automatic differentiation (AD) for particle filters (\cite{blei2018vsmc, jon2018diffpf, corenflos21, scibior2021dpf, doucet2022particlebased}) have drawn attention to AD as an alternative tool for inference in partially observed stochastic processes with the particle filter. However, these existing approaches either are not yet compatible with the desire for simulation-based inference, or are computationally expensive.

\citet{blei2018vsmc} and \citet{jon2018diffpf} drop high-variance resampling terms from the gradient, leading to bias that does not vanish asymptotically according to \citet{corenflos21}. Though the approach of \citet{corenflos21} is promising, it is computationally expensive. \citet{doucet2022particlebased} provide another promising fixed-lag smoothing-based approach that is similar in spirit to our method (MOP-$\alpha$), but requires either (1) access to the process model transition densities or (2) the special case where the transition density factors into a policy that selects actions (in the reinforcement learning sense) and a deterministic model that maps states and actions to next-timestep states.

Finally, \citet{scibior2021dpf} demonstrate that the estimators derived in \citet{poyiadjis11} can be attained with AD, albeit with a simple tweak to the particle filter that does not modify its likelihood estimates. Though their method is not fully compatible with simulation-based inference, we show that another simple tweak enables this compatibility and corresponds to a special case of our MOP-$\alpha$ algorithm.

\subsection{Our Contributions}

Previous research on AD for particle filters has struggled with various issues: 
\begin{enumerate}
    \item Bias, if the discontinuous nature of particle resampling is ignored.
    \item High Monte Carlo variance, if continuity corrections lead to numerical instability.
    \item High computational cost, for algorithms which involve pairwise interactions between particles, marginalization, smoothing or optimal transport.
    \item Reduced applicability for algorithms which lose or fail to take advantage of the simulation-based capabilities of the bootstrap filter.
\end{enumerate}

We develop a new approach to statistical inference via AD for particle filters which addresses these concerns. We develop a new theoretical framework that encompasses existing methods, and addresses the seemingly incompatible paradox of differentiating through a Monte-Carlo algorithm with discontinuous resampling.

This is done through the construction of a smooth extension to the particle filter, as well as a \textbf{Measurement Off-Policy-$\alpha$} (MOP-$\alpha$) family of algorithms that (informally) encompasses the special case of AD of a vanilla PF (recovering the gradient estimator of \citet{blei2018vsmc}) when $\alpha=0$, and the DPF algorithm from \citet{scibior2021dpf} (that recovers the gradient estimator from \citet{poyiadjis11}) when $\alpha=1$. When $\alpha \in (0,1)$, this provides opportunities for optimizing a bias/variance tradeoff.

We show these off-policy particle filters successfully target the posterior, recover desirable gradient estimates previously encountered in the literature, and possess other desirable theoretical properties. To cap off the theoretical results, we also derive a linear convergence rate for gradient methods on the particle filter in the presence of strong convexity, providing support for numerical optimization with AD for the particle filter.

In practice, we propose a hybrid algorithm that we call \textbf{Iterated Filtering with Automatic Differentiation (IFAD)} that warm-starts gradient ascent (or a similar iterative first or second-order gradient method) with a coarse solution obtained from a few rounds of iterated filtering. While this requires a differentiable simulator compatible with the reparametrization trick (as previous literature such as \citet{corenflos21} also require), future work will deal with workarounds inspired by our framework that use likelihood ratios of transition densities in place of this. Promising numerical results indicate that IFAD beats IF2 (and by the transitive property, IF1, the Liu-West filter, and particle MCMC) on a challenging scientific benchmark problem.

While the Monte-Carlo Adjusted Profile (MCAP) from \citet{Ionides_mcap} provides a profile likelihood confidence interval that takes into account both Monte Carlo profile error and statistical uncertainty in the likelihood, it is computationally inefficient. \kevin{TODO} We provide a method for incorporating the MCAP methodology (with the poor man's likelihood profile) in our IFAD algorithm that provides a more computationally efficient profile likelihood confidence interval. 
\kevin{Quantile regression for estimating the max/the envelope of the likelihood evaluations from AD, then fit MCAP/MQLE/MSLE?}

Finally, we make the first steps towards developing a Python counterpart to the popular \texttt{pomp} R package by \citet{king16, king2017pompmanual} that we call \texttt{pompy}. The software we provide includes welcome features for scientific computing such as GPU acceleration, parallel computing for single-program-multiple-data programs, just-in-time compilation, and automatic differentiation enabled by the \texttt{jax} Python package from \citet{jax2018github}.

\kevin{Terminology: Plug-and-play (clearly definable, but not well-known, define at start of paper and don't use in title/abstract), likelihood-free maximum likelihood (seemingly an oxymoron, rationale being there is no need to evaluate density, bayesians call likelihood densities, plus we only approximate the likelihood and don't really have it), simulation-based inference (loaded term)}

\subsection{Significance Statement}
Many scientific models involve highly nonlinear stochastic dynamical systems which can be observed only via noisy and incomplete measurements. Under the Markov assumption on system dynamics, previous work has provided methods of performing inference for these models. In particular, prior to this work, iterated filtering algorithms were the only class of algorithms for maximum likelihood estimation that did not require access to the system's transition probabilities, instead needing only a simulator of the system dynamics. We leverage recent advances in automatic differentiation to propose a hybrid algorithm that requires only a differentiable simulator for maximum likelihood estimation. Our new method outperforms previous approaches on a challenging problem in epidemiology. This is implemented in a precursor to a Python counterpart to the popular \texttt{pomp} package by \citet{king16, king2017pompmanual} that includes welcome features for scientific computing such as GPU acceleration, parallel computing for single-program-multiple-data programs, just-in-time compilation, and automatic differentiation.

\kevin{TODO: (1) Fix beta smoothness to d(n) at no more than O(n), (2) add story to say that no one spotted that it could have been plug-and-play because of e.g. reasons why they would stumble}

\section{Problem Setup}

Consider an unobserved Markov process $\{X_t, t \geq t_0\}$, and observations $Y_1,...,Y_N$ at timesteps $t_1,..., t_N$. The process is parameterized by an unknown parameter $\theta \in \Theta$, where $X \in \gX, Y \in \gY,$ and finally $\gX \subseteq \R^d, \Theta \subseteq \R^p$. By a similar decomposition to that in \citet{doucet2009tutorial}, we find that the joint density of $X_{0:N}, Y_{1:N}$ can be factored as
$$f_{X_{0: N}, Y_{1: N}}\left(x_{0: N}, y_{1: N} ; \theta\right)=f_{X_0}\left(x_0 ; \theta\right) \prod_{n=1}^N f_{X_n \mid X_{n-1}}\left(x_n \mid x_{n-1} ; \theta\right) f_{Y_n \mid X_n}\left(y_n \mid x_n ; \theta\right).$$

We call $f_{X_n|X_{n-1}}\left(x_{n} \mid x_{n-1}; \theta\right)$ the process model, writing $\texttt{process}\left(x_n, \theta\right)$ for the simulator corresponding to the above process model, $f_{Y_n|X_n}\left(y_n \mid x_n, \theta\right)$ the measurement model, and will write $y_n^*$ for the actual values of the observations that were observed.

The above are all defined on a single probability space $(\Omega, \Sigma, \prob),$ where $\omega \in \Omega$ denotes an element of the sample space that can be thought of as a random seed. For notational simplicity, we omit the dependence of the likelihood and other suitable calculations on $X_0.$

\paragraph{Seed-Fixing, Coupling, and the Method of Common Random Numbers, Coupling:}

In the rest of this paper, we will identify the seed with an element of the sample space. We do so because the seed determines the entire sequence of pseudo-random numbers as generated by a computer, to maintain an algorithmic and practical interpretation of our theory. When two algorithms share the same seed, this corresponds to the method of common random numbers.

\section{A Smooth Extension of the Particle Filter}

%\paragraph{Motivation:} 
%The gradient of the log-likelihood can be written as a vector $\nabla_\theta \ell(\theta)$ such that 
%\begin{equation}
    %\lim_{v \to 0} \frac{|\ell(\theta+v) - \ell(\theta) - v^T \nabla_\theta \ell(\theta)|}{\|v\|} = 0.
%\end{equation}

%We therefore need to evaluate the likelihood at a point $\theta$ and a nearby point $\theta + v$ to estimate the gradient with finite differences. But in the presence of Monte Carlo noise in evaluating the likelihood, this is difficult, unless (1) one ensures one uses the same seed for each evaluation in the neighborhood, and (2) chooses a step-size small enough that changing the the parameter does not change the resampling indices.

\paragraph{Intuition:} 
Our theoretical framework for automatic differentiation for the particle filter hinges on the following observation. Instead of seeking to differentiate $\ell(\theta)$ directly, we instead focus on differentiating through a suitable fixed-seed reweighting of a bootstrap filter. The intuition is as follows. Consider a particle filter likelihood estimate where the system (i.e. \textbf{both the state transitions and resampling}) evolves under a parameter $\phi \in \Theta$ and the seed, or sample space element, is fixed at $\omega \in \Omega.$ Conditional on $\omega$ and $\phi,$ one can reweight the likelihood estimate to instead evaluate the likelihood at another (sufficiently nearby) $\theta \in \Theta.$

Differentiating this reweighted likelihood estimate with respect to $\theta$ then amounts to differentiating through the reweighting likelihood ratios, which is differentiable if the densities used for the reweighting are differentiable. We call this simulation-and-reweighting scheme an \textbf{off-policy} likelihood estimate, as it is similar in spirit to the off-policy methods found in the reinforcement learning literature.

The above intuition is formalized below.

\kevin{Method of Common Random Numbers, Coupling}

\subsection{Doubly Off-Policy Particle Filters}

\begin{defn}[Seed-Fixing]
\label{defn:seed-fixing}
Consider a probability space $(\Omega, \Sigma, \mu)$ governing the particles and observations. Noting that the particle filter is unbiased for the likelihood when computing the expectation of the normalizing factor under the posterior $f_{X_{0:N|Y_{1:N}}}\left(x_{0: N} \mid y_{1: N}; \theta\right)$, we can write the likelihood as
$$
\begin{aligned}
\mathcal{L}(\theta) & =\int_{\Omega} \mathcal{L}(\theta, \omega, J) d \mu(\omega) \\
& =\int_{\Omega}\left(\prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f_{Y_n|X_n}\left(y_n^* \mid x_{n, j}^P(\omega), \theta\right)\right) f_{X_{0:N}, Y_{1:N}}\left(x_{0: N}^P(\omega) , y_{1: N}^*, \theta\right) d \mu(\omega) \\
& =\int_{\Omega}\left(\prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f_{Y_n|X_n}\left(y_n^* \mid x_{n, j}^P(\omega), \theta\right) f_{X_n|X_{n-1}}(x_{n,j}^P(\omega)|x_{n-1,j}^F(\omega), \theta)\right) f_{X_{0:N}| Y_{1:N}}\left(x_{n, j}^P(\omega) \mid y_n^*, \theta\right) d \mu(\omega)
\end{aligned}
$$
where $\mathcal{L}(\theta, \omega, J)$ is the Monte-Carlo estimate of the likelihood given the realization of $\mathrm{J}$ particles corresponding to $\omega \in \Omega$. For simplicity, we resample every timestep. We call $\omega$ a seed, and the deterministic quantity $\mathcal{L}(\theta, \omega, J)$ the seed-fixed estimate.
\end{defn}

Armed with this seed-fixed likelihood estimate, we modify a particle filter likelihood estimate where the system evolves under a parameter $\phi \in \Theta$ and seed $\omega \in \Omega$ to instead evaluate the likelihood at another (sufficiently nearby) $\theta \in \Theta.$

\begin{defn}[Doubly Off-Policy Particle Filter Likelihood]
\label{defn:doubly-off-policy}
Let $\lik(\theta, \phi, \omega, J)$ denote a Monte-Carlo estimate of the likelihood evaluated at $\theta \in \Theta$ with $J$ particles, but where the system evolves according to another $\phi \in \Theta$ and conditional on a fixed seed $\omega \in \Omega$. This can be computed with a suitable reweighting of the particles, which we provide the derivation for later. We then have
\begin{equation} \label{eq:loglik2}
\lik(\theta)=\E[\lik(\theta,\phi,\Omega, J)]=\int_\Omega \lik(\theta,\phi,\omega, J)\, d\mu(\omega),
\end{equation}
where we define
\begin{align}
    \lik(\theta,\phi,\omega, J) = \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J &\left(f_{Y_n|X_n}(y_n^* | x_{n,j}^{P,\phi}(\omega), \phi) \cdot 
    f_{X_n|X_{n-1}}(x_{n,j}^{P, \phi}(\omega) | x_{n-1,j}^{F,\phi}, \phi) \right.\\
    &\cdot
    \left.
    \frac{f_{Y_n|X_n}(y^*_n|x_{n,j}^{P,\phi}(\omega),\theta)}{f_{Y_n|X_n}(y^*_n|x_{n,j}^{P,\phi}(\omega),\phi)}
    \cdot \frac{f_{X_n|X_{n-1}}(x_{n,j}^{P,\phi}(\omega)|x_{n-1,j}^{F,\phi}(\omega),\theta)}{f_{X_n|X_{n-1}}(x_{n,j}^{P,\phi}(\omega)|x_{n-1,j}^{F,\phi}(\omega),\phi)}\right).
\end{align}
\end{defn}

Lastly, we require the following regularity conditions.

\begin{aspt}[Continuity of the Likelihood] $\ell(\theta)$ proper is continuous in a neighborhood $\left\{\theta: \ell(\theta)>\lambda_1\right\}$ for some $\lambda_1<\sup _{\varphi} \ell(\varphi)$.
\end{aspt}

\begin{aspt}[Bounded Process Model]
\label{assump:bounded-process}
    There exist $\underbar{M}, \bar{M}$ such that $0 < \underbar{M} \leq f_{X_n|X_{n-1}}(x_n | x_{n-1}) \leq \bar{M} < \infty$. 
\end{aspt}

\begin{aspt}[Bounded Measurement Model]
\label{assump:bounded-measurement}
There exist $\underbar{G}, \bar{G}$ such that $0<\underbar{G} \leq f_{Y_n \mid X_n}\left(y_n^* \mid x_n; \theta\right) \leq \bar{G} < \infty$ and $\infty > G' \geq G'(\theta)>0$ such that $||\nabla_\theta \log f_{Y_n \mid X_n}\left(y_n^* \mid x_n; \theta\right)||_\infty <\leq G'(\theta) \leq G'$ for all $1 \leq n \leq N, x_n \in \gX$ and $\theta \in \Theta.$
\end{aspt}

\begin{aspt}[Locally Bounded Derivatives]
\label{assump:local-bounded-derivative}
Let $\mathcal{M}$ be an open subset of $\Omega$. There exists some function $G(\theta)$ and constants $G=\sup _{\theta \in \mathcal{N}} G(\theta)<\infty$, $H=\sup _{\theta \in \mathcal{N}} H(\theta)<\infty$ such that
    $$
    \|\nabla_\theta \ell(\theta, \phi, \omega, J)\|_\infty \leq G(\theta) \leq G<\infty, \;\;
    \|\nabla^2_\theta \ell(\theta, \phi, \omega, J)\|_\infty \leq G(\theta)^2 \leq G^2<\infty
    $$
    for every $\phi \in \mathcal{M}, \theta \in \mathcal{N}_J(\phi)$, and almost every $\omega \in \Omega$.
\end{aspt}

We need the following assumption to ensure that the weighted particle filter with off-policy resampling targets the posterior:
\begin{aspt}[Bounded And Differentiable Measurement Density Ratios]
    \label{assump:diff-meas}
    For any $\phi \in \Theta$, there exists an open neighborhood $\gN(\phi)$ containing $\phi$ such that the density ratios $\frac{f_{Y_n|X_n}(y_n^*|x, \theta)}{f_{Y_n|X_n}(y_n^*|x, \phi)}$ are bounded and have more than two continuous derivatives for all timesteps $n=1,...,N$, $x \in \gX$ and all $\theta\in\gN(\phi)$. 
\end{aspt}

In the following assumption, the first condition is sufficient for DOP, and the second for MOP:
\begin{aspt}[Bounded And Differentiable Process Model]
    \label{assump:diff-process}
    Either one of the following two conditions hold:
    \begin{enumerate}
        \item For any $\phi \in \Theta$, there exists an open neighborhood $\gN(\phi)$ containing $\phi$ such that the density ratios $\frac{f_{X_n|X_{n-1}}(x'|x, \theta)}{f_{X_n|X_{n-1}}(x'|x, \phi)}$ are bounded and have more than two continuous derivatives for all timesteps $n=1,...,N$, $x, x' \in \gX$ and all $\theta\in\gN(\phi)$, or
        \item There exists a differentiable simulator with more than two continuous derivatives for the process model. 
    \end{enumerate}
\end{aspt}

We want $\phi\in\Theta$ to be a point in parameter space around which $\lik(\theta,\phi,\omega, J)$ is differentiable with respect to $\theta$ and does not have excessive Monte~Carlo variance.

We require many of these assumptions to establish the following proposition, the proof of which is a straightforward consequence of what we assumed.

\begin{prop}[Smooth Neighborhood]
\label{prop:smooth-nbhd}
There exists an open neighborhood $\mathcal{N}(\phi)$ around any $\phi \in \Theta$ where for all $\theta \in \mathcal{N}(\phi)$ and almost every $\omega \in \Omega$, the Monte Carlo estimate of the likelihood at $\theta \in \mathcal{N}(\phi)$ with the system evolving according to $\phi$ conditional on $\omega$, $\lik(\theta,\phi,\omega, J)$, has two continuous derivatives in $\theta$.
\end{prop}

\begin{proof}
    For a fixed $\omega$, the resampling indices for $\lik(\theta,\phi,\omega, J)$ and $\lik(\phi,\phi,\omega, J)$ are the same as the resampling indices in both cases correspond to that yielded by $f_{Y_n|X_n}(y_n^*|x_{n,j}^{P,\phi}(\omega);\phi)$, eliminating any discontinuities from resampling.

    By Assumptions \ref{assump:diff-meas} and \ref{assump:diff-process}, for small enough $\mathcal{N}(\phi)$ the density ratios are bounded. The likelihood only changes by a factor of the density ratios, which are bounded and possess two continuous derivatives in $\theta$ if either the process and measurement densities used in the calculation do, or if the measurement densities do and one has access to a differentiable simulator with two continuous derivatives in $\theta$. This gives us the desired result.

    As a remark, we note that the neighborhoods the sequence is differentiable in are independent of $J$, as the resampling does not change with $\theta$ -- it is dependent on $\phi$. 
\end{proof}

\subsubsection{Doubly Off-Policy Particle Filter}

Having the system evolve according to $\phi$ but evaluating the likelihood at $\theta$ requires us to reweight the conditional likelihood at each timestep by likelihood ratios corresponding to the process and measurement models. This corresponds to making two corrections in the particle weights at each timestep, by multiplying by 
\begin{equation}
    s_{n,j}=\frac{g_{n,j}^\theta}{g_{n,j}^{\phi}}=\frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi}; \theta)}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P,\phi}; \phi)}, \qquad r_{n,j}=\frac{f_{X_n|X_{n-1}}(x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \theta)}{f_{X_n|X_{n-1}}(x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \phi)},
\end{equation}
as we illustrate in Appendix \ref{app:dop}. We use the notation $x_{n,j}^{P, \phi}$ to emphasize that the particles evolve according to $\phi$.

We claim that this yields a particle filter that targets the posterior and recovers the derivative estimate of \citet{poyiadjis11, scibior2021dpf} when $\theta$ is evaluated at $\phi$. The proof can be found in Appendix \ref{app:dop}.

\begin{prop}[Correctness of Doubly Off-Policy Particle Filter Corrections]
    \label{prop:dop-correctness}

    Let the system evolve according to $\phi \in \Theta,$ that is, with the process model $\process(\cdot | \phi)$ and resampling with probabilities proportional to $f_{Y_n|X_n}(y_n^* | x_{n,j}^{P,\phi})$. Also let $\theta \in \gN(\phi)$. Then, the following corrections in the particle weights,
    \begin{equation}
    s_{n,j}=\frac{g_{n,j}^\theta}{g_{n,j}^{\phi}}=\frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi}; \theta)}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P,\phi}; \phi)}, \qquad r_{n,j}=\frac{f_{X_n|X_{n-1}}(x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \theta)}{f_{X_n|X_{n-1}}(x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \phi)},
    \end{equation}
    yield a particle filter that is unbiased for the likelihood, even when $\theta \neq \phi$.

    In particular, when $\theta$ is evaluated at $\phi$, its gradient is the same as the estimator of \citet{poyiadjis11},
    \begin{equation}
        \nabla_\theta \hat\ell(\theta) = \frac{1}{J}\sum_{j=1}^J \nabla_\theta \log f_{X_{0:N}, Y_{1:N}}(x_{1:N,j}^{A, F,\theta}, y_{1:N}^*).
    \end{equation}
\end{prop}

\paragraph{Intuition:} Having the system evolve according to $\phi$ but evaluating the likelihood at $\theta$ requires us to reweight the conditional likelihood at each timestep by likelihood ratios corresponding to the process and measurement models. This corresponds to making two corrections in the particle weights at each timestep, one for the process model and one for the measurement model. The measurement correction cannot be avoided, but one can bypass the process correction -- with the reparametrization trick. We can do when we assume the simulator is differentiable with the reparametrization trick, as \cite{corenflos21} also assume. Doing so then yields an algorithm that only requires a simulator, and does not require the process densities.

\subsection{Measurement Off-Policy Particle Filters}

With the above procedure, we have the entire system evolve according to $\phi$, and we differentiate through the likelihood ratios, which are smooth conditional on the resampling being the same (which in turn happens when $\theta \approx \phi$). However, doing so requires access to likelihood ratios, and as such Algorithm \ref{alg:dop} is therefore not compatible with simulation-based inference.

There is a workaround. Consider the case where you have access to a differentiable simulator compatible with the reparametrization trick. If we have the state transitions evolve according to $\theta,$ but have the resampling according to $\phi,$ we can offload the derivative with respect to the process model to the reparametrization trick, as in \cite{corenflos21}. This corresponds to the MOP-$\alpha$ procedure in Algorithm \ref{alg:mop}.

We therefore make the following assumption in this section only:
\begin{aspt}[Differentiable Simulator]
    $\texttt{process}(\cdot | X_{t-1}; \theta)$ is differentiable with respect to $\theta,$ and is compatible with the reparametrization trick. 
\end{aspt}

Algorithm \ref{alg:mop} provides a weighted particle filter that targets the posterior under $\theta$, but when differentiated yields either the low-bias but high-variance gradient estimator of \cite{poyiadjis11}, the high-bias but low-variance estimator of \cite{blei2018vsmc}, or an intermediate, interpolating quantity that can optimize a bias-variance tradeoff. This is controlled by a discounting parameter $\alpha$. The intuition behind this procedure is as follows.

\paragraph{Intuition:} Broadly, if $\theta\neq\phi,$ one can first run a vanilla particle filter to recover values for the particles with the process model under $\phi.$ One can then run an accordingly reweighted particle filter with the process model under $\theta$, but resampling according to the measurement model under $\phi$, with the same indices as the original vanilla particle filter. This requires one to set the same random seed for both particle filters. This then yields a slightly-unconventionally weighted particle filter that targets the posterior, but has some very nice gradient properties; as we outline in the remainder of this section.

\paragraph{Special Case:} When $\theta = \phi$, one can simply set the particles under $\phi$ to be copies of the particles under $\theta$. Gradients with respect to $\theta$ now treat functions of the particles under $\phi$ as constants, corresponding to the stop-gradient trick in \cite{scibior2021dpf}. This therefore provides a theoretical framework for their stop-gradient trick, which we happily use in our numerical experiments for simplicity.

\kevin{Develop theory for mixing-time swapping? Detect when IF2 reaches a neighborhood of the MLE, by detecting whether it has reached a local sort of stationary distribution when the perturbations are lower bounded, e.g. the autocorrelation between the mean vectors to estimate the mixing time or Gelman-Rubin. In practice, just swap when IF2 converges according to the diagnostics. Theory doesn't care, you just need to in a basin of attraction that is strongly convex.}

\subsubsection{Discounting Weights}

Observe that at Algorithm \ref{alg:mop}, at each timestep the corrected filtering importance weights $w_{n,j}^{F,\theta}$ optionally propagate to the next timestep for us to maintain the correction. We introduce a discounting parameter $\alpha$ that affects the weights, so $w_{n,j}^{P,\theta} := (w_{n-1,j}^{F,\theta})^{\alpha}$. Note that no matter what value $\alpha$ takes on, the system evolves according to $\phi$ when we perform the correction to evaluate the likelihood at $\theta$. The filter fully targets the posterior when $\theta=\phi$, or $\alpha=1$, as we show in Proposition \ref{prop:dop-correctness}. As previously mentioned, the value of $\alpha$ therefore informs a bias-variance tradeoff. We elaborate on this below.

\paragraph{Bias-Variance Tradeoff} This discounting serves the purpose of balancing a tradeoff between maintaining memory of each particle's ancestral trajectory (most extreme when $\alpha=1$), or considering only the single-step transition dynamics (when $\alpha=0$). This can be thought of as an \textbf{exponentially-weighted moving average}, where $\alpha$ controls the amount of discounting. We elaborate on two special cases below.

\kevin{This isn't a smoother, because we'd need future information.}

\paragraph{When $\alpha=1,$} the likelihood estimate becomes
\begin{equation}
    \hat{\lik}(\theta) := \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n,j}^{P,\theta}} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n-1,j}^{F,\theta}} = \left(\frac{1}{J}\sum_{j=1}^J w_{N,j}^{F,\theta}\right) \prod_{n=1}^N L_n^\phi,
\end{equation}
as we have a telescoping product. Its derivative when $\theta=\phi$ is then
\begin{equation}
    \nabla_\theta \hat{\ell}(\theta) := 
        \frac{1}{J}\sum_{j=1}^J \nabla_\theta \log f_{X_{0:N}, Y_{1:N}}(x_{1:N,j}^{A, F,\theta}, y_{1:N}^*),
\end{equation}
the average gradient of each surviving particle's cumulative importance weights. The filter targets the posterior when $\alpha=1$ even when $\theta \neq \phi,$ and therefore suffers from no asymptotic bias.

\paragraph{When $\alpha=0,$} the likelihood estimate becomes
\begin{equation}
    \hat{\lik}(\theta) := \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n,j}^{P,\theta}} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J w_{n,j}^{F,\theta} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J \frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \theta})}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi})},
\end{equation}
and though the telescoping product disappears, this reduces to a series of one-step corrections. However, these corrections are insufficient when $\theta \neq \phi,$ and the filter suffers from asymptotic bias unless $\theta=\phi.$ Its gradient when $\theta=\phi$ is therefore

    \begin{equation}
        \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta),
    \end{equation}

the gradient of the vanilla particle filter where the resampling correction is omitted as found in \cite{blei2018vsmc} -- when applied to the bootstrap filter. As \cite{scibior2021dpf} remark, this estimator only considers single-step quantities, and is "memoryless" beyond a single step. This is in contrast to the case when $\alpha=1,$ as that estimate only considers the surviving particles at time $N$, which is the other extreme where dependencies across time are fully tracked.

\subsubsection{Correctness of MOP-$\alpha$}

\begin{prop}[Correctness of MOP-$\alpha$ (Algorithm \ref{alg:mop})]
    \label{prop:mop-correctness}
    When either $\alpha=1$ or $\theta=\phi$, Algorithm \ref{alg:mop} is a particle filter that targets the posterior and is unbiased for the likelihood under $\theta$.

    In particular, when $\theta$ is evaluated at $\phi$, the likelihood estimate agrees with the bootstrap filter. In this case, its gradient when $\alpha=1$ is the same as the DPF estimator of \citet{scibior21} when applied to the bootstrap filter,
    \begin{equation}
        \frac{1}{J}\sum_{j=1}^J \nabla_\theta \log f_{Y_{1:N}|X_{0:N}}(y_{1:N}^* | x_{1:N,j}^{A, F,\theta}),
    \end{equation}

    and when $\alpha=0,$ is the gradient estimator of \cite{blei2018vsmc},

    \begin{equation}
        \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta),
    \end{equation}

    that as \cite{scibior21} show corresponds to differentiating through the fixed-seed likelihood estimate yielded by a vanilla particle filter, dropping resampling terms.

    When $\alpha=1$, the gradient estimate MOP-1 provides is consistent for the score.
\end{prop}

The proof for the above follows from Appendices A and C, as well as the technical lemmas for properly weighted particle filters.

\kevin{Take derivative of loglik with regard to alpha?}
\kevin{Talk about regime change the moment you move away from alpha=1, variance going from N2 to N, and bias going from 0 to something, but try to prove that decrease in variance outweights increase in bias. }
\kevin{Create tools for choosing alpha via binary search CV from an averaged alpha=1 target}
\kevin{suppose at each filtering step, look at gradient calculated at alpha=0, dynamically select alpha during filtering. calculate bias and variance online during filtering, and update alpha accordingly. }

\kevin{No need to recompute so many particle filters! Just take an array of weight copies, one for each alpha. Gradients for each alpha come free the moment you compute one of them. }

\kevin{Elbow method on variance, since bias is hard to compute? }
\kevin{Get variance estimate, bias estimate, construct confidence intervals on bias estimate, use that to get unbiased estimate of the MSE. }

\subsubsection{Practical Considerations}

It is only ever practical to evaluate $\theta$ at $\phi$. This removes the issue of the likelihood ratios blowing up, and yields an estimate of the likelihood that agrees with the bootstrap filter estimate. Its gradient in this case is either the estimate of \citet{poyiadjis11} when $\alpha=1$, a single-timestep estimator that corresponds to naively taking the derivative of the particle filter when $\alpha=0$ as first proposed in \citet{blei2018vsmc}

\paragraph{Bottom Line:} The DOP-$\alpha$ and MOP-$\alpha$ procedures therefore provide us with a unifying framework for reasoning about smooth extensions of weighted particle filters that target the posterior and provide reasonable gradient estimates when differentiated. This fixed-seed reweighting alleviates issues that can arise from differentiating through a quantity yielded by ostensibly non-differentiable resampling. This is done by either (1) differentiating through smooth likelihood ratios, or (2) differentiating through a differentiable simulator with the reparametrization trick (and through smooth likelihood ratios to adjust for the resampling). In particular, MOP-$\alpha$ does not require access to the transition densities, and like the bootstrap filter, only requires access to a (albeit differentiable) simulator. To our knowledge, this is the first particle filter gradient method in the literature with this capability.

\section{Practical Maximum-Likelihood Inference}

This still leaves us with the question of designing a reasonably effective procedure for likelihood maximization with these new tools. We propose a simple algorithm we call Iterated Filtering with Automatic Differentiation (IFAD) in Algorithm \ref{alg:ifad} that amounts to warm-starting gradient descent or some other reasonable first or second order procedure (using the gradient estimate given by MOP-$\alpha$) with the output of an initial search of IF2.

\paragraph{Summary:} A coarse solution is found by running IF2 with an aggressive random walk standard deviation (learning rate) and a cooling schedule that is \textbf{bounded below}, until the search stalls to an initial "convergence" to what is hopefully a neighborhood of the MLE. We can then use gradient descent, or other first or second order gradient methods, to refine this coarse solution.

The initial "convergence" happens fairly quickly in practice. In the case of the Dhaka cholera model of \cite{king08}, when a geometric cooling multiplier of 0.95 and initial random walk standard deviation of 0.02 is used, initial convergence happens within 40 iterations of IF2. In comparison, often 100 or 200 iterations of IF2 are used for an initial global search. \cite{ionides15}, for example, use 100 iterations with the Dhaka cholera model of \cite{king08}, while \cite{wheeler23} use 200 with Model 1 in \cite{Lee_haiticholera}.

\paragraph{Rationale:}

IF2 converges quickly to a neighborhood of some reasonably good local optimum in practice, but then struggles to squeeze out the last few log-likelihood units. Conversely, gradient descent (and other first or second order gradient methods) can get stuck in saddle points and less desirable local minima when the likelihood is nonconvex. Warm-starting gradient methods with IF2 lets us get close enough to the MLE for gradient methods to shine, as in most cases regularity conditions ensure that the likelihood surface is well-behaved enough in this neighborhood that the aforementioned issues with saddle points and local minima are alleviated. Combining these two methods in this way lets us enjoy the best of both worlds.

\begin{algorithm}[ht]
\centering
	\caption{Iterated Filtering with Automatic Differentiation}
    \label{alg:ifad}
	\begin{algorithmic}[1]
	     \STATE \textbf{Input:} Number of particles $J$, timesteps $N$, measurement model $p(y_n^*|x_n, \theta)$, simulator $\process(x_{n+1}|x_n, \theta),$ IF2 cooling schedule $\eta_t$, MOP-$\alpha$ discounting parameter $\alpha$.
		\STATE Initialize $\theta_0$, $t=0.$
        \STATE Run IF2 until initial "convergence" under cooling schedule $\eta_t$ to obtain $\{\Theta_j, j=1,...,J\}$, set $\theta_t := \frac{1}{J}\sum_{j=1}^J \Theta_j.$
		\WHILE{Procedure not converged:}
		\STATE Draw $\omega \in \Omega$
		\STATE Perform forward pass by running Algorithm \ref{alg:mop} with discounting parameter $\alpha$ to obtain $\loglik(\theta_t, \theta_t, \omega, J).$
		\STATE Obtain $g(\theta_t) = \nabla_{\theta_t} (-\loglik(\theta_t, \theta_t, \omega, J))$ via backpropagation, $H(\theta_t)$ such that $\lambda_{\min}(H) \geq c$.
		\STATE Obtain $\eta$ with line search or an annealing schedule.
		\STATE Update $\theta_{t+1} := \theta_t - \eta (H(\theta_t))^{-1} g(\theta_t)$, $t:=t+1.$
		\ENDWHILE
		\STATE Return $\hat{\theta} := \theta_t.$
	\end{algorithmic}
\end{algorithm}

\subsection{Theoretical Guarantee}

We first analyze the performance of gradient descent, Newton's method, or any iterative method that multiplies the particle estimate of the gradient with the inverse of a matrix with minimum eigenvalue bounded below, in the case where the log-likelihood surface is $\gamma$-strongly convex, and where the gradient estimate is a particle approximation of the score (for example, IFAD-1 and MOP-1). In this case, these methods enjoy a linear convergence guarantee:

\kevin{I think we can establish the linear convergence for the entire procedure in the same sense as \href{this}{https://arxiv.org/pdf/2311.08442.pdf}, in the sense that if IF2 is run for a sufficiently large but fixed number of iterations, once we arrive in a local strongly convex region, we get linear convergence when we swap to MOP. Show by induction that we remain in the strongly convex region (in terms of parameter space and likelihood), eqn 218. }

\begin{thm}[Convergence of Gradient Methods]
\label{thm:convergence}

Consider a variant of IFAD-1 (Algorithm \ref{alg:ifad}) where one stops if the gradient estimate $||g(\theta_n)|| \leq \sigma \epsilon$. Assume $-\ell$ is $\gamma$-strongly convex, $\gamma I \preceq \nabla_\theta^2 (-\ell) \preceq \Gamma I$. $\sigma = \frac{4 \Gamma}{(1-\beta)}> 4$. Define $G$ as in Assumption \ref{assump:local-bounded-derivative}.

Choose $J$ according to the maximum of the $J$ prescribed in Lemmas \ref{lemma:grad_bound} and (if using a Hessian estimate) $\ref{lemma:hess_bound}$ with failure probability $\delta/2$ each. Let $H(\theta_n)$ be a matrix possibly dependent on $\theta_n$ with minimum eigenvalue always greater than some $c>0$. Choose the learning rate $\eta$ and error tolerance $\epsilon$ such that, where $\beta$ is the Armijo condition hyperparameter,
\begin{equation}
    \eta \leq \frac{c(1-\beta)}{\Gamma}, \;\; \epsilon \leq \frac{c(1-\beta)}{2\Gamma}||g(\theta_n)||.
\end{equation}

Then the following holds with probability at least $1-\delta$:
\begin{equation}
\ell(\theta^*) - \ell(\theta_{n+1}) \leq \left(1-\eta\beta\frac{8\gamma}{9c}\right)(\ell(\theta^*)-\ell(\theta_n))
\end{equation}

and the algorithm eventually terminates after we obtain $||\nabla_\theta \ell(\theta_n)|| \leq (1+\sigma) \epsilon$.
\end{thm}

The proof of this result uses the concentration inequalities from \cite{delmoral2011ci}, and leans upon the analysis of \cite{mahoney16}. We defer the proof to Appendix \ref{app:ifad}

\paragraph{Takeaways for IFAD (Algorithm \ref{alg:ifad}):}

The gradient stage of IFAD-1 converges linearly to the MLE if (1) the log-likelihood surface is $\gamma$-strongly convex in a neighborhood of the MLE and (2) the IF2 stage of IFAD successfully reaches a (high-probability) basin of attraction of the MLE. This happens fairly often in practice, for example, when sufficient regularity conditions for local asymptotic normality of the MLE hold.

We conjecture that this applies to the entirety of IFAD, as IF2 converges very quickly to a neighborhood of the MLE. IF2 posesses some SGD-like qualities, including what looks like linear convergence to a ball around the MLE with width determined by the random walk standard deviation. We will explore this conjecture in future work.

While this theoretical guarantee currently only holds for IFAD-1, as it forms a particle approximation of the score as in \cite{poyiadjis11}, one can also extend this for a similar guarantee if the bias for the gradient estimate given by MOP-$\alpha$ for $\alpha < 1$ is small enough. For general $\alpha$, including MOP-0, the analysis then becomes more complicated, as one needs to handle the case of biased gradient descent. However, if future measurements $y_{n+1:N}^*$ do not provide a large amount of information on the identification of the current state $x_n$ given past and current measurements $y_{0:n}^*$, as mentioned in \cite{corenflos21}, the bias may be small enough for it to not present too much of a problem.

\subsection{Runtime}

Even with the increased computational cost of automatic differentiation, Algorithm \ref{alg:ifad} is reasonably fast, as by the cheap gradient principle from \cite{kakade2019provably}, a gradient operation takes no more than 6 times the runtime of the particle filter (in practice, this means a gradient step takes about 3.75x as much time as an IF2 iteration). As such, unlike methods like the marginal differentiable particle filter in \cite{scibior2021dpf} or the the optimal transport of resampling approach of \cite{corenflos21}, our algorithm has runtime \textbf{linear, and not quadratic}, in the number of particles (while alleviating the $O(N^2)$ variance concern of \cite{poyiadjis11} and \cite{corenflos2021ot} and asymptotic bias concern of \cite{blei2018vsmc} through optimizing $\alpha$ for a bias-variance tradeoff).

In addition, re-implementation of the particle filter, iterated filtering, and Algorithm \ref{alg:ifad} itself in \texttt{JAX} (\cite{jax2018github}) led to considerable speedups. This can be attributed to both GPU acceleration and the just-in-time (JIT) compilation functionality that \texttt{JAX} provides. When tested on a computer with an Intel i9-13900K processor, 128GB of RAM, and a NVIDIA RTX3090 GPU, each post-compilation particle filtering operation on the Dhaka cholera model from \cite{king08} with 10000 particles took roughly 379 milliseconds (with a standard deviation of 279 microseconds over seven runs), a considerable improvement over the existing performance in the widely used \texttt{pomp} package of \cite{king16}, which took between 6.19 to 6.39 seconds over ten runs on the same machine.

\section{Numerical Experiments}

\subsection{Cholera in Bangladesh}

To compare the performance of IFAD against that of IF2, we fit the very same model first proposed in \cite{king08} that \cite{ionides15} use to demonstrate the capabilities of IF2. This is becoming a standard benchmark, as \cite{wood16} and \cite{wycoff2024voronoi} include in their papers. This is a stochastic SIR compartmental model with transition and measurement uncertainty, where the population at time $t$, $H(t)$, is divided into the susceptible compartment $S(t)$, infected compartment $I(t)$, and three ($k=3$) recovered compartments $R_1(t), ..., R_k(t)$ denoting varying degrees of cholera immunity. We write $M(t)$ for the cholera deaths in each month. As in \cite{ionides15}:

\begin{itemize}
    \item The \textbf{transition dynamics} follow this series of stochastic differential equations:

    \begin{align}
        d S&=\left(k \epsilon R_k+\delta(S-H)-\lambda(t) S\right) d t+d H-(\sigma S I / H) d B, \\ 
        d I&=\left(\lambda(t) S-(m+\delta+\gamma) I\right) d t+(\sigma S I / P) d B, \\ 
        d R_1&=\left(\gamma I-(k \epsilon+\delta) R_1\right) d t, \\ 
        \vdots \\ 
        d R_k&=\left(k \epsilon R_{k-1}-(k \epsilon+\delta) R_k\right) d t,
    \end{align}

    where $B$ represents Brownian motion, $m$ is the rate at which infected individuals succumb to cholera, $\gamma$ is the rate at which infected individuals recover, $1/\epsilon$ is the mean duration of immunity, $\sigma$ represents the standard deviation of perturbations to the force of infection, and $\delta=0.02$ is the death rate.

    \item The \textbf{force of infection}, $\lambda(t)$, is modeled by splines $s_j$
    \begin{equation}
        \lambda(t)=\exp \left\{\beta_{\text {trend }}\left(t-t_0\right)+\sum_{j=1}^{6} \beta_j s_j(t)\right\}(I / P) + \exp \left\{\sum_{j=1}^{6} \omega_j s_j(t)\right\},
    \end{equation}
    where the $\beta_j$ model seasonality in the force of infection, $\beta_{\text{trend}}$ models the trend in the force of infection, and the $\omega_j$ represent seasonality of a non-human environmental reservoir of disease. 
    \item The \textbf{measurement model} for observed monthly cholera deaths is given by 
    \begin{equation}
        Y_n \sim \gN(M_n, \tau^2M_n^2),
    \end{equation}
    where $M_n$ is the true number of cholera deaths in that month.
\end{itemize}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[node distance=1cm and 1cm, auto,
>=Latex,every node/.append style={align=center}, 
int/.style={draw, minimum size=1.6cm}]
  \node [int] (H) at (0,0) {$H$};
\node [int] (S) [right=of H] {$S$};
\node [int] (I) [right=of S] {$I$};
\node (Mcoord) [right=of I, coordinate] {Mcoord};
\node [int] (R1) [right=of Mcoord] {$R_1$};
\node [int] (R2) [right=of R1] {$R_2$};
\node [int] (R3) [right=of R2] {$R_3$};
\node [circle, draw] (M) [above=of Mcoord] {$M$};
\node (deathsS) [above=of S, coordinate] {deathsS};
\node (deathsI) [above=of I, coordinate] {deathsS};
\node (deathsR1) [above=of R1, coordinate] {deathsR1};
\node (deathsR2) [above=of R2, coordinate] {deathsR2};
\node (deathsR3) [above=of R3, coordinate] {deathsR3};
\node (belowR3) [below=of R3, coordinate] {belowR3};
\node (belowS) [below=of S, coordinate] {belowS};
\path[->]
    (H) edge node {$dH$} (S)
    (S) edge node {$\lambda(t)$}  (I)
    (I) edge node {$\gamma$}  (R1)
    (R1) edge node {$k\epsilon$}  (R2)
    (R2) edge node {$k\epsilon$}  (R3)
    (R3) edge node {} (belowR3) 
    (belowR3) edge node {$k\epsilon$} (belowS) 
    (belowS) edge node {} (S)
    (I) edge node {$m$}  (M)
    (S) edge node {$\delta$}  (deathsS)
    (I) edge node {$\delta$}  (deathsI)
    (R1) edge node {$\delta$}  (deathsR1)
    (R2) edge node {$\delta$}  (deathsR2)
    (R3) edge node {$\delta$}  (deathsR3);
\end{tikzpicture}
\caption{Illustration of SIR model from \cite{king08}.}
\label{fig:sir}
\end{figure}

Figure \ref{fig:sir} depicts the above compartmental model that we use to benchmark the performance of IFAD. We present comparisons of IFAD with $\alpha=0$ (IFAD-0) and $\alpha=1$ (IFAD-1) against both (1) the results from \cite{ionides15}, and (2) our own implementation of IF2 in \texttt{JAX}.

\paragraph{Hyperparameters:} While performing an initial exploratory analysis, we found that alternative hyperparameters for IF2 (random walk standard deviation of $\sigma=0.02$, geometrically decreasing cooling rate exponent $a=0.95$) performed better than the settings \cite{ionides15} chose, and chose to use these for our implementation to achieve a fairer comparison. Our implementations of IF2 and IFAD used 10000 particles, in line with \cite{ionides15}.

To prevent exploding gradients given the high variance in the MOP-1 estimator, we scale the gradient to have norm 1 before performing vanilla gradient descent, and use a linear cooling schedule of $0.2$ to $ 0.0001$. We chose to use vanilla gradient descent for simplicity, and to demonstrate that minimal tuning was required to have IFAD perform well.

% The 60 more iterations of IF2 yielded largely not a huge improvement over the initial coarse solution of 40 iterations. This demonstrates what we have mentioned earlier: that although IF2 manages to quickly find a decent coarse solution, it struggles to squeeze out those last few units of log-likelihood improvement. The 60 more iterations of gradient descent, however, successfully refines this solution, getting much closer to the MLE on average.

\subsubsection{Results} 
We benchmarked IFAD against IF2 on a challenging global search problem. 100 initial starting parameter vectors were drawn uniformly from the same wide bounding box used in \cite{ionides15}, and we performed 100 searches each of IF2, IFAD-0, IFAD-1, MOP-0, and MOP-1 initialized from these 100 starting parameter vectors. We summarize our findings below.

\begin{table}[h!]
\centering
\begin{tabular}{||c c c||} 
 \hline
 Method & Best Log-Likelihood & Rank \\ [0.5ex] 
 \hline\hline
     IFAD-0 & -3747.87 & 1\\
     IFAD-1 & -3748.53 & 2\\
     IF2 (Ours) & -3755.91 & 3\\
     IF2 (Warm Start) & -3759.60 & 4 \\
     IF2 2015 & -3768.63 & 5\\ 
     MOP-0 Alone & -3795.74 & 6\\
     MOP-1 Alone & -3797.38 & 7\\
 \hline
\end{tabular}
\caption{Maximum log-likelihood found by IF2, IFAD, and MOP alone. IFAD performs the best among all methods. Our implementation of IF2 outperforms that of \cite{ionides15}, but still ultimately underperforms IFAD. Both IFAD-0 and IFAD-1 manage to find the MLE, either matching or beating the highest log-likelihood previously found in the \texttt{dacca()} model implemented within the \texttt{pomp} package of \cite{king16}.}
\label{table:mle}
\end{table}

\paragraph{IFAD Successfully Finds the MLE:} Both IFAD-0 and IFAD-1 successfully find the MLE, as seen in Table \ref{table:mle}. Previously, the MLE at a log-likelihood of $-3748.5$ was found with much effort, using many global and local IF2 searches and likelihood profiling. Despite being initialized for a global search, both IFAD-0 and IFAD-1 manage to find the MLE over the 100 searches. This may imply that the local search and refinement that was previously necessary for finding the MLE is \textbf{baked into} the two-stage IFAD algorithm.

Interestingly, \cite{ionides15} only achieve a maximum log-likelihood of $-3768.63$, while \cite{king08} only achieve a best possible estimate of $-3793.4$. Our results of $-3748.53$ and $-3747.87$ match or beat the highest log-likelihood previously found in the current implementation in the \texttt{dacca()} model implemented within the \texttt{pomp} package of \cite{king16}.

This can be seen most clearly in Figure \ref{fig:boxplot-search}. In the left panel within Figure \ref{fig:boxplot-search}, only IFAD has whiskers that extend to the dotted red line that represents the MLE.

\paragraph{IFAD Outperforms Both IF2 and MOP Alone:} While IF2 quickly approaches a neighborhood of the MLE within only 40 iterations (as seen in Figure \ref{fig:optim}), performing IF2 alone ultimately fails to squeeze out the last few log-likelihood units, as no IF2 search comes within 7 log-likelihood units of the MLE (as seen in Figure \ref{fig:scatter} and the left panel in Figure \ref{fig:boxplot-search}).

Conversely, performing MOP alone leads to many, many failed searches, as seen in Figure \ref{fig:hist-all} and the right panel in Figure \ref{fig:boxplot-search}. This is a difficult, nonconvex, and noisy problem, and the search gets stuck in local minima and saddle points, failing to approach the MLE.

IFAD, in comparison, successfully manages to (1) approach the MLE quickly due to the IF2 warm-start (as seen in Figure \ref{fig:optim}) and (2) refine the coarse solution found by the warm-start with MOP gradient steps to find the MLE (as seen in Figures \ref{fig:scatter}, \ref{fig:boxplot-search}, and \ref{fig:hist-all}). IFAD therefore successfully combines the best qualities of IFAD and MOP, outperforming either of them alone.

\begin{figure}[htbp!]
    \centering
    \includegraphics[scale=0.53]{../imgs/095/pairs.png}
    \includegraphics[scale=0.53]{../imgs/095/qq.png}
    \caption{Scatterplots depicting the performance of IFAD against that of IF2. \textbf{Left:} Paired searches from the same starting point. \textbf{Right:} Q-Q plot of ranked IFAD searches against ranked IF2 searches. It is clear that on average, IFAD has the edge, and manages to find the MLE while no IF2 search successfully gets within 7 log-likelihood units of it. We use a dotted red line to display the MLE in both. }

    \label{fig:scatter}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.48]{../imgs/095/boxplot.png}
    \includegraphics[scale=0.53]{../imgs/095/boxplotall.png}
    \caption{\textbf{Left:} Boxplot depicting the performance of IFAD and IF2 where we plot the results of the best run out of every ten runs. This simulates the result of the common procedure of running a few searches and choosing the best one. IFAD outperforms all other methods, and improves on the warm-start given by running 40 IF2 iterations. Our choice of hyperparameters for our implementation of IF2 outperforms the choices made by \cite{ionides15}, which we depict in the boxplot labeled "IF2 2015". As before, while IF2 successfully approaches the MLE, it ultimately fails to find it.
    \textbf{Right:} Boxplot depicting the performance of IFAD, IF2, and MOP alone without the warm-start, where we plot all runs. We see that MOP alone drastically underperforms all other methods, failing to get close to the MLE. This shows that the IF2 warm-start may be necessary to achieve good performance in challenging nonconvex and noisy problems like this one.
    \textbf{Takeaway:} In this panel, we see that (1) IF2 fails to find the MLE while the gradient steps perform better at fine-grained refinement, and (2) performing gradient steps alone without a warm-start leads to the search getting stuck in local minima and saddle points, ultimately demonstrating the necessity of a hybrid algorithm like IFAD.
    In both, as before, we use a dotted red line to display the MLE. }
    \label{fig:boxplot-search}
\end{figure}

\begin{figure}[htbp!]
    \centering
    \includegraphics[scale=0.7]{../imgs/095/optim.png}
    \caption{Optimization progress of IFAD and IF2, where the solid lines depict the median negative log-likelihood at each iteration and the shaded area depicts the best search at any iteration and the 80\% percentile. The dashed orange line depicts the median warm-start given by running 40 IF2 iterations. While running 60 more iterations of IF2 improves upon the median warm-start, doing so ultimately underperforms IFAD -- IFAD has better tail control and successfully reaches the MLE. The initial loss in log-likelihood upon transitioning from IF2 to MOP seems to help with escaping local minima. 
    As before, we use a dotted red line to display the MLE. }
    \label{fig:optim}
\end{figure}

\begin{figure}[t!]
    \centering
    \includegraphics[scale=0.7]{../imgs/095/hist.png}
    \caption{Histogram comparison of the performance of IFAD, IF2, and MOP alone. As before, we note that IFAD successfully reaches the MLE a nontrivial number of times, while IF2 ultimately fails to do so. Our implementation of IF2 outperforms that of \cite{ionides15}, largely due to improved hyperparameter choices. MOP alone performs poorly, demonstrating the necessity of warm-starting the search with IF2. 
    We once again see that IFAD combines the best of IF2 and MOP, approaching the MLE quickly due to the IF2 warm-start and successfully refining the coarse solution with gradient steps, achieving superior performance. As before, we use a dotted red line to display the MLE. }
    \label{fig:hist-all}
\end{figure}

\paragraph{Differences Between IFAD-0 and IFAD-1:} The differences between IFAD-1 and IFAD-0 (MOP-1 and MOP-0) do not seem very pronounced in this example. This could be due to a number of reasons, such as (1) future case counts $y_{{n+1}:N}^*$ not providing a large amount of information on state identification $x_n$ given current and past measurements $y_{0:n}^*$ as mentioned in \cite{corenflos21}, and (2) the gradient normalization making the trajectories given by MOP-0 and MOP-1 more similar than they would be otherwise.

\kevin{IFAD beats IF2 x percent of the time on 100 paired runs, but the bigger thing is that it gets close to the MLE without much effort. Previously, we'd have to do many searches, plus refinement (which is necessary, as we can see from the IF2 boxplot, it is easy to fool yourself into thinking that because you see the plateau that is the MLE). This includes likelihood profiling diagnostics, which are very computationally intensive. Now we can just take the best of ten searches -- IFAD does the refinement.}

\kevin{Have an outsider read the manuscript afterwards.}

\section{Discussion and Future Work}

We develop a new theoretical framework to differentiating through the particle filter that (1) encompasses the gradient estimates of \cite{poyiadjis11}, \cite{blei2018vsmc}, and \cite{scibior2021dpf} as special cases, (2) sidesteps the issue of differentiating through a Monte-Carlo algorithm with discontinuous resampling, and (3) provides an opportunity for optimizing a bias-variance tradeoff.

Prior to this work, iterated filtering algorithms were the only class of algorithms for maximum likelihood estimation that did not require access to the system's transition probabilities, instead needing only a simulator of the system dynamics. In the presence of a differentiable simulator, MOP-$\alpha$ enjoys this very same desirable property, enabling practical maximum likelihood inference in a wider variety of situations.

Furthermore, MOP-$\alpha$, and by extension IFAD, has runtime \textbf{linear, and not quadratic}, in the number of particles, unlike the marginal differentiable particle filter in \cite{scibior2021dpf} or the the optimal transport of resampling approach of \cite{corenflos21}. This does not necessarily come at a cost to Monte Carlo variance, unlike \cite{poyiadjis11} and \cite{corenflos2021ot} who suffer from variance quadratic in the length of the trajectory, as we can optimize $\alpha$ for a bias-variance tradeoff.

In the presence of strong convexity, first and second-order gradient methods that use MOP-$\alpha$ as an estimate enjoy a linear convergence rate, providing support for numerical optimization with AD for the particle filter.

This leads to a practical hybrid algorithm we call IFAD that warm-starts gradient descent, or some other first or second-order method using the gradient estimate given by MOP-$\alpha$, with IF2. When benchmarked on a challenging global search problem with the model of \cite{king08}, IFAD enjoys the best properties of IF2 and MOP-$\alpha$, successfully approaching a neighborhood of the MLE quickly without getting stuck in saddle points and (poor) local minima (like gradient descent alone would), while effectively optimizing the last few units of log-likelihood (which IF2 struggles with).

Finally, we make the first steps towards developing a Python counterpart to the popular \texttt{pomp} R package by \citet{king16, king2017pompmanual} that we call \texttt{pomp-py}. The software we provide includes welcome features for scientific computing such as GPU acceleration, parallel computing for single-program-multiple-data programs, just-in-time compilation, and automatic differentiation enabled by the \texttt{JAX} Python package from \citet{jax2018github}.

\subsection{Future Work}

While MOP-$\alpha$, and IFAD in its current incarnation, remain very general, they currently lack the capability of handling discrete latent states. We conjecture that DOP-$\alpha$ might be a possible solution (despite needing the likelihood ratio of transition densities), and will explore this in future work.

We conjecture that the linear convergence result in Theorem \ref{thm:convergence} applies to the entirety of IFAD, as IF2 converges very quickly to a neighborhood of the MLE. IF2 posesses some SGD-like qualities, including what looks like linear convergence to a ball around the MLE with width determined by the random walk standard deviation. Not much is known about the convergence rate of IF2, and an answer to this would also provide an answer to the convergence rate of IFAD beyond just the convergence rate of the gradient stage. We will explore this conjecture in future work.

In the interest of simplicity, we only explore the performance of gradient descent on the Dhaka cholera model of \cite{king08}. Exploring the use of various first and second order methods such as Newton's method, ADAM, L-BFGS, or other popular iterative optimization methods with the gradient estimates that MOP-$\alpha$ provides in the context of the particle filter could very well lead to nontrivial speedups and increased performance.

Extensions of MOP-$\alpha$ and IFAD to panel and spatiotemporal datam, possibly through a block particle filter in the same vein as \cite{ionides22} and \cite{ning23}, would go a long way towards extending the current applicability of these methods to handle a wider variety of interesting and challenging scientific problems.

While we make the first steps towards developing a Python counterpart to the popular \texttt{pomp} R package by \citet{king16, king2017pompmanual}, this remains a work in progress.

\paragraph{Acknowledgements.} We thank Nicholas Chopin for helpful comments on the proof of a strong law of large numbers for a suitably reweighted particle filter.

\kevin{Reach: PNAS. Else try for JRSSB, try to get Chopin for Annals AE. Would need to change bracketing conventions for Biometrika. If submitting to PNAS, JASA, Biometrika, or JRSSB, highlight the different statistical gradient estimator and finite sample performance, rewrite introduction, maybe create a way to use the gradients for inference rather than using profiling. Try to rewrite it as a breakthrough in statistics. Safety: JMLR}

\bibliography{paper/bib-ifad, paper/bib-ref}
\bibliographystyle{apalike}

\appendix
\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{figure}{0}

\section{Weighted Particle Filters that Target the Posterior}

\begin{defn}[Targeting]
    A random vector $(X, w)$ drawn from a distribution $g$ \textbf{targets} the distribution $\pi$ if for any measurable and bounded functional $h$,
\begin{equation}
    E_g\{h(X) \cdot w\}=E_\pi\{h(X)\}
\end{equation}  
    A set of particles $(X^J_j, w^J_j), j=1,2, \ldots,J$, \textbf{targets} $\pi$ if for any measurable and bounded functional $h$,
\begin{equation}
    \frac{\sum_{j=1}^J h(X^J_j) w^J_j}{\sum_{j=1}^J w^J_j} \stackrel{a.s.}{\to} E_\pi(h(X))
\end{equation}
as $J \to \infty$.
\end{defn}
\citet{chopin2004clt} asserted without proof that common particle filter algorithms target the filtering distribution, $f_{X_{1:N}|Y_{1:N}}$, in this sense.
\ed{PERHAPS IF WE HAD A CONCRETE PROOF OF THIS, THAT WOULD HELP TO SEE HOW IT EXTENDS TO OUR SUBSEQUENT LEMMAS. ARGUABLY, THAT IS IN FACT WHAT WE ARE DEVELOPING VIA THESE LEMMAS. I WONDER IF SOME OTHER REFERENCE, E.G. DELMORAL'S 2004 BOOK, HAS AN EXPLICIT PROOF OF CONSISTENCY FOR SOME PARTICLE FILTER ALGORITHM.}
In order to prove the consistency of our variation on the particle filter, we now present three helper lemmas.
The first follows from standard importance sampling arguments, the second from integrating out the marginal, and the third from Bayes' theorem. 
We state Lemma~\ref{lem:change-measure-proper-weights} assuming multinomial resampling, which is convenient for the proof though other resampling strategies may be preferable in practice REFS.

\begin{lem}[Change of Particle Measure]
    \label{lem:change-measure-proper-weights}
    Suppose that $\{(\tilde X_j^J,U_j^J),j=1,\dots,J\}$ targets $f_X$. Now, let $\{(Y_j^J,V_j^J),j=1,\dots,J\}$ be a multinomial sample drawn from $\{(\tilde X_j^J,U_j^J)\}$ where $(\tilde X_j^J,U_j^J)$ is represented, on average, proportional to $\pi^J_j J$ times. Write
    \[
    (Y_j^J,V_j^J) = \big(\tilde X^J_{a(j)},U^J_{a(j)}/\pi^J_{a(j)}\big),
    \]
    where $a$ is called the ancestor function. If the importance sampling weights $U_j/\pi_j$ are bounded, then $\{(Y^J_j,V^J_j),j=1,\dots,J\}$ targets $f_X$.
\end{lem}

\begin{proof}

    Note that as the $Y_j^J$ are a subsample from $X_j^J$, $h$ can be a function of $Y$ as well as it is one for $X$. We then expand
    $$\frac{\sum_j h(Y_j^J) V_j^J}{\sum_j V_j^J} = \frac{\sum_j h(\tilde X_{k_j}^J)\frac{U_{k_j}^J}{\pi_{k_j}^J}}{\sum_j \frac{U_{k_j}^J}{\pi_{k_j}^J}}.$$
    By hypothesis,
    $$\frac{\sum_j h(\tilde X_j^J)U_j^J}{\sum_j U_j^J} \stackrel{a.s.}{\to} \E_{f_X}[f(X)].$$

    We want to show
    \begin{equation}\label{eq:lemma1:h}
    \frac{\sum_j h(X_{k_j}^J)\frac{U_{k_j}^J}{\pi_{k_j}^J}}{\sum_j \frac{U_{k_j}^J}{\pi_{k_j}^J}} - \frac{\sum_j h(\tilde X_j^J)U_j^J}{\sum_j U_j^J} \stackrel{a.s.}{\to} 0.
    \end{equation}
    For this, it is sufficient to show that
   $$ \sum_j h(\tilde X_{k_j}^J)\frac{U_{k_j}^J}{\pi_{k_j}^J}
    -  \sum_j h(\tilde X_j^J)\frac{U_j^J}{\pi_j^J}\pi_j^J \stackrel{a.s.}{\to} 0 $$
    since an application of this result with $h(x)=1$ provides almost sure convergence of the denominator in (\ref{eq:lemma1:h}).
    Write $g(\tilde X_j^J) = h(\tilde X_j^J)\frac{U_{k_j}^J}{\pi_{k_j}^J}$. We therefore need to show that 
    $$\sum_j Z_j^J := \sum_j \left(g(\tilde X_{k_j}^J) -  g(\tilde X_j^J) \pi_j^J \right) \stackrel{a.s.}{\to} 0.$$

    Because the functional $h$ and importance sampling weights $u_{k_j}^J/\pi_{k_j}^J$ are bounded, we have that $\E[\left(Z_j^J\right)^4] < \infty$. We can then follow the argument of \cite{chopin20} from this point on, where 
    {THE Zs ARE ONLY CONDITIONALLY INDEPENDENT, SO WE COULD PUT IN THE CONDITIONING EVENT (I.E., THE Xs AND Us) HERE TO BE MORE ACCURATE}
    $$\E\left[\left(\sum_j Z_j^J\right)^4\right] 
    = J\E[(Z_1^J)^4] + 3J(J-1)\left(\E[(Z_j^J)^2]\right) \leq CJ^2,$$ for some $C>0$. By Markov, 
    $$\mathbb{P}\left(\left|\frac{1}{J} \sum_{j=1}^J Z_j^J\right|>\epsilon\right) 
    \leq \frac{1}{\epsilon^4J^4 } 
    \mathbb{E}\left[\left(\sum_{j=1}^J Z_j^J\right)^4\right] \leq \frac{C}{\epsilon^4J^2},$$
    and as these terms are summable we can apply Borel-Cantelli to conclude that these deviations happen only finitely often for every $\epsilon>0,$ giving us the almost-sure convergence for

    $$ \sum_j h(X_{k_j}^J)\frac{u_{k_j}^J}{\pi_{k_j}^J}
    -  \sum_j h(X_j^J)\frac{u_j^J}{\pi_j^J}\pi_j^J \stackrel{a.s.}{\to} 0.$$

    Similarly, we also have that
    $$ \sum_j \frac{u_j^J}{\pi_j^J}\pi_j^J
    - \sum_j \frac{u_{k_j}^J}{\pi_{k_j}^J} \stackrel{a.s.}{\to} 0,$$

    and the result is proved.

    \kevin{MAY HAVE TO DEAL WITH CONDITIONING ON THE FILTRATION AT THE LAST TIMESTEP, AND THEN USING THE TOWER PROPERTY. EXPECTATION WILL BE THE SAME REGARDLESS FOR THE FOURTH MOMENT.}

   % Let $h$ be an integrable function. 
    %\begin{align*}
    %    \E\left[\sum_{j=1}^J h(Y_j) v_j\right] 
    %    &= \sum_{j=1}^J \E\left[h(X_{a(j)}) \frac{u_{a(j)}}{\pi_{a(j)}}\right] \\
    %    &= \sum_{j=1}^J \E\left[h(X_{j}) u_{j}\right],
    %\end{align*}
    %and similarly for the denominator. Now, the numerator and denominator of the reweighted quantity have the same expectation as the numerator and denominator of the original quantity, so they must converge to $ E_\pi(h(X))$ almost surely as well. 
\end{proof}

\textbf{Remark:} Note that Lemma \ref{lem:change-measure-proper-weights} permits $\pi_{1:J}$ to depend on $\{(X_j,u_j)\}$ as long as the resampling is carried out independently of $\{(X_j,u_j)\}$, conditional on $\pi_{1:J}$.

\begin{lem}[Particle Marginals]
    \label{lem:marginal-proper-weights}
    Suppose that $\{(\tilde X_j^J,U_j^J),j=1,\dots,J\}$ targets $f_X$. Also suppose that $\tilde Z_j^J \sim f_{Z|X}(\cdot | \tilde X_j^J)$ where $f_{Z|X}$ is a conditional probability density function corresponding to a joint density $f_{X,Z}$ with marginal densities $f_X$ and $f_Z$. Then, if the $U_j^J$ are bounded, $\{(\tilde Z_j^J,U_j^J)\}$ targets $f_Z$.
\end{lem}
\begin{proof}
    We want to show that, for any measurable bounded $h$, 
     $$\frac{\sum_j h(\tilde Z_j^J) \, U_j^J}{\sum_j U_j^J} \stackrel{a.s.}{\to} \E[h(Z)] = \E\big[\E[h(Z) | X]\big].$$
    By hypothesis, for any measurable and bounded functional $g$ with domain $\gX$,
    \begin{equation}\label{eq:lemma2:g}
    \frac{\sum_j g(\tilde X_j^J)\, U_j^J}{\sum_j U_j^J} \stackrel{a.s.}{\to} \E[g(X)].
    \end{equation}
    Let $\bar{U}_j^J = \frac{J U_j^J}{\sum_j U_j^J}$. Examine the numerator and denominator of the quantity $$\frac{J^{-1}\sum_j h(\tilde Z_j^J) \, U_j^J}{J^{-1}\sum_j U_j^J} = \frac{J^{-1}\sum_j h(\tilde Z_j^J) \, \bar{U}_j^J}{J^{-1}\sum_j \bar{U}_j^J}.$$
    The denominator converges to $1$ almost surely. The numerator, on the other hand, is
    $$\frac{1}{J}\sum_j h(\tilde Z_j^J)\, \bar{U}_j^J,$$
    by the same fourth moment argument to the above lemma converges almost surely to the limit of its expectation
    \begin{eqnarray*}        
    \lim_{J\to\infty} \E\left[\frac{1}{J}\sum_j h(\tilde Z_j^J)\, \bar{U}_j^J\right] 
    &=& \lim_{J\to\infty} \E \left[\frac{1}{J}\sum_j  
      \E\left[ h(\tilde Z_j^J)\, \bar{U}_j^J\big|\tilde X_j^J, \bar U_j^J\right]
    \right]
    \\
    &=& \lim_{J\to\infty} \E \left[\frac{1}{J}\sum_j  \E\big[h(Z) \big| X=\tilde X_j^J \big] \, \bar U_j^J\right].
    \end{eqnarray*}
    Applying (\ref{eq:lemma2:g}) with $g(x) = \E\big[h(Z) | X=x\big]$, the average on the right hand size converges almost surely to $\E\big\{\E[h(Z)|X]\big\}=\E[h(Z)].$
    It remains to swap the limit and expectations. We can do so with the bounded convergence theorem, and therefore obtain   
    $$\frac{1}{J}\sum_j h(\tilde Z_j^J)\,\bar{U}_j^J \stackrel{a.s.}{\to} \E[h(Z)].$$

\end{proof}

\begin{lem}[Particle Posteriors]
    \label{lem:posterior-proper-weights}
    Suppose that $\{(X_j^J,u_j^J),j=1,\dots,J\}$ targets $f_X$. Also suppose that $(X^{\prime J}_j,u^{\prime J}_j) = \big(X_j^J,u_j^J \cdot f_{Z|X}(z^*|X_j^J)\big)$. Then, if $u_j^J \cdot f_{Z|X}(z^*|X_j^J)\big)$ and $u_j^J \cdot f_{Z|X}(z^*|X_j^J)\big) / f_Z(z^*)$ are bounded, $\{(X^{\prime J}_j,u^{\prime J}_j)\}$ targets $f_{X|Z}(\cdot | z^*)$.
\end{lem}

\begin{proof}

    $$ \frac{J^{-1} \sum_j h(X_j^J) \bar{u}_j^{' J}}{J^{-1} \sum_j \bar{u}_j^{' J}}$$

    by

    \begin{equation}
    \frac{J^{-1} \sum_j h(X_j^J) f_{Z|X}(z^*|X_j^J) {u}_j^{J}}{J^{-1} \sum_j {u}_j^{J}}
    \times \left( \frac{J^{-1} \sum_j f_{Z|X}(z^*|X_j^J) {u}_j^{J}}{J^{-1} \sum_j {u}_j^{J}} \right)^{-1}
    \end{equation}

    Again, we want to show that
     $$\frac{\sum_j h(X_j^J) \cdot u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{\sum_j u_j^J \cdot f_{Z|X}(z^*|X_j^J)} \stackrel{a.s.}{\to} \E_{f_{X|Z}}[h(X)|z^*].$$

    We already have that
    $$\frac{\sum_j h(X_j^J)u_j^J}{\sum_j u_j^J} \stackrel{a.s.}{\to} \E_{f_X}[h(X)].$$

    %= \frac{u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{\sum_j f_{Z|X}(z^*|X_j^J)f_X(X_j^J)}

    Consider the following:
    \begin{equation}
    \frac{J^{-1} \sum_j h(X_j^J) f_{Z|X}(z^*|X_j^J) {u}_j^{J}}{J^{-1} \sum_j {u}_j^{J}}
    \times \left( \frac{J^{-1} \sum_j f_{Z|X}(z^*|X_j^J) {u}_j^{J}}{J^{-1} \sum_j {u}_j^{J}} \right)^{-1}.
    \end{equation}

    Write $\bar{u}_j^{J} = \frac{J u_j^J}{\sum_j u_j^J}$. Define the weights $u^{' J}_j := u_j^J \cdot f_{Z|X}(z^*|X_j^J),$ and the corresponding self-normalized weights $\bar{u}_j^{' J} := \frac{J u^{' J}_j}{\sum_j u^{' J}_j}.$ The denominator of the below quantity 
    $$ \frac{J^{-1} \sum_j h(X_j^J) \bar{u}_j^{' J}}{J^{-1} \sum_j \bar{u}_j^{' J}}$$
    converges to $1$, while the numerator, by the same fourth moment argument as the first lemma, converges almost surely to the limit of its expectation, which is
    \begin{align*}
        \lim_{J\to\infty} \E\left[J^{-1} \sum_j h(X_j^J) \bar{u}_j^{' J} \right]
        &= \lim_{J\to\infty} \E\left[J^{-1} \sum_j h(X_j^J) \frac{J u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{\sum_j u_j^J \cdot f_{Z|X}(z^*|X_j^J)}\right] \\
        &= \lim_{J\to\infty} \E\left[J^{-1} \sum_j h(X_j^J) \frac{J \frac{u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{f_Z(z^*)}}{\sum_j \frac{u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{f_Z(z^*)}}\right]\\
        &= \lim_{J\to\infty} \E\left[ \sum_j h(X_j^J) \frac{\frac{u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{f_Z(z^*)}}{\sum_j \frac{u_j^J \cdot f_{Z|X}(z^*|X_j^J)}{f_Z(z^*)}}\right]\\
        &= \E_{X|Z}[h(X)|z^*],
    \end{align*}

    where we can swap the limit and expectations with the bounded convergence theorem.
\end{proof}

\begin{prop}[MOP-1 Targets the Posterior]
    When $\alpha=1$ or $\phi=\theta$, MOP-$\alpha$ targets the posterior. 
\end{prop}
\begin{proof}
    When $\theta=\phi$, regardless of the value of $\alpha$, the ratio $\frac{g_{n,j}^\theta}{g_{n,j}^\phi}=1,$ and this reduces to the vanilla particle filter estimate.

    When $\alpha=1$, and $\theta\neq\phi,$ the proof is as follows. Recursively applying Lemmas \ref{lem:change-measure-proper-weights}, \ref{lem:marginal-proper-weights}, and \ref{lem:posterior-proper-weights}, we obtain that 
    %to step~\ref{mop:step1}, Lemma~2 step~ {mop:weight:update} and Lemma~3 to step~\ref{mop:step2} we obtain that
    the MOP-1 filter targets the posterior.
    Specifically, suppose inductively that $\big\{\big(X^{F,\theta}_{n-1,j},w^{F,\theta}_{n-1,j}\big)\big\}$ is properly weighted for $f_{X_{n-1}|Y_{1:n-1}}(x_{n-1}|y^*_{1:n-1};\theta)$.
    Then, Lemma \ref{lem:marginal-proper-weights} tells us that $\big\{\big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j}\big)\big\}$ targets $f_{X_{n}|Y_{1:n-1}}(x_{n}|y^*_{1:n-1};\theta)$.
    Lemma \ref{lem:posterior-proper-weights} tells us that $\big\{\big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j} g^\theta_{n,j} \big)\big\}$ therefore targets  $f_{X_{n}|Y_{1:n}}(x_{n}|y^*_{1:n};\theta)$.
    Lemma \ref{lem:change-measure-proper-weights} guarantees that the resampling rule, given by 
    \[
    \big(X^{F,\theta}_{n,j},w^{F,\theta}_{n,j}\big) = \big(X^{P,\theta}_{n,a(j)}, w^{P,\theta}_{n,a(j)} g^\theta_{n,a(j)}\big/ g^\phi_{n,a(j)}\big),
    \]
    with resampling weights proportional to $g^\phi_{n,j}$, therefore also targets $f_{X_{n}|Y_{1:n}}(x_{n}|y^*_{1:n};\theta)$.
\end{proof}

This has addressed filtering, but not quite yet the likelihood evaluation. For this we use the following lemma.

\begin{lem}[Likelihood Proper Weighting]
    \label{lem:lik-proper-weight}
  $f_{Y_n|Y_{1:n-1}}(y_n^*|y_{1_n-1}^*;\theta)$ is consistently estimated by either the before-resampling estimate,
\begin{equation}\label{L1}
L_n^{B,\theta} =  \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}},
\end{equation}
or by the after-resampling estimate,
\begin{equation}\label{L2}
L_n^{A,\theta} = L^\phi_n \frac{\sum_{j=1}^Jw^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}.
\end{equation}
where $L^\phi_n$ is as defined in the various algorithms.
\end{lem}

Here, (\ref{L1}) is a direct consequence of our earlier result that $\{ \big(X^{P,\theta}_{n,j},w^{P,\theta}_{n,j}\big) \}$ targets $f_{X_{n}|Y_{1:n-1}}(x_{n}|y^*_{1:n-1};\theta)$.
To see  (\ref{L2}),
we write the numerator of (\ref{lem:change-measure-proper-weights}) as
\[
L^\phi_n \sum_{j=1}^J \left[ \frac{g^\theta_{n,j}}{g^\phi_{n,j}} w^{P,\theta}_{n,j}\right] \frac{g^\phi_{n,j}}{L_n^\phi}
= L^\phi_n \sum_{j=1}^J w_{n,j}^{FC,\theta} \frac{g^\phi_{n,j}}{L_n^\phi}
\]
Using Lemma \ref{lem:change-measure-proper-weights}, we resample according to probabilities $\frac{g^\phi_{n,j}}{L_n^\phi}$ to see this is properly estimated by
\[
L^\phi_n \sum_{j=1}^J w^{F,\theta}_{n,j},
\]
from which we obtain (\ref{L2}).

Using Lemma \ref{lem:lik-proper-weight}, we obtain a likelihood estimate,
\[
L^{A,\theta} = \prod_{n=1}^N \left( L^\phi_n \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J w^{P,\theta}_{n,j}}\right).
\]
Since $w^{F,\theta}_{n,j}=w^{P,\theta}_{n+1,j}$, this is a telescoping product. The remaining terms are
$\sum_{j=1}^J w^{P,\theta}_{0,j} = J$ on the denominator and $\sum_{j=1}^J w^{F,\theta}_{N,j}$ on the numerator.
This derives the MOP/DOP estimate.

\kevin{IS THIS UNBIASED?}

$L^{B,\theta}$ should generally be preferred in practice, since there is no reason to include the extra variability from resampling when calculating the conditional log likelihood, but it lacks the nice telescoping product that lets us derive exact expressions for the gradient later.

\section{Doubly Off-Policy Particle Filters}
\label{app:dop}

The proof for Proposition \ref{prop:dop-correctness} follows from the below two lemmas. Evaluating the off-policy likelihood is possible with a weighted particle filter by using resampling weight $p(y^*_n|x_n;\phi)$ and making two corrections in the particle weight, through multiplying $w_{t,j}$ by the two likelihood ratios below,

\begin{equation}
    \label{eq:dmeas-ratio}
    s=p(y^*_n|x_n;\theta)/p(y^*_n|x_n;\phi)
\end{equation}
and
\begin{equation}
    \label{eq:rproc-ratio}
    r=p(x_n|x_{n-1};\theta)/p(x_n|x_{n-1};\phi).
\end{equation}

The following result showcases the correctness of corrections \ref{eq:dmeas-ratio} and \ref{eq:rproc-ratio}.

\begin{lem}
    For $\theta$, $\phi$ that satisfy Proposition \ref{prop:smooth-nbhd}, the following correction in the particle filter weights,
    \begin{equation}
        w_{t,j} := \tilde{w}_{t,j} \cdot p(y^*|x_{t,j}^P,\phi)\cdot r \cdot s
    \end{equation}
    successfully recovers $\lik(\theta)$.
\end{lem}

\begin{proof}
    Recall $\lik(\theta) = \E[\lik(\theta,\phi,\Omega, J)]=\int_\Omega \lik(\theta,\phi,\omega, J)\, d\mu(\omega)$ and $\lik(\theta) = \int_\Omega \lik(\theta, \omega, J) d\mu(\omega)$. We want to evaluate
    \begin{align*}
        \lik(\theta,\omega, J)
        &= \E\left[ \prod_{t=1}^T \lik_t(\theta) \middle| \omega\right]
        = \E\left[ \prod_{t=1}^T p(y_t^*|y_{1:t-1}^*,\theta) \middle| \omega\right] \\
        &= \E\left[ \prod_{t=1}^T \frac{1}{J} \sum_{j=1}^J  p(y_t^*|x_{t,j}^P,\theta) p(x_{t,j}^P|y_{1:t-1}^*,\theta) \middle| \omega\right] \\
        &= \E\left[ \prod_{t=1}^T\frac{1}{J}  \sum_{j=1}^J  p(y_t^*|x_{t,j}^P,\theta) p(x_{t,j}^P|x_{t-1}^F,\theta) \middle| \omega\right] 
    \end{align*}
    where $p(x_t^P|y_{1:t-1}^*,\theta)$ is the prediction distribution of particles, and the filtering distribution of particles is proportional to $p(y_t^*|x_{t,j}^P,\theta)$, the likelihood of particle $j$ at time $t$.

    Consider first the case with no resampling. Let $\lik^j(\theta)$, $\lik^j(\phi)$ be the likelihood of trajectory $j$ under $\theta$ and $\phi$ respectively. As the prediction and filtering particles are then the same, we use $x_{t,j}$ to refer to both. As such, conditional on $\omega$,
    \begin{align*}
        \lik^j(\theta) 
        &= \prod_{t=1}^T p(y_t^*|x_{t,j}, \theta) p(x_{t,j}|x_{t-1,j}, \theta) \\
        &= \prod_{t=1}^T p(y_t^*|x_{t,j}, \phi) \frac{p(y_t^*|x_{t,j}, \theta)}{p(y_t^*|x_{t,j}, \phi)} p(x_{t,j}|x_{t-1,j}, \theta) \frac{p(x_{t,j}|x_{t-1,j}, \theta) }{p(x_{t,j}|x_{t-1,j}, \phi) } \\
        &= \prod_{t=1}^T p(y_t^*|x_{t,j}, \phi) \cdot s \cdot p(x_{t,j}|x_{t-1,j}, \phi)\cdot r
    \end{align*}

    and therefore, if trajectories were proposed with equal probability with particles conditional on $\omega$,
    $$ \lik(\theta, \omega, J) = \frac{1}{J} \sum_{j=1}^J \prod_{t=1}^T p(y_t^*|x_{t,j}, \phi) \cdot s \cdot p(x_{t,j}|x_{t-1,j}, \phi)\cdot r $$
    The correction is therefore sufficient for recovering $\lik(\theta)$, with trajectories collected with the process model at $\phi$ and measurements evaluated with the measurement model at $\phi$.

    Consider now the case with resampling. Without loss of generality we can resample every timestep. Here, $x_{t,j}^F(\phi) \sim p(y^*|x_{t,j}^P, \phi)$ and $x_{t,j}^P(\phi) \sim p(x_t|x_{t-1}^F,\phi)$, and similarly for $\theta$. We use this notation to emphasize that \textbf{both} the measurement model and the previously drawn particles depend on $\phi$ or $\theta$.

    \begin{align*}
        \lik_t(\theta) &= \E_{x_{t,j}^P \sim p(x_t|x_{t-1}^F, \theta)} \left[\frac{1}{J}\sum_{j=1}^J p(y_t^*|x_{t,j}^P,\theta) \right] \\
        &= \E_{x_{t,j}^P \sim p(x_t|x_{t-1}^F, \theta)} \left[\frac{1}{J}\sum_{j=1}^J p(y_t^*|x_{t,j}^P,\phi) \frac{p(y_t^*|x_{t,j}^P,\theta)}{p(y_t^*|x_{t,j}^P,\phi)} \right] \\
        &= \E_{x_{t,j}^P \sim p(x_t|x_{t-1}^F, \phi)} \left[\frac{1}{J}\sum_{j=1}^J p(y_t^*|x_{t,j}^P,\phi) \frac{p(y_t^*|x_{t,j}^P,\theta)}{p(y_t^*|x_{t,j}^P,\phi)} \frac{p(x_{t,j}|x_{t-1,j}, \theta) }{p(x_{t,j}|x_{t-1,j}, \phi) }\right] \\
        &= \E_{x_{t,j}^P \sim p(x_t|x_{t-1}^F, \phi)} \left[\frac{1}{J}\sum_{j=1}^J p(y_t^*|x_{t,j}^P,\phi) \cdot s \cdot r \right] \\
        &= \E_{\omega}\E_{x_{t,j}^P \sim \process(x_{t-1}^F, \omega, \phi)} \left[\frac{1}{J}\sum_{j=1}^J p(y_t^*|x_{t,j}^P,\phi) \cdot s \cdot r \middle| \omega \right]
    \end{align*}

    and again the correction is sufficient.

    Therefore,

    \begin{align}
         \mathcal{L}(\theta) 
         &= \int_\Omega \left(
         \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f(y_n^* | x_{n,j}^P(\omega), \theta) p(x_{n,j}^P(\omega) | x_{n-1,j}^P(\omega), \theta) \right) \, d\mu(\omega) \\
         &= \int_\Omega \left(
         \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f(y_n^* | x_{n,j}^{P,\theta}(\omega), \theta) p(x_{n,j}^{P,\theta}(\omega) | x_{n,j}^P(\omega), \theta) \right) \, d\mu(\omega) \\
         &= \int_\Omega \left(
         \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f(y_n^* | x_{n,j}^{P,\theta}(\omega), \phi) p(x_{n,j}^{P,\theta}(\omega) | y_n^*, \phi) \cdot s  \right) \, d\mu(\omega) \\
         &= \int_\Omega \left(
         \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f(y_n^* | x_{n,j}^{P,\phi}(\omega), \phi) p(x_{n,j}^{P,\phi}(\omega) | y_n^*, \phi) \cdot r \cdot s \right) \, d\mu(\omega) \\
         &= \int_\Omega \left(
         \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f(y_n^* | x_{n,j}^{P,\phi}(\omega), \phi) p(x_{n,j}^{P, \phi}(\omega) | y_n^*, \phi) \cdot r \cdot s \right) \, d\mu(\omega) \\
         &= \int_\Omega \left(
         \prod_{n=1}^N \frac{1}{J} \sum_{j=1}^J f(y_n^* | x_{n,j}^{P,\phi}(\omega), \phi) \cdot r \cdot s \right)  p(x_{1:N} | y_{1:N}^*, \phi) \,  d\mu(\omega) \\
         &=: \int_\Omega \lik(\theta,\phi,\omega, J)\, d\mu(\omega).
     \end{align}

     The proof for an arbitrary functional under the particle measure follows analogously.
\end{proof}

\begin{lem}[Gradient of Doubly Off-Policy Particle Filter]
    Let the system evolve according to $\phi \in \Theta,$ that is, with the process model $\process(\cdot | \phi)$ and resampling with probabilities proportional to $f_{Y_n|X_n}(y_n^* | x_{n,j}^{P,\phi})$. Also let $\theta \in \gN(\phi)$. When $\theta$ is evaluated at $\phi$, its gradient is the same as the estimator of \citet{poyiadjis11},
    \begin{equation}
        \nabla_\theta \hat\ell(\theta) = \frac{1}{J}\sum_{j=1}^J \nabla_\theta \log f_{X_{0:N}, Y_{1:N}}(x_{1:N,j}^{A, F,\theta}, y_{1:N}^*).
    \end{equation}
\end{lem}

\begin{proof}
    It is known, from \cite{doucet2009tutorial}, that 
    $$p_\theta(x_{1:N}|y_{1:N}) = \prod_{n=1}^N \frac{p(x_n|x_{n-1})p(y_n|x_n)}{p(y_n|y_{1:n-1})} = \prod_{n=1}^N \frac{p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)}{\lik_n(\theta)}.$$

    From the above lemma, and the fact that the particle filter is unbiased for the likelihood, we know that
    \begin{align*}
        \lik(\theta)
        &= \int_{x_{1:N}} \prod_{n=1}^N p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n) p_\theta(x_{1:N}|y_{1:N}) dx_{1:N} \\
        &= \int_{x_{1:N}} \left(\prod_{n=1}^N  p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n) \right) \frac{p_\theta(x_{1:N}|y_{1:N})}{p_\phi(x_{1:N}|y_{1:N})}p_\phi(x_{1:N}|y_{1:N}) dx_{1:N}\\
        &= \int_{x_{1:N}} \left(\prod_{n=1}^N \lik_n(\theta) \right) \left(\prod_{n=1}^N \frac{p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n) \lik_n(\phi)}{p_\phi(x_n|x_{n-1})p_\phi(y_n|x_n) \lik_n(\theta)}\right)p_\phi(x_{1:N}|y_{1:N}) dx_{1:N} \\ 
        &= \int_{x_{1:N}} \left(\prod_{n=1}^N \lik_n(\phi) \frac{p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)}{p_\phi(x_n|x_{n-1})p_\phi(y_n|x_n) }\right)p_\phi(x_{1:N}|y_{1:N}) dx_{1:N}.
    \end{align*}

    Therefore, under regularity conditions that permit differentiating under the integral sign, using the log-derivative trick, we have
    \begin{align*}
        \nabla_\theta \lik(\theta)
        &= \int_{x_{1:N}} \nabla_\theta \hat\lik(\theta) \;p_\phi(x_{1:N}|y_{1:N})dx_{1:N}\\
        &= \int_{x_{1:N}} \nabla_\theta \left(\prod_{n=1}^N \lik_n(\phi) \frac{p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)}{p_\phi(x_n|x_{n-1})p_\phi(y_n|x_n) }\right)p_\phi(x_{1:N}|y_{1:N}) dx_{1:N} \\
        &= \int_{x_{1:N}} \nabla_\theta \left(\prod_{n=1}^N \left(\frac{p_\phi(x_n|x_{n-1})p_\phi(y_n|x_n)}{\lik_n(\phi)}\right)^{-1} p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)p_\phi(x_{1:N}|y_{1:N}) dx_{1:N} \\
        &= \int_{x_{1:N}} \nabla_\theta \left(\frac{1}{p_\phi(x_{1:N}|y_{1:N}) }\prod_{n=1}^N p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)p_\phi(x_{1:N}|y_{1:N}) dx_{1:N} \\
        &= \int_{x_{1:N}} \nabla_\theta \left(\prod_{n=1}^N p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right) dx_{1:N} \\
        &= \int_{x_{1:N}} \left(\prod_{n=1}^N p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)\nabla_\theta \log\left(\prod_{n=1}^N p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right) dx_{1:N} \\
        &= \int_{x_{1:N}} p_\theta(x_{1:N},y_{1:N}) \left(\sum_{n=1}^N \nabla_\theta \log\left(p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)\right) dx_{1:N} \\
        &= \int_{x_{1:N}} \left(\sum_{n=1}^N \nabla_\theta \log\left(p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)\right) p_\theta(x_{1:N}|y_{1:N}) \cdot \lik(\theta) \; dx_{1:N}. 
    \end{align*}

    Using the log-derivative trick again,
    \begin{align*}
        \nabla_\theta \ell(\theta)
        &= \frac{\nabla_\theta \lik(\theta)}{\lik(\theta)} \\
        &= \int_{x_{1:N}} \left(\sum_{n=1}^N \nabla_\theta \log\left(p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)\right) p_\theta(x_{1:N}|y_{1:N}) dx_{1:N} \\
        &= \E_{x_{1:N}|y_{1:N}, \theta}\left[\nabla_\theta \left(\sum_{n=1}^N\log\left(p_\theta(x_n|x_{n-1})p_\theta(y_n|x_n)\right)\right)\right] \\
        &= \E_{x_{1:N}|y_{1:N}, \theta}\left[\nabla_\theta\log p_\theta(x_{1:N},y_{1:N})\right] \\
        &= \lim_{J\to\infty} \E_{x_{1:N}|y_{1:N}, \theta}\left[\frac{1}{J}\sum_{j=1}^J\nabla_\theta \log p_\theta(x_{1:N,j}^{A,F},y_{1:N}^*)\right].
    \end{align*}

    But it also holds that
    \begin{align*}
        \nabla_\theta \ell(\theta)
        &= \frac{\nabla_\theta \lik(\theta)}{\lik(\theta)} \\
        &= \int_{x_{1:N}} \frac{\nabla_\theta \hat\lik(\theta)}{\lik(\theta)} \;p_\phi(x_{1:N}|y_{1:N})dx_{1:N} \\
        &= \int_{x_{1:N}} \nabla_\theta \hat\ell(\theta) \;p_\phi(x_{1:N}|y_{1:N})dx_{1:N}\\
        &= \lim_{J\to\infty} \E_{x_{1:N}|y_{1:N}, \phi}\left[\nabla_\theta \hat\ell(\theta)\right].
    \end{align*}

    and we see that $$\frac{1}{J}\sum_{j=1}^J\nabla_\theta \log p_\theta(x_{1:N,j}^{A,F},y_{1:N}^*)$$ and $$\nabla_\theta \hat\ell(\theta)$$ both converge to the same thing in expectation. They are both particle approximations under the posterior where the variance of each is bounded, therefore one can apply the particle CLT from \cite{chopin2004clt} to obtain that they are both consistent estimates of the score.

    In fact, when $\theta=\phi,$ they are the same in distribution. \kevin{claim}

    \kevin{This is a far weaker result than what I really wanted. This just gives consistency. Want equality in distribution.}
\end{proof}

\begin{lem}

Consider the case of DOP-$\alpha$ when $\alpha=0$ and $\theta=\phi$. The gradient estimate is then

    \begin{equation}
        \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n,X_n|X_{n-1}}(y_n^*, x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \theta\right),
    \end{equation}

    yielding the gradient estimator of \cite{blei2018vsmc} that \cite{scibior2021dpf} show is the gradient of a vanilla particle filter when the resampling terms are dropped.

\end{lem}

\begin{proof}
    When $\alpha=0,$ the likelihood estimate becomes
\begin{equation}
    \hat{\lik}(\theta) := \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n,j}^{P,\theta}} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J r_{n,j}s_{n,j}}{\sum_{j=1}^J r_{n,j}} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J r_{n,j}s_{n,j}.
\end{equation}

Its gradient when $\theta=\phi$ is therefore 
\begin{align*}
    \nabla_\theta \hat{\ell}(\theta) &:= \sum_{n=1}^N \nabla_\theta \log\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J r_{n,j}s_{n,j}\right) \\
    &= \sum_{n=1}^N \frac{\nabla_\theta \left(L_n^\phi \frac{1}{J} \sum_{j=1}^J r_{n,j}s_{n,j}\right)}{\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J r_{n,j}s_{n,j}\right)} \\
    &= \sum_{n=1}^N \frac{\sum_{j=1}^J \nabla_\theta r_{n,j}s_{n,j}}{\sum_{j=1}^J r_{n,j}s_{n,j}} \\
    &= \sum_{n=1}^N \frac{1}{J} \sum_{j=1}^J \frac{\nabla_\theta f_{Y_n,X_n|X_{n-1}}(y_n^*, x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \theta)}{f_{Y_n, X_n|X_{n-1}}(y_n^*, x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \phi)} \\
    &= \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n,X_n|X_{n-1}}(y_n^*, x_{n,j}^{P, \phi}|x_{n-1,j}^{F, \phi}; \theta\right),
\end{align*}
where we use the log-derivative trick, observe that $\sum_{j=1}^J r_{n,j}s_{n,j} = J$ when $\theta=\phi$, and use the log-derivative trick where $\theta=\phi$ again. 
\end{proof}

\subsection{Off-Policy Weighted Differentiable Particle Filter}

We are finally ready to introduce the off-policy weighted differentiable particle filter. In the event where a seed is fixed, all expressions in Algorithm \ref{alg:dop} can be considered to be conditional on $\omega \in \Omega$.

\begin{algorithm}[ht]
\centering
	\caption{Doubly Off-Policy-$\alpha$}
    \label{alg:dop}
	\begin{algorithmic}[1]
	     \STATE \textbf{Input:} Number of particles $J$, timesteps $N$, measurement model $f_{Y_n|X_n}(y_n^*|x_n, \theta)$, simulator $\process(x_{n+1}|x_n, \theta)$, evaluation parameter $\theta$, behavior parameter $\phi$, seed $\omega$.
		\STATE Initialize filter particles ${X}_{0,j}^{F,\phi}\sim {f}_{{X}_{0}}\left(\cdot\giventh{\phi}\right)$, relative weights $w^{F,\theta}_{0,j}= 1$ for $j$ in $\seq{1}{J}$. Fix $\omega.$
		\FOR{$n=1,...,N$}
            \STATE Prediction weights with discounting: $w_{n,j}^{P,\theta} = \big(w_{n-1,j}^{F,\theta}\big)^\alpha$ for $j\ \mathrm{in}\ \seq{1}{J}$
            \label{dop-alpha:discount}
            \STATE Simulate for prediction:
            ${X}_{n,j}^{P,\phi}\sim {f}_{{X}_{n}|{X}_{n-1}}\big(\cdot|{X}_{n-1,j}^{F};{\phi}\big)$ for $j\ \mathrm{in}\ \seq{1}{J}$ \label{dop-alpha:step1}
            \STATE Adjust weights: $\displaystyle w_{n,j}^{P,\theta} = w_{n,j}^{P,\theta} \times
            \frac{{f}_{{X}_{n}|{X}_{n-1}}\big({X}_{n,j}^{P,\phi}|{X}_{n-1,j}^{F};{\theta}\big)}{{f}_{{X}_{n}|{X}_{n-1}}\big({X}_{n,j}^{P,\phi}|{X}_{n-1,j}^{F};{\phi}\big)}$ for $j$ in $1:J$
            \label{dop-alpha:dproc}
            \STATE Evaluate measurement density:
            $g^{\theta}_{n,j}={f}_{{Y}_{n}|{X}_{n}}(y_{n}^{*}|{X}_{n,j}^{P,\phi}\giventh{\theta})$ for $j$ in $\seq{1}{J}$
            \STATE Before-resampling conditional likelihood: $\displaystyle L_n^{B,\theta,\alpha} = \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
            \STATE Conditional likelihood under $\phi$: 
            $L_n^{\phi} = \frac{1}{J}\sum_{m=1}^{J}g^{\phi}_{n,m}$
            \label{dop-alpha:Lphi}
            \STATE Normalize weights:
            $\displaystyle \tilde{g}^{\phi}_{n,j}= \frac{g^{\phi}_{n,j}}{JL_n^{\phi}}$
            for $j\ \mathrm{in}\ \seq{1}{J}$
            \STATE Apply systematic resampling to select indices $k_{1:J}$ with $\prob\big(k_{j}=m\big) =\tilde{g}^{\phi}_{n,m}$ \label{dop-alpha:systematic}
            \STATE Resample particles: ${X}_{n,j}^{F,\phi}={X}_{n,k_{j}}^{P,\phi}$
            \STATE Filter weights corrected for resampling:
            $\displaystyle w^{FC,\theta}_{n,j}= w^{P,\theta}_{n,j} \times \frac{ g^{\theta}_{n,j}}{ g^{\phi}_{n,j}}$ for $j\ \mathrm{in}\ \seq{1}{J}$ \label{dop-alpha:weight:update}
            \STATE Resample filter weights:
            $w_{n,j}^{F,\theta}= {w}_{n,k_{j}}^{FC,\theta}$
            for $j$ in $\seq{1}{J}$ \label{dop-alpha:step2}
            \STATE After-resampling conditional likelihood: $\displaystyle L_n^{A,\theta,\alpha} = L_n^\phi \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
            \ENDFOR
		\RETURN likelihood estimate $\hat{\lik}(\theta) = \lik(\theta, \phi, \omega, J) := \prod_{n=1}^N L_n^{A,\theta,\alpha}$, filtering distribution $\{(X_{N,j}^{F, \theta}, w^{F,\theta}_{N,j})\}.$
	\end{algorithmic}
\end{algorithm}

Here, we consider a doubly off policy particle filter, meaning that both \code{rprocess} and \code{dmeasure} are computed at $\phi$ and particles are reweighted to correspond to $\theta$.

\section{Measurement Off-Policy Particle Filters}

\begin{algorithm}[ht]
\centering
	\caption{Measurement Off-Policy-$\alpha$}
    \label{alg:mop}
	\begin{algorithmic}[1]
	     \STATE \textbf{Input:} Number of particles $J$, timesteps $N$, measurement model $f_{Y_n|X_n}(y_n^*|x_n, \theta)$, simulator $\process(x_{n+1}|x_n, \theta)$, evaluation parameter $\theta$, behavior parameter $\phi$, seed $\omega$.
        \IF{$\theta \neq \phi$}
            \STATE Run a particle filter once with parameters $\phi$ to obtain $X_{n,j}^{P,\phi}, X_{n,j}^{F,\phi}$, with seed $\omega$.
        \ELSE
        \STATE Set $X_{n,j}^{P,\phi}, X_{n,j}^{F,\phi}$ to be copies of $X_{n,j}^{P,\theta}, X_{n,j}^{P,\theta}$ in the rest of the algorithm.
        \ENDIF
		\STATE Initialize filter particles ${X}_{0,j}^{F,\theta}\sim {f}_{{X}_{0}}\left(\cdot\giventh{\theta}\right)$, relative weights $w^{F,\theta}_{0,j}= 1$ for $j$ in $\seq{1}{J}$. Fix $\omega.$
		\FOR{$n=1,...,N$}
            \STATE Prediction weights with discounting: $w_{n,j}^{P,\theta} = \big(w_{n-1,j}^{F,\theta}\big)^\alpha$ for $j\ \mathrm{in}\ \seq{1}{J}$
            \label{mop-alpha:discount}
            \STATE Simulate for prediction:
            ${X}_{n,j}^{P,\theta}\sim {f}_{{X}_{n}|{X}_{n-1}}\big(\cdot|{X}_{n-1,j}^{F};{\theta}\big)$ for $j\ \mathrm{in}\ \seq{1}{J}$ \label{mop-alpha:step1}
            \STATE Evaluate measurement density:
            $g^{\theta}_{n,j}={f}_{{Y}_{n}|{X}_{n}}(y_{n}^{*}|{X}_{n,j}^{P,\theta}\giventh{\theta})$ for $j$ in $\seq{1}{J}$
            \STATE Before-resampling conditional likelihood: $\displaystyle L_n^{B,\theta,\alpha} = \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
            \STATE Conditional likelihood under $\phi$: 
            $L_n^{\phi} = \frac{1}{J}\sum_{m=1}^{J}g^{\phi}_{n,m}$
            \label{mop-alpha:Lphi}
            \STATE Normalize weights:
            $\displaystyle \tilde{g}^{\phi}_{n,j}= \frac{g^{\phi}_{n,j}}{JL_n^{\phi}}$
            for $j\ \mathrm{in}\ \seq{1}{J}$
            \STATE Apply systematic resampling to select indices $k_{1:J}$ with $\prob\big(k_{j}=m\big) =\tilde{g}^{\phi}_{n,m}$ \label{mop-alpha:systematic}
            \STATE Resample particles: ${X}_{n,j}^{F,\phi}={X}_{n,k_{j}}^{P,\phi}$
            \STATE Filter weights corrected for resampling:
            $\displaystyle w^{FC,\theta}_{n,j}= w^{P,\theta}_{n,j} \times \frac{ g^{\theta}_{n,j}}{ g^{\phi}_{n,j}}$ for $j\ \mathrm{in}\ \seq{1}{J}$ \label{mop-alpha:weight:update}
            \STATE Resample filter weights:
            $w_{n,j}^{F,\theta}= {w}_{n,k_{j}}^{FC,\theta}$
            for $j$ in $\seq{1}{J}$ \label{mop-alpha:step2}
            \STATE After-resampling conditional likelihood: $\displaystyle L_n^{A,\theta,\alpha} = L_n^\phi \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
            \ENDFOR
		\RETURN likelihood estimate $\hat{\lik}(\theta) = \lik(\theta, \phi, \omega, J) := \prod_{n=1}^N L_n^{A,\theta,\alpha}$, filtering distribution $\{(X_{N,j}^{F, \theta}, w^{F,\theta}_{N,j})\}.$
	\end{algorithmic}
\end{algorithm}

\begin{lem}
    \label{lem:mop-1-formula}
    Consider the case of MOP-$\alpha$ when $\alpha=1$ and $\theta=\phi$. The gradient estimate is then

    \begin{equation}
        \nabla_\theta \hat{\ell}(\theta) := \frac{1}{J}\sum_{j=1}^J \nabla_\theta \log f_{Y_{1:N}|X_{1:N}}\left(y_{1:N}^* | x_{1:n,j}^{A, F,\theta}\right),
    \end{equation}

    yielding the gradient estimators of \cite{poyiadjis11, scibior2021dpf} when applied to the bootstrap filter. 
\end{lem}

\begin{proof}

Consider the case of MOP-$\alpha$ when $\alpha=1$ and $\theta=\phi$. The likelihood estimate becomes
\begin{equation*}
    \hat{\lik}(\theta) := \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n,j}^{P,\theta}}= \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n-1,j}^{F,\theta}} = \left(\frac{1}{J}\sum_{j=1}^J w_{N,j}^{F,\theta}\right) \prod_{n=1}^N L_n^\phi,
\end{equation*}
as we have a telescoping product. Then, as

$$\nabla_\theta \hat\ell(\theta) = \frac{\nabla_\theta \hat\lik(\theta)}{\hat\lik(\theta)} = \frac{\nabla_\theta\left(\frac{1}{J}\sum_{j=1}^J w_{N,j}^{F,\theta}\right) \prod_{n=1}^N L_n^\phi}{\prod_{n=1}^N L_n^\phi} =  \frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,j}^{F,\theta},$$

The derivative of the log-likelihood estimate is then
\begin{equation*}
    \nabla_\theta \hat{\ell}(\theta) := \frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,j}^{F,\theta},
\end{equation*}
which we decompose as follows.

Observe that as $\alpha=1$,

$$w_{n,j}^{P,\theta} = w_{n-1,j}^{F,\theta}\frac{g_{n,j}^\theta}{g_{n,j}^\phi} = \prod_{i=1}^n \frac{g_{i,j}^{A,P,\theta}}{g_{i,j}^{A,P,\phi}},$$
where we use the $(\cdot)^A$ superscript to denote the ancestral trajectory of the $j$-th prediction or filtering particle at timestep $n$.

Using the log-derivative identity again,
$$\frac{\nabla_\theta w_{n,j}^{P,\theta}}{w_{n,j}^{P,\theta}} = \nabla_\theta \log w_{n,j}^{P,\theta} = \nabla_\theta \log \left(\prod_{i=1}^n \frac{g_{i,j}^{A,P,\theta}}{g_{i,j}^{A,P,\phi}}\right) = \nabla_\theta \sum_{i=1}^n \left(\log g_{i,j}^{A,P,\theta} - \log g_{i,j}^{A,P,\phi}\right) = \sum_{i=1}^n \nabla_\theta \log g_{i,j}^{A,P,\theta}.$$

So 
$$  \nabla_\theta \sum_{n=1}^N \log g_{n,j}^{A,\theta} = \nabla_\theta \log\left(\prod_{n=1}^N g_{n,j}^{A,P,\theta}\right) =  \nabla_\theta \log\left(\prod_{n=1}^N f_{Y_n|X_n}\left(y_n^* | x_{n,j}^{A, P,\theta}\right)\right) = \nabla_\theta \log f_{Y_{1:N}|X_{1:N}}\left(y_{1:N}^* | x_{1:n,j}^{A, P,\theta}\right),$$

and 
$$\nabla_\theta w_{N,j}^{P,\theta} = w_{N,j}^{P,\theta} \sum_{n=1}^N \nabla_\theta \log g_{n,j}^{A,P,\theta} = w_{N,j}^{P,\theta} \nabla_\theta \log f_{Y_{1:N}|X_{1:N}}\left(y_{1:N}^* | x_{1:n,j}^{A, P,\theta}\right).$$

Substituting, we have that
\begin{equation*}
    \nabla_\theta \hat{\ell}(\theta) := \frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,j}^{F,\theta} =\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,k_j}^{P,\theta} = \frac{1}{J}\sum_{j=1}^J w_{N,k_j}^{P,\theta} \nabla_\theta \log f_{Y_{1:N}|X_{1:N}}\left(y_{1:N}^* | x_{1:n,k_j}^{A, P,\theta}\right),
\end{equation*}
and finally, observing that when $\theta=\phi$ we have that $w_{N,j}^{F,\theta}=1$, that 
\begin{equation*}
    \nabla_\theta \hat{\ell}(\theta) := \frac{1}{J}\sum_{j=1}^J \nabla_\theta \log f_{Y_{1:N}|X_{1:N}}\left(y_{1:N}^* | x_{1:n,j}^{A, F,\theta}\right),
\end{equation*}
yielding the gradient estimators of \cite{poyiadjis11, scibior2021dpf} when applied to the bootstrap filter. 
\end{proof}

Note that one would believe that the variance blows up the moment $\theta\neq\phi$. This is true, as 
\begin{equation*}
    \nabla_\theta \hat{\ell}(\theta) := \frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,j}^{F,\theta} =\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,k_j}^{P,\theta} = \frac{1}{J}\sum_{j=1}^J w_{N,k_j}^{P,\theta} \nabla_\theta \log f_{Y_{1:N}|X_{1:N}}\left(y_{1:N}^* | x_{1:n,k_j}^{A, P,\theta}\right),
\end{equation*}
and when $\theta\neq\phi$, $w_{N,k_j}^{P,\theta} = O(c^N)$. When $\theta=\phi$, this is a special case of the \cite{poyiadjis11} estimator, which has $O(N^2)$ variance by a property of functionals of the particle filter by \cite{delMoral03}.

\begin{lem}
\label{lem:mop-0-formula}

Consider the case of MOP-$\alpha$ when $\alpha=0$ and $\theta=\phi$. The gradient estimate is then

    \begin{equation}
        \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right),
    \end{equation}

    yielding the gradient estimator of \cite{blei2018vsmc} that \cite{scibior2021dpf} show is the gradient of a vanilla particle filter when the resampling terms are dropped -- when applied to the bootstrap filter.

\end{lem}

\begin{proof}
    When $\alpha=0,$ the likelihood estimate becomes
\begin{equation}
    \hat{\lik}(\theta) := \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n,j}^{P,\theta}} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J s_{n,j} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J \frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \theta})}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi})}.
\end{equation}

Its gradient when $\theta=\phi$ is therefore 
\begin{align*}
    \nabla_\theta \hat{\ell}(\theta) &:= \sum_{n=1}^N \nabla_\theta \log\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right) \\
    &= \sum_{n=1}^N \frac{\nabla_\theta \left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right)}{\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right)} \\
    &= \sum_{n=1}^N \frac{\sum_{j=1}^J \nabla_\theta s_{n,j}}{\sum_{j=1}^J s_{n,j}} \\
    &= \sum_{n=1}^N \frac{1}{J} \sum_{j=1}^J \frac{\nabla_\theta f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)}{f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \phi}; \phi)} \\
    &= \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right),
\end{align*}
where we use the log-derivative trick, observe that $\sum_{j=1}^J s_{n,j} = J$ when $\theta=\phi$, and use the log-derivative trick where $\theta=\phi$ again. 
\end{proof}

\kevin{I have no clue how to prove this. Help!}
\begin{conj}[MOP-1 Estimate Is Unbiased for the Likelihood and the Gradient of the Likelihood]
    Consider the case of MOP-$\alpha$ when $\alpha=1$, and resampling is carried out with either multinomial, residual, or systematic resampling at every timestep. Assume the gradient estimate obtained by the gradient of the likelihood estimate with J particles, $\nabla_\theta \hat\lik_J(\theta)$ is uniformly bounded for all $J$, and that the likelihood has two continuous derivatives. Then the likelihood estimate is unbiased for the likelihood and the gradient of the likelihood estimate is unbiased for the gradient of the likelihood. That is, it holds that
    $$\E[\hat\lik_J(\theta)] = \lik(\theta),\;\; \E[\nabla_\theta \hat\lik_J(\theta)] = \nabla_\theta \lik(\theta).$$
\end{conj}

\begin{proof}
    First, we need to establish the unbiasedness of the likelihood estimate. This is 
    $$\hat\lik_J(\theta) := \left(\frac{1}{J}\sum_{j=1}^J w_{N,j}^{F,\theta}\right)\prod_{n=1}^N L_n^\phi = \left(\frac{1}{J}\sum_{j=1}^J \prod_{n=1}^N \frac{g_{n,j}^{A,\theta}}{g_{n,j}^{A,\phi}} \right)\prod_{n=1}^N L_n^\phi.$$

    We know from \cite{chopin20} that $\E[L_n^{\phi}] = \lik(\phi).$ It then suffices to show that
    $$\E\left[\left(\frac{1}{J}\sum_{j=1}^J \prod_{n=1}^N \frac{g_{n,j}^{A,\theta}}{g_{n,j}^{A,\phi}} \right)\right] = \frac{\lik(\theta)}{\lik(\phi)}.$$

    To obtain this, we need to see that the resampling under $\phi$, with corrected weights, is unbiased. That is, we need to show that for any bounded measurable functional $\varphi : \gX \to \R$,
    $$\E\left[\frac{\sum_{j=1}^J w_{n,j}^{F, \theta} \varphi(x_{n,j}^{\theta, F})}{\sum_{j=1}^J w_{n,j}^{F, \theta}} \bigg| \gF_n \right] = \frac{\sum_{j=1}^J w_{n,j}^{P, \theta} \varphi(x_{n,j}^{P, \theta})}{\sum_{j=1}^J w_{n,j}^{P, \theta} }. $$

    To see this, 
    \begin{align*}
        \sum_{j=1}^J w_{n,j}^{P, \theta} \varphi(x_{n,j}^{P, \theta}) 
        &= \sum_{j=1}^J w_{n-1,j}^{F, \theta}\frac{g_{n,j}^{\theta}}{g_{n,j}^{\phi}}  \varphi(x_{n,j}^{P, \theta})
    \end{align*}

%     Observe that our differentiable simulator $\process(\cdot, \theta)$ and choice of weights now satisfy:

%     - A Feynman-Kac model $\left\{M_t, G_t\right\}$ such that:

% - the weight function $G_t$ may be evaluated pointwise (for all $t$ );

% - it is possible to simulate from $\mathbb{M}_0\left(\mathrm{~d} x_0\right)$ and from $M_t\left(x_{t-1}, \mathrm{~d} x_t\right)$ (for any $x_{t-1}$ and $t$ ).

% - The number of particles $N$.

% - The choice of an unbiased resampling scheme (e.g. multinomial, residual, stratified, systematic), that is, an algorithm to generate variables $A_t^{1: N}$ in $1: N$ such that $\mathbb{E}\left[\sum_{m=1}^N \mathbbm{1}\left\{A_t^m=n\right\}\right]=N W_{t-1}^n$; see previous chapter for more details.
\end{proof}

\begin{lem}[Consistency of MOP-1 Gradient Estimate]
    Consider the case of MOP-$\alpha$ when $\alpha=1$. Assume the gradient estimate obtained by the gradient of the log-likelihood estimate with J particles, $\nabla_\theta \hat\ell_J(\theta) = \frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,j}^{F,\theta}$ is uniformly bounded by $G(\theta)$ for all $J$, and that the likelihood has two continuous derivatives. Then the gradient estimate, $\nabla_\theta \hat\ell_J(\theta)$, is consistent for the score.
    \label{lem:mop-grad-consistency}
\end{lem}
\begin{proof}

    % We know that the variance of the MOP-1 gradient estimate $\nabla_\theta \hat\ell_J(\theta) = \frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{N,j}^{F,\theta}$ goes to zero coordinate-wise as $J\to\infty$ by the particle filter law of large numbers for bounded functionals of particles (\kevin{Prove general LLN for functionals of MOP under boundedness assumptions in L2, but maybe Hoeffding may be enough to show this, because we only want convergence in probability, we just need to show the expectation converges}), so it suffices to show that it is asymptotically unbiased.

    % Write $\hat\pi_{n,J}$ for the filtering distribution at timestep $n$ and with $J$ particles, $\hat\pi_{1:N,J}$ for the sequence of filtering distributions from timesteps $1$ to $N$, and $\pi_n, \pi_{1:N}$ for the respective posteriors.

    % From Lemma \ref{lem:lik-proper-weight}, we know that $f_{Y_n|Y_{1:n-1}}(y_n^*|y_{1:n-1}^*, \theta)$ is consistently estimated by the conditional likelihood estimate obtained by MOP-1, for all $\theta$ in an open neighborhood of $\phi$. The likelihood is simply the product of the conditional likelihoods, and similarly for the estimates, so the likelihood $\lik(\theta)$ is also consistently estimated by the likelihood estimate obtained by MOP-1, $\hat\lik_J(\theta)$, for all $\theta$ in an open neighborhood of $\phi$, $\gN(\phi)$.

    % We therefore have that 
    % $$\lim_{J\to\infty} \E[\hat\lik_J(\theta)] = \lik(\theta)$$
    % pointwise for all $\theta \in \gN(\phi)$. Given that the gradient is bounded uniformly for all $J$ and consistency for the likelihood estimate holds for all $\theta$ in an open neighborhood of $\phi$ (and therefore all $\theta'$ in a sufficiently small neighborhood of $\theta$), the bounded convergence theorem allows us to switch the expectation and derivative:
    % $$\nabla_\theta \E[\hat\lik_J(\theta)] 
    % = \E[\nabla_\theta \hat\lik_J(\theta)].$$

    % Take limits on both sides to find that 
    % $$\lim_{J \to \infty} \nabla_\theta \E[\hat\lik_J(\theta)] 
    % = \lim_{J \to \infty} \E[\nabla_\theta \hat\lik_J(\theta)].$$

    % We then need to swap the limit and derivative on the left-hand side. This requires uniform convergence of $\lim_{J \to \infty} \nabla_\theta \E[\hat\lik_J(\theta)]$, i.e. that $$\sup_{\theta \in \gN(\phi)} ||  \nabla_\theta \E[\hat\lik_J(\theta)] - \lim_{J \to \infty} \nabla_\theta \E[\hat\lik_J(\theta)]||_{\infty} \to 0 \text{ as } J \to \infty,$$
    % or that $\forall \epsilon>0$ $\exists N\in \mathbb{N}$ such that $n>N => ||  \nabla_\theta \E[\hat\lik_J(\theta)] - \lim_{J \to \infty} \nabla_\theta \E[\hat\lik_J(\theta)]||_{\infty} < \epsilon$ $\forall \theta \in \gN(\phi).$ However, it is easier to instead use an indirect argument.

    % We then need to swap the limit and derivative on the left-hand side. This requires uniform convergence.

    Fix $\omega \in \Omega$, and set $\phi = \theta$, where $\theta$ is the point we wish to evaluate the gradient at. The sequence $(\nabla_\theta \hat\lik_J(\theta, \omega))_{J \in \mathbb{N}}$ is uniformly bounded over all $J$, by Assumption \ref{assump:local-bounded-derivative}. The sequence is also uniformly equicontinuous. To see this, by Assumption \ref{assump:local-bounded-derivative}, the second derivative of $\hat\lik_J(\theta, \omega)|_{\theta=\theta'}$ is also bounded by $H$ for almost every $\omega\in \Omega$ and every $\theta' \in \gN(\phi)$, including $\theta=\phi$. A set of functions with derivatives bounded by the same constant is uniformly Lipschitz, and therefore uniformly equicontinuous. So the sequence $(\nabla_\theta \hat\lik_J(\theta, \omega))_{J \in \mathbb{N}}$ is uniformly equicontinuous for almost every $\omega \in \Omega$. 
    Explicitly, for almost every $\omega \in \Omega$ and every $\epsilon>0$, there exists some $\delta(\omega)>0$ such that for every $||\theta - \theta'||_{\infty}<\delta$ and every $J \in \mathbb{N}$ we have that
    $$||\nabla_\theta \hat\lik_J(\theta, \omega)-\nabla_\theta \hat\lik_J(\theta', \omega)||_\infty < \epsilon.$$

    \kevin{Our previous understanding was that the target is differentiable, and the sequence is differentiable in increasingly small open neighborhoods. This may not be true. The neighborhoods the sequence is differentiable in should be independent of $J$, as the resampling does not change with $\theta$ as it is dependent on $\phi$. Either way, the target is in each of the neighborhoods, and this may be sufficient in itself.}

    Then, by Arzela-Ascoli, there exists a uniformly convergent subsequence. We claim that there is only one subsequential limit. We know that when as the gradient is bounded by Assumption \ref{assump:local-bounded-derivative} we can treat the gradient as a bounded functional. So by Lemma \ref{lem:posterior-proper-weights} the sequence $(\nabla_\theta \hat\lik_J(\theta, \omega))_{J \in \mathbb{N}}$ converges pointwise for $\theta=\phi$ and almost every $\omega \in \Omega$, and there is therefore only one subsequential limit. The sequence therefore converges uniformly to its limit $\lim_{J \to \infty} \nabla_\theta \hat\lik_J(\theta, \omega).$ Therefore, with uniform convergence for the derivatives established, we can swap the limit and derivative, and obtain that for almost every $\omega \in \Omega$, 
    $$\lim_{J \to \infty} \nabla_\theta \hat\lik_J(\theta, \omega) = \nabla_\theta \lim_{J \to \infty} \hat\lik_J(\theta, \omega).$$

    From Lemma \ref{lem:lik-proper-weight}, we know that
    $$\hat{\lik}_J(\theta, \omega) \stackrel{a.s.}{\to} \lik(\theta).$$

    We then have that for almost every $\omega \in \Omega$, 
    $$\lim_{J \to \infty} \nabla_\theta \hat\lik_J(\theta, \omega) = \nabla_\theta \lim_{J \to \infty} \hat\lik_J(\theta, \omega) = \nabla_\theta \lik(\theta),$$
    as we wanted. The result then follows by the continuous mapping theorem.

    % Taking expectations on both sides and using the bounded convergence theorem to interchange limits and expectations, we see that
    % $$\lim_{J \to \infty} \nabla_\theta \E[\hat\lik_J(\theta)] = \E[\lim_{J \to \infty} \nabla_\theta \hat\lik_J(\theta)] = \E[\nabla_\theta \lim_{J \to \infty} \hat\lik_J(\theta)] = \nabla_\theta \lim_{J \to \infty} \E[\hat\lik_J(\theta)].$$

    % It then follows that
    % $$\nabla_\theta\lik(\theta) 
    % = \nabla_\theta \lim_{J\to\infty} \E[\hat\lik_J(\theta)] 
    % = \lim_{J\to\infty} \nabla_\theta \E[\hat\lik_J(\theta)] 
    % = \lim_{J\to\infty} \E[\nabla_\theta \hat\lik_J(\theta)].$$

    % We then have mean-square consistency for the gradient of the likelihood estimate. The result then follows by the continuous mapping theorem.

    \kevin{Explain that the difference in formula between this and Poyiadjis is due to us taking the derivative w.r.t. the particles in the chain rule. }

    \kevin{Alternative argument: Show gradient is unbiased, after showing the likelihood is unbiased in an open neighborhood of phi. That should hold by Chopin's argument for unbiased resampling to nonconstant weights. Gradient unbiased if can swap expectation and derivative, I think with two continuous derivatives and boundedness, might need BCT/DCT. No need to swap limit and derivative in this case, or worrying about uniform limits.}

\end{proof}

\kevin{Show bias is order N, variance is order N when alpha below 1. That implies optimal choice of alpha is not 1. }

\section{Analysis of MOP-$\alpha$ Error and Variance}

\subsection{Analysis of MOP-$0$ Variance}

We made Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} to leverage the results from \cite{karjalainen2023} on the forgetting of the particle filter. This is required to show error bounds on the gradient estimates that we provide -- namely that the error of the MOP-$1$ estimator that corresponds to that of \cite{poyiadjis11} is $O(N^2)$, and that the variance of MOP-$\alpha$ for any $\alpha<1$ is $O(N)$.

We can decompose the MOP-$\alpha$ estimator as follows. When $\theta=\phi$,
$$\hat{\mathcal{L}}(\theta):=\prod_{n=1}^N L_n^{A, \theta, \alpha}=\prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n, j}^{F, \theta}}{\sum_{j=1}^J w_{n, j}^{P, \theta}}=\prod_{n=1}^N L_n^{A, \theta, \alpha}=\prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n, j}^{F, \theta}}{\sum_{j=1}^J (w_{n-1, j}^{F, \theta})^\alpha}.$$

To illustrate the proof strategy, we first analyze the special case when $\alpha=0.$ We then have:
\begin{equation}
    \hat{\lik}(\theta) := \prod_{n=1}^N L_n^{A, \theta, \alpha} = \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta}}{\sum_{j=1}^J w_{n,j}^{P,\theta}} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J s_{n,j} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J \frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \theta})}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi})}.
\end{equation}

Similarly to the proof of Lemma \ref{lem:mop-0-formula}, its gradient when $\theta=\phi$ is therefore 
\begin{align*}
    \nabla_\theta \hat{\ell}(\theta) &:= \sum_{n=1}^N \nabla_\theta \log\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right) \\
    &= \sum_{n=1}^N \frac{\nabla_\theta \left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right)}{\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right)} \\
    &= \sum_{n=1}^N \frac{\sum_{j=1}^J \nabla_\theta s_{n,j}}{\sum_{j=1}^J s_{n,j}} \\
    &= \sum_{n=1}^N \frac{1}{J} \sum_{j=1}^J \frac{\nabla_\theta f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)}{f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \phi}; \phi)} \\
    &= \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right),
\end{align*}
where we use the log-derivative trick, observe that $\sum_{j=1}^J s_{n,j} = J$ when $\theta=\phi$, and use the log-derivative trick where $\theta=\phi$ again. The point here was that we establish that the gradient of the log-likelihood estimate is given by the sum of terms over all $N$ and $J$:
$$\nabla_\theta \hat{\ell}(\theta) = \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right).$$

Therefore,
\begin{align*}
    \text{Var}(\nabla_\theta \hat\ell(\theta)) &= \frac{1}{J^2}\text{Var}\left(\sum_{n=1}^N\sum_{j=1}^{J}\nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right)\right) \\
    &= \frac{1}{J^2}\sum_{n=1}^N\text{Var}\left(\sum_{j=1}^{J}\nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right)\right) \\
    &+ 2\sum_{m<n}\text{Cov}\left(\frac{1}{J}\sum_{j=1}^{J}\nabla_\theta \log\left(f_{Y_m|X_{m}}(y_m^*|x_{m,j}^{F, \theta}; \theta)\right), \frac{1}{J}\sum_{j=1}^{J}\nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right)\right) \\
    &= \sum_{n=1}^N \text{Var}(\nabla_\theta\hat\ell_n(\theta)) + 2\sum_{m<n} \text{Cov}(\nabla_\theta\hat\ell_m(\theta),\nabla_\theta\hat\ell_n(\theta)).
\end{align*}

Here, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen2023} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
$$\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},$$
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}.

\kevin{How does the dobrushin contraction coefficient relate to the alpha-mixing coefficients? If you have a bound on the mixing time $b$, then the alpha mixing coefficients should just be $\alpha(t) \leq e^{-t/b}$.}

By Lemma \ref{lem:dobrushin-implies-alpha-mixing}, the particle filter itself is strongly mixing, with $\alpha$-mixing coefficients $\alpha(k) \leq (1-\epsilon)^{\floor{k/(c\log(J))}}$. Therefore, functions of particles are strongly mixing as well, with $\alpha$-mixing coefficients bounded by the original (to see this, observe that the $\sigma$-algebra of the functionals is contained within the original $\sigma$-algebra). Therefore, by Davydov's inequality, noting that $\nabla_\theta\hat\ell_n(\theta)\leq G'(\theta)$ by Assumption \ref{assump:bounded-measurement}, and WLOG assuming $\E[(\nabla_\theta\hat\ell_m(\theta))^4]^{1/4}\leq\E[(\nabla_\theta\hat\ell_n(\theta))^4]^{1/4}$, we see that
\begin{align*}
    \text{Cov}(\nabla_\theta\hat\ell_m(\theta), \nabla_\theta\hat\ell_n(\theta)) 
    &\leq \alpha(n-m)^{1/2}\E[(\nabla_\theta\hat\ell_m(\theta))^4]^{1/4}\E[(\nabla_\theta\hat\ell_n(\theta))^4]^{1/4}\\
    &\leq \alpha(n-m)^{1/2}\E[(\nabla_\theta\hat\ell_n(\theta))^4]^{1/2}.
\end{align*}

To bound this, we use the fact that $$\E[(\nabla_\theta\hat\ell_n(\theta))^4] = \E[(\nabla_\theta\hat\ell_n(\theta)-\E[\nabla_\theta\hat\ell_n(\theta)])^4]+\E[(\nabla_\theta\hat\ell_n(\theta))^2]^2$$
alongside Lemma 2 of \cite{karjalainen2023}, which shows that
$$\E[(\nabla_\theta\hat\ell_n(\theta)-\E[\nabla_\theta\hat\ell_n(\theta)])^4] \lesssim \frac{G'(\theta)^4}{J^2}, \;\; \E[(\nabla_\theta\hat\ell_n(\theta)-\E[\nabla_\theta\hat\ell_n(\theta)])^2] \lesssim \frac{G'(\theta)^2}{J},$$
to find that 
$$\E[(\nabla_\theta\hat\ell_n(\theta))^4]^{1/2} \lesssim  \sqrt{\frac{G'(\theta)^4}{J^2}+\left(\frac{G'(\theta)^2}{J}\right)^2} = \frac{G'(\theta)^2}{J},$$
and conclude that 
$$\text{Cov}(\nabla_\theta\hat\ell_m(\theta), \nabla_\theta\hat\ell_n(\theta)) \leq (1-\epsilon)^{\frac{1}{2}\floor{\frac{|n-m|}{c\log(J)}}}\frac{G'(\theta)^2}{J}.$$

Putting it all together, we see that
\begin{align*}
    \text{Var}(\nabla_\theta \hat\ell(\theta)) &= \sum_{n=1}^N \text{Var}(\nabla_\theta\hat\ell_n(\theta)) + 2\sum_{m<n} \text{Cov}(\nabla_\theta\hat\ell_m(\theta),\nabla_\theta\hat\ell_n(\theta))\\
    &\leq \frac{NG'(\theta)^2}{J} + 2\sum_{n=1}^{N-1} (N-n)(1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c\log(J)}}}\frac{G'(\theta)^2}{J} \\
    &\leq \frac{G'(\theta)^2}{J} \left(N + 2\sum_{n=1}^{N-1} (N-n) (1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c\log(J)}}}\right) \\
    &\lesssim \frac{G'(\theta)^2N}{J}.
\end{align*}

\paragraph{Remark:} Note that the factor of $N$ that pops up here is due to the use of the unnormalized gradient. If one divides the gradient estimate by $\sqrt{N}$ before usage, the variance does not depend on the horizon. If one divides by $N$, the error is $O(1/\sqrt{NJ})$!

\kevin{https://math.stackexchange.com/questions/3274996/is-a-transformed-mixing-process-is-still-mixing}

%$$\text{Cov}(x_{n,1:J}^P, x_{m, 1:J}^P) \leq (1-\epsilon)^{\floor{\frac{|n-m|}{c\log(J)}}}$$

\kevin{Ed: Claim 11.4.2, Chopin, is the key to reasoning about this. Establish geometric mixing, that $\text{Cov}(\nabla_\theta\hat\ell_n, \nabla_\theta\hat\ell_m) \leq \rho^{|n-m|}V_n$, where $V_n$ is asymptotically stable (the latter only holds for $\alpha<1$, as only then do we get asymptotic stability, otherwise it increases linearly in $n$). Geometrically decreasing covariances and stable variances over time are the same thing. Then, we can bound the total variance by $\text{Var}(\sum_n \nabla_\theta \hat\ell_n(\theta)) = \sum_{m,n} \text{Cov}(\nabla_\theta\hat\ell_n(\theta), \nabla_\theta\hat\ell_m(\theta)) = NV_n + 2(N-1)\rho V_n + 2(N-3)\rho^2V_n + ...$}

\subsection{Analysis of MOP-$\alpha$ Variance}

Next, we analyze the case for $\alpha \in (0,1)$. The proof strategy is as follows. Instead of analyzing the variance of the MOP-$\alpha$ gradient estimate proper, we will analyze the variance of a modified estimator where we truncate the weights at $k$ timesteps, so this estimator only looks at most $k$ timesteps back while accumulating the importance weights. This enables us to establish strong mixing for the truncated estimator, in order to get a bound on the variance that is $O(NJ^{-1})$. We emphasize that this estimator is only a theoretical construct -- we do not actually use this estimator.

Coupled with a bound ensuring that the $k$-truncated estimator provides an estimate that is close to that of MOP-$\alpha$ proper, we get a bound on the variance of MOP-$\alpha$ that comprises of the variance of the $k$-truncated estimator plus the error, for any $k \leq N$. It then holds that the final variance bound on MOP-$\alpha$ is given by the minimum over all $k$, and is never larger than $O(1+NJ^{-1})$.

Using the log-derivative trick and that $w_{n,j}^{F,\theta} = w_{n,j}^{F,\theta,k} = 1$ when $\theta=\phi$,
\begin{align*}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,k}\right)\right\rVert_{\infty}\\
    &= \left\lVert\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta}}{{\sum_{j=1}^J w_{n,j}^{F,\theta}}}-\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,k}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,k}}}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta}-\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,k}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta}}{w_{n,j}^{F,\theta}}-\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,k}}{w_{n,j}^{F,\theta,k}}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta}\right)-\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,k}\right)\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\log\left(w_{n,j}^{F,\theta}\right)-\log\left(w_{n,j}^{F,\theta,k}\right)\right)\right\rVert_{\infty}\\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=1}^n\alpha^{(n-i)}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=n-k}^n{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=1}^{n-k}\alpha^{(n-i)}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) \right)\right\rVert_{\infty}\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_{\infty}\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}G'(\theta)\\
    &\leq G'(\theta)\frac{\alpha^k-\alpha^n}{1-\alpha},
\end{align*}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}.

We can then bound $||\nabla_\theta\hat\ell(\theta) - \hat s_k(\theta)||$:
\begin{align*}
    &\left\lVert\nabla_\theta\hat\ell(\theta) - \hat s_k(\theta) \right\rVert_\infty\\
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \log\left(\frac{\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) - \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) \right\rVert_\infty\\
    &= \Bigg\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)- \log\left(\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)\\
    &\qquad\qquad -\log\left(\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right) + \log\left(\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)\Bigg)\Bigg\rVert_\infty \\
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,k}\right)
    -\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,k}\right)\Bigg)\right\rVert_\infty \\
    &\leq \sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,k}\right)\right)\right\lVert_{\infty}
    +\sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,k}\right)\right)\right\rVert_\infty,
\end{align*}
and each of these two terms is bounded by $\frac{\alpha^k}{1-\alpha}NG'(\theta)$. So we now know that
$$\left\lVert\nabla_\theta\hat\ell(\theta) - \hat s_k(\theta) \right\rVert_\infty \leq  \frac{2\alpha^k}{1-\alpha}NG'(\theta),$$
which is our desired error bound.

Now, we bound the variance of $\hat s_k(\theta)$. 
Recall that we defined
$$\hat s_k(\theta) = \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) = \sum_{n=1}^N \nabla_\theta \left(\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta, k}\right)-\log\left(\sum_{j=1}^J w_{n-1,j}^{A, F,\theta, k}\right)\right).$$

We can then decompose this expression with the log-derivative trick while noting that $w_{n,j}^{F,\theta, k} = 1$ whenever $\theta=\phi$, to see that
\begin{align*}
    \hat s_k(\theta) 
    &= \sum_{n=1}^N \nabla_\theta \left(\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta, k}\right)-\log\left(\sum_{j=1}^J w_{n-1,j}^{A, F,\theta, k}\right)\right) \\
    &= \sum_{n=1}^N \left(\frac{\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta, k}}{\sum_{j=1}^J w_{n,j}^{F,\theta, k}}-\frac{\sum_{j=1}^J \nabla_\theta w_{n-1,j}^{A,F,\theta, k}}{\sum_{j=1}^J w_{n-1,j}^{A,F,\theta, k}}\right) \\
    &= \sum_{n=1}^N \left(\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta, k}-\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n-1,j}^{A,F,\theta, k}\right)\\
    &= \frac{1}{J}\sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \left(w_{n,j}^{F,\theta, k}- w_{n-1,j}^{A,F,\theta, k}\right).
\end{align*}

It now follows that we need to bound the variance at a single timestep, namely
$$\text{Var}\left(\frac{1}{J}\sum_{j=1}^J\nabla_\theta \left(w_{n,j}^{F,\theta, k}- w_{n-1,j}^{A,F,\theta, k}\right)\right).$$

We use the log-derivative trick yet again to find, noting that $w_{n,j}^{F,\theta, k}=1$ when $\theta=\phi$, that
\begin{align*}
    \nabla_\theta w_{n,j}^{F,\theta, k} = \frac{\nabla_\theta w_{n,j}^{F,\theta, k}}{w_{n,j}^{F,\theta, k}} = \nabla_\theta \log(w_{n,j}^{F,\theta, k}) = \sum_{i=n-k}^n \alpha^{n-i} \nabla_\theta \log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) = \sum_{i=n-k}^n \alpha^{n-i} \nabla_\theta \log\left(g_{i,j}^{A,F,\theta}\right).
\end{align*}

Then,
\begin{align*}
    \text{Var}\left(\nabla_\theta \left(w_{n,j}^{F,\theta, k}- w_{n-1,j}^{A,F,\theta, k}\right)\right) 
    &= \text{Var}\left(\nabla_\theta \left(\sum_{i=n-k}^n \alpha^{n-i} \nabla_\theta \log\left(g_{i,j}^{A,F,\theta}\right)- \sum_{i=n-k-1}^{n-1} \alpha^{n-i} \nabla_\theta \log\left(g_{i,j}^{A,F,\theta}\right)\right)\right) \\
    &= \text{Var} \left(\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1}\log\left(g_{i-1,j}^{A,F,\theta}\right)\right)\right).
\end{align*}

Note that $$\hat s_{n,k}(\theta) := \frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \nabla_\theta \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)$$ is a function bounded by $C\frac{G'(\theta)(1-\alpha^k)}{1-\alpha} =: CG'(\theta)r$ in each coordinate for some constant $C$ (by Assumption \ref{assump:bounded-measurement}) that depends only on $k$ timesteps, and not $n$. We can therefore invoke Lemma 2 of \cite{karjalainen2023} to bound the $L_2$ error of each of the $k$ (from time $n-k$ to time $n$) functionals from its expectation under the posterior by $CrG'(\theta)J^{-1/2}k$. That is, we have that
$$\E\left[||\hat s_{n,k}(\theta) - \E_{\pi}\hat s_{n,k}(\theta)||_\infty^2\right]^{1/2} \leq \frac{Cr^2G'(\theta) k}{\sqrt{J}}.$$

By the bias-variance decomposition, this in turn bounds the variance at a single timestep by $C^2r^2G'(\theta)^2J^{-1}k^2$. Concretely, we have that
$$\text{Var}\left(\frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)\right) \lesssim \frac{r^2k^2G'(\theta)^2}{J}.$$

It now remains to bound the variance of $\hat s_k(\theta)$ by considering the covariance of each of the $N$ terms that comprise it. We decompose
\begin{align*}
    \text{Var}(\nabla_\theta \hat s_k(\theta)) &= \text{Var}\left(\sum_{n=1}^N \frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)\right) \\
    &= \sum_{n=1}^N\text{Var}\left(\hat s_{n,k}(\theta)\right)+ 2\sum_{m<n}\text{Cov}\left(\hat s_{m,k}(\theta), \hat s_{n,k}(\theta)\right) \\
    &\lesssim \frac{r^2k^2G'(\theta)^2N}{J} + 2\sum_{m<n}\text{Cov}\left(\hat s_{m,k}(\theta), \hat s_{n,k}(\theta)\right).
\end{align*}

Similarly to the proof of the MOP-$0$ case, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen2023} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
$$\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},$$
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}. By Lemma \ref{lem:dobrushin-implies-alpha-mixing}, the particle filter itself is strongly mixing, with $\alpha$-mixing coefficients $\alpha(l) \leq (1-\epsilon)^{\floor{l/(c\log(J))}}$. Therefore, functions of particles are strongly mixing as well, with $\alpha$-mixing coefficients bounded by the original (to see this, observe that the $\sigma$-algebra of the functionals is contained within the original $\sigma$-algebra).

We now derive the $\alpha$-mixing coefficients of $(\hat s_{n,k}(\theta))_{n=1}^N$. Observe that $\hat s_{n,k}$ is strong mixing at lag $k+1$, as all weights beyond $k$ timesteps are truncated. We therefore have that the mixing time of $\hat s_k(\theta)$ is $O(1+k+\log(J))$, and that the $\alpha$-mixing coefficients for $(\hat s_{n,k}(\theta))_{n=1}^N$ are given by $\alpha(l) \leq (1-\epsilon)^{\floor{l/(c(1+k+\log(J)))}}$.

Therefore, by Davydov's inequality and Lemma 2 of \cite{karjalainen2023}, and noting that $\hat s_{n,k}(\theta) \leq CrG'(\theta)$ by Assumption \ref{assump:bounded-measurement}, by a similar argument to the MOP-0 case, we find that

$$\text{Cov}\left(\hat s_{m,k}(\theta), \hat s_{n,k}(\theta)\right) \lesssim (1-\epsilon)^{\frac{1}{2}\floor{\frac{|n-m|}{c(1+k+\log(J))}}}\frac{r^2G'(\theta)^2}{J}.$$

Concretely, noting that $\nabla_\theta\hat s_{n,k}(\theta)\leq CrG'(\theta)$ by Assumption \ref{assump:bounded-measurement}, and WLOG assuming $\E[(\nabla_\theta\hat\hat s_{m,k}(\theta))^4]^{1/4}\leq\E[(\nabla_\theta\hat\hat s_{n,k}(\theta))^4]^{1/4}$, we apply Davydov's inequality to see that
\begin{align*}
    \text{Cov}(\nabla_\theta\hat s_{m,k}(\theta), \nabla_\theta\hat s_{n,k}(\theta)) 
    &\leq \alpha(n-m)^{1/2}\E[(\nabla_\theta\hat s_{m,k}(\theta))^4]^{1/4}\E[(\nabla_\theta\hat s_{n,k}(\theta))^4]^{1/4}\\
    &\leq \alpha(n-m)^{1/2}\E[(\nabla_\theta\hat s_{n,k}(\theta))^4]^{1/2}.
\end{align*}

To bound this, we use the fact that $$\E[(\nabla_\theta\hat s_{n,k}(\theta))^4] = \E[(\nabla_\theta\hat s_{n,k}(\theta)-\E[\nabla_\theta\hat s_{n,k}(\theta)])^4]+\E[(\nabla_\theta\hat s_{n,k}(\theta))^2]^2$$
alongside Lemma 2 of \cite{karjalainen2023}, which shows that
$$\E[(\nabla_\theta\hat s_{n,k}(\theta)-\E[\nabla_\theta\hat s_{n,k}(\theta)])^4] \lesssim \frac{r^4G'(\theta)^4}{J^2}, \;\; \E[(\nabla_\theta\hat s_{n,k}(\theta)-\E[\nabla_\theta\hat s_{n,k}(\theta)])^2] \lesssim \frac{r^2G'(\theta)^2}{J},$$
to find that 
$$\E[(\nabla_\theta\hat s_{n,k}(\theta))^4]^{1/2} \lesssim  \sqrt{\frac{r^4G'(\theta)^4}{J^2}+\left(\frac{r^2G'(\theta)^2}{J}\right)^2} = \frac{r^2G'(\theta)^2}{J},$$
and conclude that 
$$\text{Cov}(\nabla_\theta\hat s_{m,k}(\theta), \nabla_\theta\hat s_{n,k}(\theta)) \leq (1-\epsilon)^{\frac{1}{2}\floor{\frac{|n-m|}{c(1+k+\log(J))}}}\frac{r^2G'(\theta)^2}{J}.$$

Putting it all together, we see that
\begin{align*}
    \text{Var}(\nabla_\theta \hat s_{k}(\theta)) &= \sum_{n=1}^N \text{Var}(\nabla_\theta\hat s_{n,k}(\theta)) + 2\sum_{m<n} \text{Cov}(\nabla_\theta\hat s_{m,k}(\theta),\nabla_\theta\hat s_{n,k}(\theta))\\
    &\leq \frac{Nr^2k^2G'(\theta)^2}{J} + 2\sum_{n=1}^{N-1} (N-n)(1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c(1+k+\log(J))}}}\frac{r^2G'(\theta)^2}{J} \\
    &\leq \frac{r^2k^2G'(\theta)^2}{J} \left(N + 2\sum_{n=1}^{N-1} (N-n) (1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c(1+k+\log(J))}}}\right) \\
    &\lesssim \frac{k^2r^2G'(\theta)^2N}{J}.
\end{align*}

We will now use this result to bound $\text{Var}(\nabla_\theta \hat s_k(\theta)$ proper. For random variables $X,Y$ where $|X-Y|$ is bounded almost surely by some $M$, we have that 
$$X-\E X - 2M \leq Y - \E Y \leq X - \E X + 2M$$
so,
$$(Y-\E Y)^2 \leq (X-\E X - 2M)^2 + (X-\E X +2M)^2.$$
Taking expectations, we get
$$\text{Var}(Y) \leq \text{Var}(X) + 8M^2.$$

It follows from this result and the result we proved earlier, namely that
$$||\nabla_\theta\hat\ell(\theta) - \hat s_k(\theta)||_\infty \leq \frac{2\alpha^k}{1-\alpha}NG'(\theta),$$
that the variance of MOP-$\alpha$ proper is bounded by, for any $k \leq N$,
that the variance of MOP-$\alpha$ proper is bounded by
$$\text{Var}(\nabla_\theta \hat\ell(\theta)) \lesssim \frac{k^2r^2G'(\theta)^2N}{J} + \frac{16\alpha^{2k}}{(1-\alpha)^2}N^2G'(\theta)^2.$$

As the above holds for any $k \leq N$,
$$\text{Var}(\nabla_\theta \hat\ell(\theta)) \lesssim \min_{k\leq N} \left(\frac{k^2G'(\theta)^2N}{(1-\alpha)^2J} + \frac{16\alpha^{2k}}{(1-\alpha)^2}N^2G'(\theta)^2 \right).$$

% To interpret this result, we can choose $k$ such that
% \begin{align*}
%     &\frac{16\alpha^{2k}}{(1-\alpha)^2}N^2G'(\theta)^2 \leq 1,\\
%     &\alpha^{2k} \leq \frac{(1-\alpha)^2}{16N^2G'(\theta)^2},\\
%     &k \geq \frac{\log\left(16N^2G'(\theta)^2\right)-2\log(1-\alpha)}{2\log\left(\frac{1}{\alpha}\right)}.
% \end{align*}

% To interpret this result, if there exists some $k$ such that $$N \geq k \geq \frac{\log\left(16NG'(\theta)\right)-\log(1-\alpha)}{\log\left(\frac{1}{\alpha}\right)},$$
% we can choose such a $k$, as well as $k=N$, to see that
% \begin{align*}
%     \text{Var}(\nabla_\theta \hat\ell(\theta)) 
%     &\lesssim \min_{k\leq N}\left(\frac{c_d^2k^2r^2G'(\theta)^2N}{J} + \frac{\alpha^{2k}}{(1-\alpha)^2}N^2G'(\theta)^2 \right)\\
%     &\lesssim \min\left\{1+\frac{c_d^2G'(\theta)^2N}{J}, \frac{c_d^2G'(\theta)^2r^2N^3}{J}\right\} \\
%     &\leq \min\left\{1+\frac{c_d^2G'(\theta)^2N}{J}, \frac{c_d^2G(\theta)^2r^2N}{J}\right\},
% \end{align*}
% as $NG'(\theta)\leq G(\theta)$ as a consequence of Assumption \ref{assump:local-bounded-derivative}.

\subsection{Analysis of MOP-$\alpha$ Bias}

Using the log-derivative trick and that $w_{n,j}^{F,\theta,1} = w_{n,j}^{F,\theta,\alpha} = 1$ when $\theta=\phi$,
\begin{align*}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,1}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}\right)\right\rVert_{\infty}\\
    &= \left\lVert\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,1}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,1}}}-\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}}}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,1}-\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,\alpha}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,1}}{w_{n,j}^{F,\theta,1}}-\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,\alpha}}{w_{n,j}^{F,\theta,\alpha}}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,1}\right)-\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,\alpha}\right)\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\log\left(w_{n,j}^{F,\theta,1}\right)-\log\left(w_{n,j}^{F,\theta,\alpha}\right)\right)\right\rVert_{\infty}\\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=1}^n\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=1}^n{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=1}^n\left(1-\alpha^{(n-i)}\right)\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) \right)\right\rVert_{\infty}.
\end{align*}

Now we establish the bound:
\begin{align*}
     \nabla_\theta \left(\sum_{i=1}^n\left(1-\alpha^{(n-i)}\right)\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) \right)
     &\leq \sum_{i=1}^n\left(1-\alpha^{(n-i)}\right)\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_{\infty}\\
    &\leq \sum_{i=1}^n\left(1-\alpha^{(n-i)}\right)G'(\theta)\\
    &\leq G'(\theta)\left(\frac{n-\alpha n + \alpha^n - 1}{1-\alpha}\right)
\end{align*}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}.

By a similar decomposition to that of the error of the $k$-truncated gradient to the MOP-$\alpha$ gradient estimate,
\begin{align*}
    &\left\lVert\nabla_\theta\hat\ell^1(\theta) - \nabla_\theta\hat\ell^\alpha(\theta) \right\rVert_\infty\\
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \log\left(\frac{\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)}{\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)}\right) - \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) \right\rVert_\infty\\
    &= \Bigg\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)- \log\left(\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\\
    &\qquad\qquad -\log\left(\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right) + \log\left(\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)\Bigg)\Bigg\rVert_\infty \\
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha}\right)
    -\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha}\right)\Bigg)\right\rVert_\infty \\
    &\leq \sum_{n=1}^N \left\lVert\nabla_\theta \Bigg(\left|\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha}\right)\right|
    +\left|\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha}\right)\right|\Bigg)\right\rVert_\infty.
\end{align*}

From our arguments above, we know that 
$$\left\lVert\nabla_\theta\hat\ell^1(\theta) - \nabla_\theta\hat\ell^\alpha(\theta) \right\rVert_\infty \leq 2NG'(\theta)\left(\frac{N-\alpha N + \alpha^N - 1}{1-\alpha}\right).$$

So the bias of MOP-$\alpha$ is the bias of MOP-$1$ arising from the logarithm in the gradient of the log-likelihood plus no more than $$2NG'(\theta)\left(\frac{N-\alpha N + \alpha^N - 1}{1-\alpha}\right),$$
where we note that the bias vanishes as $\alpha$ goes to $1$ as
$$\lim_{\alpha\to1}\left(\frac{N-\alpha N + \alpha^N - 1}{1-\alpha}\right) = 0.$$

\kevin{This subsection is not useless I think! It shows that the bias vanishes continuously as $\alpha$ goes to $1$.}

\subsection{An Improved Bound by Mixing}

We decompose the bias $||\nabla_\theta\ell(\theta) - \E\nabla_\theta \hat\ell^\alpha(\theta)||_\infty \leq \E||\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)||_\infty$ into
$$\E||\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)||_\infty \leq \E||\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)||_\infty + \E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_\infty + \E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_\infty.$$

The broad idea is that the first term can be controlled by a mixing argument, the second term garners $O(Nk)$ error, and the third term is $O\left(\frac{N}{1-\alpha}\right)$. It is unclear whether the bias can be reduced further -- the gradient is a sum of $N$ terms, and is therefore $O(N)$ itself.

The first term is the problem term. We will first bound the ground truth of conditional scores from particle approximation conditional on correct filtering distribution at time $n-k$, and then bound that from the particle approximation resulting from an arbitrary filtering distribution at time $n-k$. That is, where $\hat\pi_n$ is the particle approximation at time $n$ of the posterior $\pi_n$ and we write $\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)$ for the truncated MOP-$1$ conditional score estimate given the correct filtering distribution at time $n-k$,

\begin{align*}
    &\E||\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)||_\infty \\
    &\leq \sum_{n=1}^N \E||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_\infty \\
    &\leq \sum_{n=1}^N \E||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)||_\infty + \sum_{n=1}^N \E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_\infty.
\end{align*}

The first term is a particle approximation dependent on $k$ timesteps, so by Lemma 2 of \cite{karjalainen2023}, this is bounded by
$$\E||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)||_\infty \leq \frac{CG'(\theta)k}{\sqrt{J}}.$$

The second term amounts to bounding the difference between functionals of two different particle measures that mix under the same Markov kernel. Here, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen2023} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
$$\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},$$
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}.

Then, we can bound 
$\E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_\infty$ by
$$\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu \textbf{M}_{n,n+k}-\nu \textbf{M}_{n,n+k}\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}} = \beta_{TV}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},$$
implying that
\begin{align*}
    \E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_\infty 
    &\lesssim \sup_{\mu, \nu} \sup_{||\psi||_\infty \leq 1/2} G'(\theta)\left|(\mu \textbf{M}_{n,n+k})(\psi)-(\nu \textbf{M}_{n,n+k})(\psi)\right| \\
    &\leq \sup_{\mu, \nu} G'(\theta)\|\mu \textbf{M}_{n,n+k}-\nu \textbf{M}_{n,n+k}\|_{\mathrm{TV}} \\
    &\leq G'(\theta)(1-\epsilon)^{\floor{k/(c\log(J))}} ||\hat\pi_{n-k} - \pi_{n-k}||_{\text{TV}} \\
    &\leq G'(\theta)(1-\epsilon)^{\floor{k/(c\log(J))}} \sup_{||\psi||_{\infty} \leq 1/2} |\hat\pi_{n-k}(\psi) - \pi_{n-k}(\psi)| \\
    &\lesssim G'(\theta)(1-\epsilon)^{\floor{k/(c\log(J))}}.
\end{align*}

Therefore, we have that
\begin{align*}
    &\E||\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)||_\infty \\
    &\leq \sum_{n=1}^N \E||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)||_\infty + \sum_{n=1}^N \E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_\infty\\
    &\lesssim N\frac{CG'(\theta)k}{\sqrt{J}} + NG'(\theta)(1-\epsilon)^{\floor{k/(c\log(J))}}.
\end{align*}

Now that the first term $\E||\nabla_\theta\ell(\theta)$ is taken care of, it remains to bound $$\E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_\infty \text{ and } \E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_\infty.$$

We see, by a similar argument to the variance bound, that
\begin{align*}
    &\E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_\infty\\
    &= \E\left\lVert\sum_{n=1}^N \nabla_\theta \log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)}\right) - \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) \right\rVert_\infty\\
    &= \E\Bigg\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)- \log\left(\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\\
    &\qquad\qquad -\log\left(\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right) + \log\left(\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)\Bigg)\Bigg\rVert_\infty \\
    &= \E\left\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1,k}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)
    -\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1,k}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\Bigg)\right\rVert_\infty \\
    &\leq \sum_{n=1}^N \E\left\lVert\nabla_\theta \Bigg(\left|\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1,k}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right|
    +\left|\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1,k}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\right|\Bigg)\right\rVert_\infty.
\end{align*}

Each of these terms can be bounded using the log-derivative trick and that $w_{n,j}^{F,\theta,1,k} = w_{n,j}^{F,\theta,\alpha,k} = 1$ when $\theta=\phi$,
\begin{align*}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,1,k}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_{\infty}\\
    &= \left\lVert\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,1,k}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,1,k}}}-\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}}}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,1,k}-\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,\alpha,k}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,1,k}}{w_{n,j}^{F,\theta,1,k}}-\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,\alpha,k}}{w_{n,j}^{F,\theta,\alpha,k}}\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,1,k}\right)-\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\log\left(w_{n,j}^{F,\theta,1,k}\right)-\log\left(w_{n,j}^{F,\theta,\alpha,k}\right)\right)\right\rVert_{\infty}\\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=n-k}^n\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=n-k}^n{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\right\rVert_{\infty} \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=n-k}^{n}(1-\alpha^{(n-i)})\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) \right)\right\rVert_{\infty}\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_{\infty}\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})G'(\theta)\\
    &\leq G'(\theta)\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right),
\end{align*}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}.

So, 
\begin{align*}
    &\E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_\infty\\
    &\leq \sum_{n=1}^N \E\left\lVert\nabla_\theta \Bigg(\left|\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1,k}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right|
    +\left|\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1,k}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\right|\Bigg)\right\rVert_\infty \\
    &\leq 2NG'(\theta)\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right) 
    \leq 2NG'(\theta)k.
\end{align*}

Now, the third term, $\E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_\infty$, fortunately has already been dealt with. From the variance argument, we know that
$$||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_\infty \leq 2\frac{\alpha^k}{1-\alpha}NG'(\theta),$$
and taking expectations on both sides yields our desired result.

Therefore, our decomposition yields the bias bound
\begin{align*}
    &||\nabla_\theta\ell(\theta) - \E\nabla_\theta \hat\ell^\alpha(\theta)||_\infty \leq \E||\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)||_\infty \\
    &\leq \E||\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)||_\infty + \E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_\infty + \E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_\infty \\
    &\lesssim \min_{k \leq N} \left(N\frac{CG'(\theta)k}{\sqrt{J}} + NG'(\theta)(1-\epsilon)^{\floor{k/(c\log(J))}} + 2NG'(\theta)\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right) + 2\frac{\alpha^k}{1-\alpha}NG'(\theta)\right) \\
    &\lesssim \min_{k \leq N} NG'(\theta)\left(\frac{k}{\sqrt{J}}+(1-\epsilon)^{\floor{k/(c\log(J))}}+k+\frac{\alpha^k  + \alpha^{k+1} - \alpha}{1-\alpha}\right).
\end{align*}

\subsection{Bounding the MSE of MOP-$\alpha$}

\kevin{First term sums up $k$ terms, $O(Nk^2)$ error, second term bounded by $k$ applications of the Markov kernel, so just apply forgetting of the particle filter result. }

We can obtain an $O(N)$ MSE bound from an argument that is similar to the bias argument. Simply observe that 
$$||\nabla_\theta\ell_\theta - \nabla_\theta \hat\ell^\alpha(\theta)||_2^2 =  \left\lVert\sum_{n=1}^N \nabla_\theta \hat\ell_n^\alpha(\theta) - \sum_{n=1}^N \nabla_\theta \ell_n(\theta)\right\rVert_2^2 \leq \sum_{n=1}^N\left\lVert \nabla_\theta \hat\ell_n^\alpha(\theta) -  \nabla_\theta \ell_n(\theta)\right\rVert_2^2,$$

and decompose
$$\sum_{n=1}^N\left\lVert \nabla_\theta \hat\ell_n^\alpha(\theta) -  \nabla_\theta \ell_n(\theta)\right\rVert_2^2 \leq \sum_{n=1}^N||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2 + \sum_{n=1}^N||\nabla_\theta\hat s_{n,k}^1(\theta) - \nabla_\theta\hat s_{n,k}^\alpha(\theta)||_2^2 + \sum_{n=1}^N||\nabla_\theta\hat s_{n,k}^\alpha(\theta) -  \nabla_\theta\hat\ell_n^\alpha(\theta)||_2^2.$$

As before, we bound the first term, $\sum_{n=1}^N||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2$, by decomposing it into two terms, 
$$\sum_{n=1}^N||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2 \leq \sum_{n=1}^N ||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)||_2^2 + \sum_{n=1}^N ||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2.$$

The first term is a particle approximation dependent on $k$ timesteps, so by Lemma 2 of \cite{karjalainen2023}, this is bounded by
$$\E||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)||_2^2 \leq \frac{CpG'(\theta)^2k^2}{J}.$$

The second term amounts to bounding the difference between functionals of two different particle measures that mix under the same Markov kernel. Here, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen2023} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
$$\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},$$
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}.

Then, we can bound 
$\E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2$ by
$$\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu \textbf{M}_{n,n+k}-\nu \textbf{M}_{n,n+k}\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}} = \beta_{TV}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},$$
implying that
\begin{align*}
    \E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2 
    &\lesssim \sup_{\mu, \nu} \sup_{||\psi||_\infty \leq 1/2} pG'(\theta)^2\left|(\mu \textbf{M}_{n,n+k})(\psi)-(\nu \textbf{M}_{n,n+k})(\psi)\right| \\
    &\leq \sup_{\mu, \nu} pG'(\theta)^2\|\mu \textbf{M}_{n,n+k}-\nu \textbf{M}_{n,n+k}\|_{\mathrm{TV}} \\
    &\leq pG'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}} ||\hat\pi_{n-k} - \pi_{n-k}||_{\text{TV}} \\
    &\leq pG'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}} \sup_{||\psi||_{\infty} \leq 1/2} |\hat\pi_{n-k}(\psi) - \pi_{n-k}(\psi)| \\
    &\lesssim pG'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}}.
\end{align*}

Therefore, we have that
\begin{align*}
    &\E||\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)||_2^2 \\
    &\leq \sum_{n=1}^N \E||\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)||_2^2 + \sum_{n=1}^N \E||\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)||_2^2\\
    &\lesssim N\frac{CpG'(\theta)^2k^2}{J} + NpG'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}}.
\end{align*}

Now that the first term $\E||\nabla_\theta\ell(\theta)||_2^2$ is taken care of, it remains to bound $$\E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_2^2 \text{ and } \E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_2^2.$$

We see, by a similar argument to the bias bound, that
\begin{align*}
    &\E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_2^2\\
    &\leq \sum_{n=1}^N \E\left\lVert\nabla_\theta \Bigg(\left|\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1,k}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right|
    +\left|\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1,k}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\right|\Bigg)\right\rVert_2^2.
\end{align*}

Each of these terms can be bounded using the log-derivative trick and that $w_{n,j}^{F,\theta,1,k} = w_{n,j}^{F,\theta,\alpha,k} = 1$ when $\theta=\phi$,
\begin{align*}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,1,k}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_{2}^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})pG'(\theta)^2\\
    &\leq pG'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right),
\end{align*}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}.

So, 
\begin{align*}
    \E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_\infty
    \leq 2NpG'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right) 
    \leq 2NpG'(\theta)^2k.
\end{align*}

Now, we address the third term, $\E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_\infty$. From the variance argument, we know that we can bound $||\nabla_\theta\hat\ell^\alpha(\theta) - \hat s_k^\alpha(\theta)||$:
\begin{align*}
    &\left\lVert\nabla_\theta\hat\ell^\alpha(\theta) - \hat s_k^\alpha(\theta) \right\rVert_2^2\\
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \log\left(\frac{\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) - \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) \right\rVert_2^2\\
    &\leq \sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right)\right\lVert_2^2
    +\sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta, \alpha}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha, k}\right)\right)\right\rVert_2^2.
\end{align*}

Using the log-derivative trick and that $w_{n,j}^{F,\theta,\alpha} = w_{n,j}^{F,\theta,\alpha,k} = 1$ when $\theta=\phi$,
\begin{align*}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}pG'(\theta)^2\\
    &\leq pG'(\theta)^2\frac{\alpha^k-\alpha^n}{1-\alpha},
\end{align*}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}. Putting it together and taking expectations on both sides, we obtain
$$\E\left\lVert\nabla_\theta\hat\ell(\theta) - \hat s_k(\theta) \right\rVert_2^2 \leq  \frac{2\alpha^k}{1-\alpha}NpG'(\theta)^2,$$
which is our desired error bound.

Therefore, our decomposition yields the MSE bound
\begin{align*}
    &\E||\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)||_2^2 \\
    &\leq \E||\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)||_2^2 + \E||\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)||_2^2 + \E||\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)||_2^2 \\
    &\lesssim \min_{k \leq N} \left(N\frac{CpG'(\theta)^2k^2}{J} + NpG'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}} + 2NpG'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right) + 2\frac{\alpha^k}{1-\alpha}NpG'(\theta)^2\right) \\
    &\lesssim \min_{k \leq N} NpG'(\theta)^2\left(\frac{k^2}{J}+(1-\epsilon)^{\floor{k/(c\log(J))}}+k+\frac{\alpha^k  + \alpha^{k+1} - \alpha}{1-\alpha}\right).
\end{align*}

\kevin{Trying to bound MOP-(1,k) from MOP-alpha yields exactly the same result as above! Seems like I can't get rid of the k in the third term within the parentheses.}
% \begin{align*}
%     &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,1,k}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}\right)\right\rVert_{2}^2\\
%     &\leq \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=n-k}^n\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=1}^n{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\right\rVert_2^2 \\
%     &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=n-k}^n\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=n-k}^n{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=1}^{n-k}{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\right\rVert_2^2 \\
%     &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2 + \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2 \\
%     &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})pG'(\theta)^2 + \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}pG'(\theta)^2\\
%     &\leq pG'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)+\alpha^k}{1-\alpha}\right),
% \end{align*}

\section{Finite-Particle Analysis of Gradient Methods with Algorithm \ref{alg:ifad}}
\label{app:ifad}

The analysis in this section roughly follows the analysis in \cite{mahoney16}, except with the caveat that none of the matrix concentration bounds they use apply here as the particles are dependent. We instead use the concentration inequality from \cite{delmoral2011ci} to bound the gradient and Hessian estimates. In this section, we fix $\omega \in \Omega$ only within each filtering iteration, evaluate Algorithm \ref{alg:mop} at $\theta=\phi$, and analyze Algorithm \ref{alg:ifad} post-iterated filtering.

The convergence analysis in Theorem \ref{thm:convergence} is limited to the case where $-\ell$ is $\gamma$-strongly convex. Though it is true that in a neighborhood of the optimum local asymptotic normality holds and the log-likelihood is strongly convex in this neighborhood, in practice likelihood surfaces for POMPs are often highly nonconvex globally. The convergence to an optimum, local or global, must therefore be sensitive to initialization.

\subsection{Bounding the Gradient}
\kevin{Add a few words to explain that the gradient of the log-likelihood w.r.t. parameters is a functional of the particle measure under boundedness assumptions and fixed parameters, appeal to authority and say doucet did it too. }

\kevin{Gradient linear map, bound conditional likelihoods at time n. Alternatively, convert del moral and doucet expectation bound to high-probability bound. https://arxiv.org/pdf/1905.11546.pdf for trick. Alternatively, concatenate previous states. }

\begin{lem}[Concentration of Measure for Gradient Estimate]
    \label{lemma:grad_bound}
    Consider the gradient estimate obtained by MOP-1, which we know by Lemma \ref{lem:mop-grad-consistency} is consistent for the score, where $\theta = \phi$. For $||\nabla_\theta \hat{\loglik}(\theta) - \nabla_\theta \loglik(\theta)||_2$ to be bounded by $\epsilon$ with probability $1-\delta$, we require
    \begin{align}
    J > \max\left\{2G(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\},
    \end{align}
    where $NG'(\theta) \leq G(\theta)$ are defined in Assumptions \ref{assump:bounded-measurement} and \ref{assump:local-bounded-derivative}, $h(t) = \frac{1}{2}(t - \log(1+t))$, and $\beta_N$ and $r_N$ are two additional finite model-specific constants that do not depend on $J$, but do depend on $N$ and $p$, as defined in \cite{delmoral2011ci}.

    Equivalently, with probability at least $1-\delta$, it holds that
    \begin{align}
        ||\nabla_\theta \hat\ell(\theta) - \nabla_\theta \ell(\theta)||_2 \leq G(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right).
    \end{align}
\end{lem}

\paragraph{Remark:} According to \cite{delmoral2011ci}, under some regularity conditions, $r_N$ and $\beta_N$ are linear in the trajectory length $N$. This corresponds to the finding by \cite{poyiadjis11} that the variance of the estimate is at least quadratic in the trajectory length, and their remark that the result of \cite{delMoral03} establishes that the $L_p$ error is bounded by $O(N^2J^{-1/2})$ (equivalently, the variance is bounded by $O(N^4J^{-1})$) after accounting for the sum over timesteps. The MOP-$1$ variance upper bound is therefore in fact $O(N^4)$, in contrast to the MOP-$\alpha$, where $\alpha<1$, upper bound of $O(N)$.

Either way, usually in practice one considers the estimate normalized through dividing it by $N$, in which case we only need
$$J > \max\left\{2G'(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G'(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\}.$$
Equivalently,
$$||N^{-1}(\nabla_\theta \hat\ell(\theta) - \nabla_\theta \ell(\theta))||_2 \leq G'(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right).$$

It is well-known that bounds like these hide an exponential dependence in the dimension of the state space $d$. We take care to keep track of the terms depending on $d$ when they arise in front of us, but are unable to do so for the terms that arise from the analysis of \cite{delmoral2011ci}.

\kevin{This is slightly alarming, in that the estimate really has $O(N^4)$ variance after accounting for summing over timesteps. In contrast, MOP-0 has only $O(N)$ variance after summing over timesteps. }

\begin{proof}

We will seek to use the concentration inequality of \cite{delmoral2011ci} to bound the deviation of the gradient estimate from the gradient of the negative log-likelihood in the sup norm with a union bound. Fix some $\theta \in \gN(\phi)$.

Recall that as $w_{N, j}^{F, \theta}=1$ when $\theta=\phi$,
$$
\nabla_\theta \hat{\ell}(\theta):=\frac{1}{J} \sum_{j=1}^J \nabla_\theta w_{N, j}^{F, \theta}=\frac{1}{J} \sum_{j=1}^J\sum_{n=1}^N  w_{N, k_j}^{P, \theta} \nabla_\theta \log g_{n,k_j}^{A,\theta} = \frac{1}{J} \sum_{j=1}^J\sum_{n=1}^N \nabla_\theta \log g_{n,k_j}^{A,\theta}.
$$

Define $\varphi_n^i(x_{n,j}^{F,\theta}) := \frac{\partial}{\partial\theta_i} \log g_{n,k_j}^{A,\theta}$, which is a functional of the filtering particles $x_{n,j}^{F,\theta} = x_{n,k_j}^{P,\theta}$. These are bounded measurable functionals bounded by $G'(\theta)$ by Assumption \ref{assump:bounded-measurement}. Therefore, these have bounded oscillation, satisfying the requirement that $\text{osc} \left(\frac{\partial}{\partial\theta_i} \varphi_i(x_{n,j}^{P,\theta}) \right) \leq G'(\theta)$. \cite{delmoral2011ci} in fact assume $\text{osc}(f) \leq 1$, so we simply scale their bound accordingly.

\kevin{Put some justification/make the setup of delmoral2011ci concrete here. Concentration of gradient estimate under particle measure to limit under posterior measure}

Apply the Hoeffding-type concentration inequality from Del Moral and Rio \cite{delmoral2011ci} and a union bound over each $\varphi_n^i(x_{n,j}^{F,\theta})$, totaling $N$ timesteps and $p$ parameters, to find that

\begin{align}
    \max_{n=1,...,N} \left\lVert\frac{1}{J}\sum_{j=1}^J\nabla_\theta \log g_{n,k_j}^{A,\theta} - \nabla_\theta \ell_n(\theta) \right\rVert_{\infty} \leq G'(\theta)\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)
\end{align}

with probability at least $1-2Np\exp(-t)$. Note that the concentration inequality only considers the error from the expectation under the filtering distribution -- but we invoke the consistency of MOP-$1$ shown in Lemma \ref{lem:mop-grad-consistency} to establish that the expectation under the filtering distribution is in fact the score. It therefore holds that with the same probability, that when summing over $N$, as $NG'(\theta) \leq G(\theta)$,

\begin{align}
    \left\lVert\frac{1}{J}\sum_{j=1}^J\sum_{n=1}^N\nabla_\theta \log g_{n,k_j}^{A,\theta} - \nabla_\theta \ell(\theta) \right\rVert_{\infty} 
    &\leq G'(\theta)N\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)\\
    &\leq G(\theta)\left(\frac{r_N}{J}(1+h^{-1}(t)) + \sqrt{\frac{2t}{J}}\beta_N \right)
\end{align}

We split the $\delta$ failure probability among these $2Np$ terms, to find $\delta\leq2Np\exp(-t) => t\leq\log(2Np/\delta)$. Here, $h(t) = \frac{1}{2}(t - \log(1+t))$. The two additional model-specific parameters are $\beta_t$ and $r_t$, which do not depend on $J$.

The analogous bound for the 2-norm follows from scaling the right-hand side by $\sqrt{p}$, to require

\begin{align}
    ||\nabla_\theta \hat\ell(\theta) - \nabla_\theta \ell(\theta)||_2 \leq G(\theta)\left(\frac{r_N\sqrt{p}}{J}(1+h^{-1}(\log(2p/\delta))) + \sqrt{\frac{2p\log(2p/\delta)}{J}}\beta_N\right)
\end{align}

and we therefore need 
\begin{align}
    J > \max\left\{2G(\theta)\frac{r_N\sqrt{p}}{\epsilon}\left(1+h^{-1}\left(\log\left(\frac{2p}{\delta}\right)\right)\right), 8G(\theta)^2\beta_N^2\frac{p\log(2p/\delta)}{\epsilon^2}\right\}.
\end{align}

\end{proof}

\subsection{Bounding Hessian Estimates}

Should one choose to use a second-order method involving a particle Hessian estimate, we provide a guarantee for its positive-definiteness below.

\begin{lem}[Minimum Eigenvalue Bound for Hessian Estimate]
    \label{lemma:hess_bound}
    Assume that the Hessian of the negative log-likelihood $H=\sum_{j=1}^J \E H_j$ has a minimum eigenvalue $0<\gamma<1$, and that $\E \lambda_{\min} (H_j) = \gamma' > 0$.

    We then require 
    \begin{equation}
        J > \max\left\{\frac{2r_t(1+h^{-1}(t)) + 2c}{\gamma'}, \frac{2(2t\beta_t^2+c)^2}{\gamma'^2}\right\} \geq  \frac{r_t(1+h^{-1}(t))}{\gamma'} + \sqrt{2tJ}\beta_t/\gamma' + c/\gamma'
    \end{equation}

    for $\hat{H}(\theta)$ to be invertible and positive definite with minimum eigenvalue greater than or equal to $c \in (0, \sum_{j=1}^J \E\lambda_{\min}(H_j))$ and with probability at least $1-\exp(-t)$.
\end{lem}
\begin{proof}
Write $\hat{H}(\theta) = \hat{H} = \sum_{j=1}^J H_j$ for the estimate of the negative of the Hessian, where $H_j$ is an element of the outer sum over the $J$ particles.

 As the negative log-likelihood is convex, we want to bound the minimum eigenvalue of $\hat{H}(\theta)$ from below with high probability, so that all the eigenvalues of $\hat{H}(\theta)$ are positive with high probability. This ensures that the estimated Hessian is invertible and positive-definite.

From \cite{boyd2004convex}, we know that the minimum eigenvalue of a symmetric matrix is concave. Therefore, it suffices to show that the first inequality in the below expression

\begin{equation}
    0 < \sum_{j=1}^J \lambda_{\min} (H_j) \leq  \lambda_{\min}\left(\sum_{j=1}^J H_j\right) = \lambda_{\min} (\hat{H})
\end{equation}

holds with high probability.

Apply the particle Hoeffding concentration inequality from \cite{delmoral2011ci} to find that

\begin{align}
    \frac{1}{J}\sum_{j=1}^J \lambda_{\min}(H_j) - \E_{\tilde{\pi}_t}\lambda_{\min}(H_j) &= \frac{1}{J}\sum_{j=1}^J \lambda_{\min}(H_j) - \gamma' \geq -\frac{r_t}{J}(1+h^{-1}(t)) - \sqrt{\frac{2t}{J}}\beta_t \\
    \sum_{j=1}^J \lambda_{\min}(H_j) &\geq -r_t(1+h^{-1}(t)) - \sqrt{2tJ}\beta_t + J\gamma'
\end{align}

with probability at least $1-\exp(-t)$. Here, $h(t) = \frac{1}{2}(t - \log(1+t))$. The two additional model-specific parameters are $\beta_t$ and $r_t$, which do not depend on $J$.

We additionally require, for $c \in (0, \sum_{j=1}^J \E\lambda_{\min}(H_j))$,
\begin{align}
    \sum_{j=1}^J \lambda_{\min}(H_j) \geq -r_t(1+h^{-1}(t)) - \sqrt{2tJ}\beta_t + J\gamma' \geq c \\
    J\gamma' \geq c + r_t(1+h^{-1}(t)) + \sqrt{2tJ}\beta_t 
\end{align}

We therefore require $$J > \max\left\{\frac{2r_t(1+h^{-1}(t)) + 2c}{\gamma'}, \frac{2(2t\beta_t^2+c)^2}{\gamma'^2}\right\} \geq  \frac{r_t(1+h^{-1}(t))}{\gamma'} + \sqrt{2tJ}\beta_t/\gamma' + c/\gamma'$$

for $\hat{H}(\theta)$ to be invertible and positive definite with minimum eigenvalue greater than or equal to $c$ with probability at least $1-\exp(-t)$.

\end{proof}

\subsection{Convergence Analysis of Theorem \ref{thm:convergence}}

\begin{proof}

In this analysis, we largely follow the proof of Theorem 6 in \cite{mahoney16}.

Define $\theta_\eta = \theta_n + \eta p_n$, where $p_n=-(H(\theta_n))^{-1}g(\theta_n)$. As in Roosta-Khorasani and Mahoney \cite{mahoney16}, we want to show there is some iteration-independent $\tilde{\eta}>0$ such that the Armijo condition
\begin{equation}
    f(\theta_n+\eta p_n) \leq f(\theta_n) + \eta\beta p_n^Tg(\theta_n)
\end{equation}
holds for any $0< \eta < \tilde{\eta}$ and some $\beta \in (0,1)$.

By an argument found in the beginning of the proof of Theorem 6 in \cite{mahoney16}, we have that choosing $J$ such that $||\nabla_\theta\hat{\ell}(\theta_n) - \nabla_\theta \ell(\theta_n)|| \leq \epsilon$ and $\lambda_{\min}(H(\theta_n)) \geq c$ with probability $(1-\delta)^2$ for each $n$ yields

\begin{align}
    f(\theta_\eta)-f(\theta_n) \leq \eta p_n^Tg(\theta_n) + \epsilon\eta||p_n|| + \eta^2 \Gamma ||p_n||^2 / 2
\end{align}

Now, for some $c>0$, with probability $1-\delta$
\begin{equation}
    p_n^Tg(\theta_n) = -p_n^TH(\theta_n)p_n \geq -c||p_n||^2
\end{equation}
and we can obtain a decrease in the objective.

Substituting this into the previous expression,
\begin{align}
    f(\theta_\eta)-f(\theta_n) \leq -\eta p_n^TH(\theta_n)p_n + \epsilon\eta||p_n|| + \eta^2 \Gamma ||p_n||^2 / 2
\end{align}

and the Armijo condition then becomes
\begin{align}
    -\eta p_n^TH(\theta_n)p_n + \epsilon\eta||p_n|| + \eta^2 \Gamma ||p_n||^2 / 2 &\leq \eta \beta p_n^Tg(\theta_n) = - \eta \beta p_n^TH(\theta_n)p_n \\
    \epsilon||p_n|| + \eta \Gamma ||p_n||^2 / 2 &\leq (1- \beta) p_n^TH(\theta_n)p_n \\
    \epsilon + \eta \Gamma ||p_n|| / 2 &\leq c(1- \beta) ||p_n||
\end{align}

which holds and guarantees an iteration-independent lower bound if 
\begin{equation}
    \eta \leq \frac{c(1-\beta)}{\Gamma}, \;\; \epsilon \leq \frac{c(1-\beta)}{2\Gamma}||g(\theta_n)|| \leq \frac{c(1-\beta)}{2}||p_n||
\end{equation}

Now, first note that
\begin{equation}
||g(\theta_n)|| - ||\nabla_\theta f(\theta_n)|| \leq ||g(\theta_n) - \nabla_\theta f(\theta_n)|| \leq \epsilon => ||\nabla_\theta f(\theta_n)|| \geq ||g(\theta_n)|| - \epsilon
\end{equation} and
\begin{equation}
||\nabla_\theta f(\theta_n)||-||g(\theta_n)|| \leq ||\nabla_\theta f(\theta_n)-g(\theta_n)|| \leq \epsilon => ||g(\theta_n)|| \geq ||\nabla_\theta f(\theta_n)|| - \epsilon
\end{equation}

There are now two cases. If the algorithm terminates and $||g(\theta_n)|| \leq \sigma \epsilon$, we can derive

\begin{equation}
    ||\nabla_\theta f(\theta_n)|| \leq ||g(\theta_n)|| + \epsilon = \sigma\epsilon+\epsilon = (\sigma+1)\epsilon
\end{equation}

If the algorithm does not terminate, then $||g(\theta_n)|| > \sigma \epsilon$. Notice that

\begin{align}
    \epsilon \geq ||g(\theta_n) - \nabla_\theta f(\theta_n)|| &\geq ||g(\theta_n)|| - ||\nabla_\theta f(\theta_n)|| \\
    ||\nabla_\theta f(\theta_n)|| + \epsilon &\geq ||g(\theta_n)|| \geq \sigma \epsilon \\
    ||\nabla_\theta f(\theta_n)|| &\geq \sigma \epsilon - \epsilon = (\sigma - 1)\epsilon \\
    \frac{||\nabla_\theta f(\theta_n)||}{\sigma-1} &\geq \epsilon
\end{align}

and now 
\begin{align}
    ||\nabla_\theta f(\theta_n)|| - \epsilon
    &\geq  ||\nabla_\theta f(\theta_n)|| - \frac{||\nabla_\theta f(\theta_n)||}{\sigma-1} = \left(1-\frac{1}{\sigma-1}\right)||\nabla_\theta f(\theta_n|| = \frac{\sigma-2}{\sigma-1}||\nabla_\theta f(\theta_n)|| \geq \frac{2}{3}||\nabla_\theta f(\theta_n)||
\end{align}

From $||A^{-1}|| = 1/\sigma_{\min}(A)$,
\begin{align}
    p_n^TH(\theta_n)p_n &= (-(H(\theta_n))^{-1}g(\theta_n))^TH(\theta_n)(-(H(\theta_n))^{-1}g(\theta_n)) \\
    &= g(\theta_n)^T(H(\theta_n))^{-1}g(\theta_n) \\
    &\geq \frac{1}{c}||g(\theta_n)||^2 \\
    &\geq \frac{1}{c}(||\nabla_\theta f(\theta_n)|| - \epsilon)^2 \\
    &\geq \frac{4}{9c}||\nabla_\theta f(\theta_n)||^2
\end{align}

From the assumption that $f$ is $\gamma$-strongly convex, $\gamma I \preceq \nabla_\theta^2 -\ell \preceq \Gamma I$, by an implication of $\gamma$-strong convexity we have
\begin{align}
    f(\theta_n) - f(\theta^*) \leq \frac{1}{2\gamma}||\nabla_\theta f(\theta_n)||^2
\end{align}
and we put together:
\begin{equation}
    f(\theta_n) - f(\theta^*) \leq \frac{1}{2\gamma}||\nabla_\theta f(\theta_n)||^2 \leq \frac{9c}{4}\frac{1}{2\gamma}p_n^TH(\theta_n)p_n
\end{equation}
\begin{equation}
    \frac{8\gamma}{9c}(f(\theta_n) - f(\theta^*)) \leq \frac{4}{9c}||\nabla_\theta f(\theta_n)||^2 \leq p_n^TH(\theta_n)p_n
\end{equation}
\begin{equation}
    f(\theta_n) - f(\theta^*) \leq \frac{9c}{8\gamma}p_n^TH(\theta_n)p_n
\end{equation}

\begin{equation}
    -\frac{8\gamma}{9c}(f(\theta_n) - f(\theta^*)) \geq -\frac{4}{9c}||\nabla_\theta f(\theta_n)||^2 \geq -p_n^TH(\theta_n)p_n
\end{equation}

But from earlier, as the Armijo condition is fulfilled with our choice of $\eta$ and $\epsilon$,
\begin{align}
    f(\theta_{n+1})-f(\theta_n) &\leq -\eta p_n^TH(\theta_n)p_n + \epsilon\eta||p_n|| + \eta^2 \Gamma ||p_n||^2 / 2 \\
    &\leq -\eta\beta p_n^TH(\theta_n)p_n \\
    &\leq -\eta\beta\frac{8\gamma}{9c}(f(\theta_n) - f(\theta^*)) 
\end{align}

and therefore
\begin{align}
    f(\theta_{n+1}) - f(\theta^*) 
    &= f(\theta_{n+1})-f(\theta_n)+f(\theta_n)- f(\theta^*) \\
    &\leq f(\theta_n)- f(\theta^*) -\eta\beta\frac{8\gamma}{9c}(f(\theta_n) - f(\theta^*)) \\
    &= (1-\eta\beta\frac{8\gamma}{9c})(f(\theta_n) - f(\theta^*)) 
\end{align}

\end{proof}

\section{Miscellanous Technical Lemmas}

\begin{lem}
    \label{lem:dobrushin-implies-alpha-mixing}
    Setting $X$ to be the particle collection at time $m$, and $Y$ at time $n$, we have that the alpha mixing coefficients,
    $$\int | f_{XY} - f_X f_Y | < \alpha,$$
    are bounded by the Dobrushin coefficient
    $$\int | f_{Y|X}(y|x_1) - f_{Y|X}(y|x_2) |dy < \alpha$$ for all $x_1$, $x_2$.
\end{lem}

\begin{proof}
    We rewrite the alpha-mixing assertion as $$\int { \int |f_Y|X(y|x) - f_Y(y)| dy } f_X dx.$$

    We claim that the Dobrushin coefficient implies $$\int | f_Y|X(y|x_1) - f_Y(y) |dy < \alpha$$ for all $x_1$. This is shown as follows:
    \begin{align*}
        \int | f_{Y|X}(y|x) - f_Y(y) |dy 
        &= \int \left|  \int [f_{Y|X}(y|x_1) - f_{Y|X}(y|x)] f_X(x)dx \right|dy \\
        &< \int   \int  | f_{Y|X}(y|x_1) - f_{Y|X}(y|x)] | dy f_X(x)dx \\
        &< \int \alpha f_X(x)dx \\
        &=\alpha
    \end{align*}

    We then have the desired result.
\end{proof}

\end{document}