### 2024-06-10 paper/appendix/feynman.tex
Updates from Overleaf

In this section, we introduce the Feynman-Kac convention of \cite{delMoral04} that has since become commonplace \cite{karjalainen23} for the analysis of the particle filter. The mathematical formalization and notation introduced here will be adopted