### 2024-07-10 paper/appendix/biasvar.tex

In this section, we prove various rates on the bias, variance, and MSE of MOP-$\alpha$. 
First, we note the following relation between the Dobrushin contraction coefficient and the alpha-mixing coefficients in our context below.
\begin{lem}
    \label{lem:dobrushin-implies-alpha-mixing}
    Setting $X$ to be the particle collection at time $m$, and $Y$ at time $n$, we have that the alpha mixing coefficients,
    \begin{equation}
    \int \big| f_{XY}(x,y) - f_X (x)\, f_Y(y) \big|\, dx\, dy < \alpha,
    \end{equation}
    are bounded by the Dobrushin coefficient, i.e we have
    \begin{equation}
    \int \big| f_{Y|X}(y|x_1) - f_{Y|X}(y|x_2) \big| \, dy < \alpha
    \end{equation} 
    for all $x_1$, $x_2$.
\end{lem}

\begin{proof}
    We rewrite the alpha-mixing assertion as 
    \begin{equation}
    \int { \int \big|f_{Y|X}(y|x) - f_Y(y)\big| \, dy } \, f_X(x) \, dx.
    \end{equation}
    We claim that the Dobrushin coefficient implies 
    \begin{equation}
    \int \big| f_{Y|X}(y|x_1) - f_Y(y) \big| \, dy < \alpha
    \end{equation} 
    for all $x_1$. This is shown as follows:
    \begin{align}
        \int \big| f_{Y|X}(y|x) - f_Y(y) \big| \, dy 
        &= \int \left|  \int \big[f_{Y|X}(y|x_1) - f_{Y|X}(y|x)\big] f_X(x) \, dx \right|dy \\
        &< \int   \int  \big| f_{Y|X}(y|x_1) - f_{Y|X}(y|x)\big| \, dy \, f_X(x)\, dx \\
        &< \int \alpha f_X(x) \, dx \\
        &=\alpha
    \end{align}
    We then have the desired result.
\end{proof}

\subsection{Warm Up: MOP-$0$ Variance Bound}

Note that we made Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} to leverage the results from \cite{karjalainen23} on the forgetting of the particle filter. 
This is required to show error bounds on the gradient estimates that we provide, namely that the error of the MOP-$1$ estimator, corresponding to the estimator of \cite{poyiadjis11}, is $O(N^4)$, and that the variance of MOP-$\alpha$ for any $\alpha<1$ is $O(N)$.

We can decompose the MOP-$\alpha$ estimator as follows. 
When $\theta=\phi$,
\begin{equation}
\hat{\mathcal{L}}(\theta):=\prod_{n=1}^N L_n^{A, \theta, \alpha}=\prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n, j}^{F, \theta}}{\sum_{j=1}^J w_{n, j}^{P, \theta}}=\prod_{n=1}^N L_n^{A, \theta, \alpha}=\prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n, j}^{F, \theta}}{\sum_{j=1}^J (w_{n-1, j}^{F, \theta})^\alpha}.
\end{equation}
Now, to illustrate the proof strategy for the general case in an easier context, we first analyze the special case when $\alpha=0.$ We now prove the variance bound for this case, presented below. To our knowledge, this result for the variance of the gradient estimate of \cite{naesseth18} is novel.

\begin{thm}[Variance of MOP-$0$ Gradient Estimate]
The variance of the gradient estimate from MOP-$0$, i.e. the algorithm of \cite{naesseth18}, is:
 \begin{equation}
     \Var(\nabla_\theta \hat\ell^0(\theta)) \lesssim \frac{Np \, G'(\theta)^2}{J}.
 \end{equation}
\end{thm}

\begin{proof}
When $\alpha=0$, we have:
\begin{align}
    \hat{\lik}^0(\theta) := \prod_{n=1}^N L_n^{A, \theta, 0} &= \prod_{n=1}^N L_n^\phi \cdot \frac{\sum_{j=1}^J w_{n,j}^{F,\theta,0}}{\sum_{j=1}^J w_{n,j}^{P,\theta,0}} 
    \\&
    = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J s_{n,j} = \prod_{n=1}^N L_n^\phi \cdot \frac{1}{J}\sum_{j=1}^J \frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \theta})}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi})}.
\end{align}
Similarly to the proof of Lemma \ref{lem:mop-0-formula}, we define 
\begin{equation}
s_{n,j}=\frac{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \theta})}{f_{Y_n|X_n}(y_n^*|x_{n,j}^{P, \phi})},
\end{equation}
and from the exact same arguments conclude that its gradient when $\theta=\phi$ is therefore 
\begin{align}
    \nabla_\theta \hat{\ell}^0(\theta) &:= \sum_{n=1}^N \nabla_\theta \log\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right) \\
    &= \sum_{n=1}^N \frac{\nabla_\theta \left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right)}{\left(L_n^\phi \frac{1}{J} \sum_{j=1}^J s_{n,j}\right)} \\
    &= \sum_{n=1}^N \frac{\sum_{j=1}^J \nabla_\theta s_{n,j}}{\sum_{j=1}^J s_{n,j}} \\
    &= \sum_{n=1}^N \frac{1}{J} \sum_{j=1}^J \frac{\nabla_\theta f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)}{f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \phi}; \phi)} \\
    &= \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right),
\end{align}
where we use the derivative of the logarithm, observe that $\sum_{j=1}^J s_{n,j} = J$ when $\theta=\phi$, and use the derivative of the logarithm where $\theta=\phi$ again. We do this to establish that the gradient of the log-likelihood estimate is given by the sum of terms over all $N$ and $J$:
\begin{equation}
\nabla_\theta \hat{\ell}^0(\theta) = \frac{1}{J} \sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right).
\end{equation}
Therefore, for a given $\theta_i$ in the parameter vector $\theta$,
\begin{align}
    \Var\left(\frac{\partial}{\partial_{\theta_i}} \hat\ell^0(\theta)\right) &= \frac{1}{J^2}\Var\left(\sum_{n=1}^N\sum_{j=1}^{J}\frac{\partial}{\partial_{\theta_i}} \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right)\right) \\
    & \hspace{-20mm}= \frac{1}{J^2}\sum_{n=1}^N\Var\left(\sum_{j=1}^{J}\frac{\partial}{\partial_{\theta_i}} \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right)\right) \\
    &\hspace{-20mm}+ 2\sum_{m<n}\Cov\left(\frac{1}{J}\sum_{j=1}^{J}\frac{\partial}{\partial_{\theta_i}} \log\left(f_{Y_m|X_{m}}(y_m^*|x_{m,j}^{F, \theta}; \theta)\right), \frac{1}{J}\sum_{j=1}^{J}\frac{\partial}{\partial_{\theta_i}} \log\left(f_{Y_n|X_{n}}(y_n^*|x_{n,j}^{F, \theta}; \theta)\right)\right) \\
    &\hspace{-20mm}= \sum_{n=1}^N \Var\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right) + 2\sum_{m<n} \Cov\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_m^0(\theta),\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right).
\end{align}
Here, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen23} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
\begin{equation}
\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},
\end{equation}
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}.

%\kevin{How does the dobrushin contraction coefficient relate to the alpha-mixing coefficients? If you have a bound on the mixing time $b$, then the alpha mixing coefficients should just be $\alpha(t) \leq e^{-t/b}$.}

By Lemma \ref{lem:dobrushin-implies-alpha-mixing}, the particle filter itself is strong mixing, with $\alpha$-mixing coefficients $a(k) \leq (1-\epsilon)^{\floor{k/(c\log(J))}}$. Therefore, functions of particles are strongly mixing as well, with $\alpha$-mixing coefficients bounded by the original (to see this, observe that the $\sigma$-algebra of the functionals is contained within the original $\sigma$-algebra). Therefore, by Davydov's inequality, noting that $\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\leq G'(\theta)$ by Assumption \ref{assump:bounded-measurement}, and without loss of generality labeling $m$ and $n$ such that $\E\big[(\frac{\partial}{\partial_{\theta_i}}\hat\ell_m^0(\theta))^4\big]^{1/4}\leq\E\big[(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta))^4\big]^{1/4}$, we see that
\begin{align}
    \Cov\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_m^0(\theta), \frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right) 
    &\leq a(n-m)^{1/2} \, \E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_m^0(\theta)\right)^4\right]^{1/4}
    \,
    \E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^4\right]^{1/4}\\
    &\leq a(n-m)^{1/2}\, \E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^4\right]^{1/2}.
\end{align}
To bound this, we use the fact that 
\begin{equation}
\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^4\right] = \E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^4-\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^2\right]^2\right]+\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^2\right]^2
\end{equation}
alongside Lemma 2 of \cite{karjalainen23}, which shows that
\begin{eqnarray}
\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^4-\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^2\right]^2\right] &\lesssim& \frac{G'(\theta)^4}{J^2}, 
\\
\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)-\E\left[\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right]\right)^2\right] &\lesssim& \frac{G'(\theta)^2}{J}.
\end{eqnarray}
It follows that 
\begin{equation}
\E\left[\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)^4\right]^{1/2} \lesssim  \sqrt{\frac{G'(\theta)^4}{J^2}+\left(\frac{G'(\theta)^2}{J}\right)^2} = \frac{G'(\theta)^2}{J},
\end{equation}
and we conclude that 
\begin{equation}
\Cov\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_m^0(\theta), \frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right) \leq (1-\epsilon)^{\frac{1}{2}\floor{\frac{|n-m|}{c\log(J)}}}\frac{G'(\theta)^2}{J}.
\end{equation}
Putting it all together, we see that
\begin{align}
    \Var\left(\frac{\partial}{\partial_{\theta_i}} \hat\ell^\alpha(\theta)\right) &= \sum_{n=1}^N \Var\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right) + 2\sum_{m<n} \Cov\left(\frac{\partial}{\partial_{\theta_i}}\hat\ell_m^0(\theta),\frac{\partial}{\partial_{\theta_i}}\hat\ell_n^0(\theta)\right)\\
    &\leq \frac{NG'(\theta)^2}{J} + 2\sum_{n=1}^{N-1} (N-n)(1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c\log(J)}}}\frac{G'(\theta)^2}{J} \\
    &\leq \frac{G'(\theta)^2}{J} \left(N + 2\sum_{n=1}^{N-1} (N-n) (1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c\log(J)}}}\right) \\
    &\lesssim \frac{G'(\theta)^2N}{J}.
\end{align}
It then follows that as $\theta \in \Theta \subseteq \R^p$, 
\begin{equation}
    \Var\big(\nabla_{\theta} \hat\ell^\alpha(\theta)\big) \lesssim \frac{Np \, G'(\theta)^2}{J}.
\end{equation}
\end{proof}

\paragraph{Remark:} Note that the factor of $N$ that pops up here is due to the use of the unnormalized gradient. If one divides the gradient estimate by $\sqrt{N}$ before usage, the variance does not depend on the horizon. If one divides by $N$, the error is $O(1/\sqrt{NJ})$.

The variance here of $\frac{Np \, G'(\theta)^2}{J}$ can be thought of as a lower bound for the variance of MOP-$\alpha$. We will later see that the variance of MOP-$\alpha$ contains this term, and an additional term that corresponds to the memory of the MOP-$\alpha$ gradient estimate.

%Although the variance of MOP-$0$ is desirable, it unfortunately has large uncontrollable asymptotic bias. Error bounds for the bias and MSE of MOP-$0$ \cite{naesseth18} are also fortunately special cases (where $k=1$, $\alpha=0$) of the bias and MSE analysis we conduct for MOP-$\alpha$. We will later see that these showcase the large uncontrollable asymptotic bias of MOP-$0$.

\subsection{MOP-$\alpha$ Variance Bound}

We are now in a position to tackle the MOP-$\alpha$ variance bound. Here, we analyze the case for $\alpha \in (0,1)$. The proof strategy is as follows. Instead of analyzing the variance of the MOP-$\alpha$ gradient estimate proper, we will analyze the variance of a modified estimator where we truncate the weights at $k$ timesteps, so this estimator only looks at most $k$ timesteps back while accumulating the importance weights. We write
\begin{equation}
    \hat s_k^\alpha(\theta) := \sum_{n=1}^N \log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right)
\end{equation}
for the log-likelihood estimate from the $k$-truncated estimator with discount factor $\alpha$. It is the sum of components
\begin{equation}
\hat s_{n,k}^\alpha(\theta) := \frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \nabla_\theta \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)
\end{equation} 
over timesteps $N$.

This enables us to establish strong mixing for the truncated estimator, in order to get a bound on the variance that is $O(NJ^{-1})$. We emphasize that this truncated estimator is only a theoretical construct -- we do not actually use the truncated estimator. The truncated estimator, as we note in the discussion, possesses similar theoretical guarantees as the MOP-$\alpha$ estimator, but would require the user to specify the number of timesteps to truncate at instead of a discount factor.

Coupled with a bound ensuring that the $k$-truncated estimator provides an estimate that is close to that of MOP-$\alpha$ proper, we get a bound on the variance of MOP-$\alpha$ that comprises of the variance of the $k$-truncated estimator plus the error, for any $k \leq N$. It then holds that the final variance bound on MOP-$\alpha$ is given by the minimum over all $k$, and is never larger than $O(\psi(\alpha)+NJ^{-1})$ for some function $\psi$ increasing in $\alpha$.

\begin{thm}[Variance of MOP-$\alpha$ Gradient Estimate]
    When $\alpha \in (0,1)$, the variance of the gradient estimate from MOP-$\alpha$ is
    \begin{equation}\Var\big(\nabla_\theta \hat\ell^\alpha(\theta)\big) \lesssim \min_{k\leq N} \left(\frac{k^2G'(\theta)^2Np}{(1-\alpha)^2J} + \frac{\alpha^{k}}{1-\alpha}Np \, G'(\theta)^2 \right).
    \end{equation}
\end{thm}
\begin{proof}

Using the derivative of the logarithm and that $w_{n,j}^{F,\theta,\alpha} = w_{n,j}^{F,\theta,\alpha,k} = 1$ when $\theta=\phi$,
\begin{align}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_2^2\\
    &= \left\lVert\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}}}-\frac{\nabla_\theta\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}}{{\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}}}\right\rVert_2^2 \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,\alpha}-\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,\alpha,k}\right\rVert_2^2 \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,\alpha}}{w_{n,j}^{F,\theta,\alpha}}-\frac{1}{J}\sum_{j=1}^J \frac{\nabla_\theta w_{n,j}^{F,\theta,\alpha,k}}{w_{n,j}^{F,\theta,\alpha,k}}\right\rVert_2^2\\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,\alpha}\right)-\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_2^2.
\end{align}

This lets us bound the cumulative weight discrepancies by
\begin{align}
    &  \left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_2^2\\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,\alpha}\right)-\frac{1}{J}\sum_{j=1}^J \nabla_\theta \log\left(w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_2^2 \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\log\left(w_{n,j}^{F,\theta,\alpha}\right)-\log\left(w_{n,j}^{F,\theta,\alpha,k}\right)\right)\right\rVert_2^2\\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=1}^n\alpha^{(n-i)}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) - \sum_{i=n-k}^n{\alpha^{(n-i)}}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)\right)\right\rVert_2^2 \\
    &= \left\lVert\frac{1}{J}\sum_{j=1}^J \nabla_\theta \left(\sum_{i=1}^{n-k}\alpha^{(n-i)}\log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) \right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}p \, G'(\theta)^2\\
    &\leq p \, G'(\theta)^2\frac{\alpha^k-\alpha^n}{1-\alpha},
\end{align}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}.
We can then bound $\|\nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta)\|$ as follows:
\begin{align}
    &\left\lVert\nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta) \right\rVert_2^2
    \\ \nonumber
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \log\left(\frac{\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) - \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) \right\rVert_2^2
    \\ \nonumber
    &= \Bigg\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\log\left(\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)- \log\left(\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)
    \\
    &\qquad\qquad -\log\left(\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right) + \log\left(\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}\right)\Bigg)\Bigg\rVert_2^2 
    \\ 
    &= \left\lVert\sum_{n=1}^N \nabla_\theta \Bigg(\!\log\!\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha}\right) \!-\! \log\!\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)
    \!-\! \log\!\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha}\right) \!+\! \log \! \left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\!\Bigg)\right\rVert_2^2 
    \\ \nonumber
    &\leq \sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right)\right\lVert_{\infty}
    \\ 
    & \qquad\qquad +\sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\right)\right\rVert_2^2,
\end{align}
and each of these two terms is bounded by $\frac{\alpha^k}{1-\alpha}Np \, G'(\theta)^2$. So we now know that
\begin{equation}\left\lVert\nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta) \right\rVert_2^2 \leq  \frac{2\alpha^k}{1-\alpha}Np \, G'(\theta)^2,\end{equation}
which is our desired error bound. 
Now, we bound the variance of $\nabla_\theta \hat s_k^\alpha(\theta)$. 
Recall that we defined
\begin{eqnarray}
\nabla_\theta \hat s_k^\alpha(\theta) &=& \sum_{n=1}^N \nabla_\theta\log\left(\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) 
\\
&=& \sum_{n=1}^N \nabla_\theta \left(\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)-\log\left(\sum_{j=1}^J w_{n-1,j}^{A, F,\theta, \alpha, k}\right)\right).
\end{eqnarray}
We can then decompose this expression with the derivative of the logarithm while noting that $w_{n,j}^{F,\theta,\alpha,k} = 1$ whenever $\theta=\phi$, to see that
\begin{align}
    \nabla_\theta \hat s_k^\alpha(\theta) 
    &= \sum_{n=1}^N \nabla_\theta \left(\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)-\log\left(\sum_{j=1}^J w_{n-1,j}^{A, F,\theta, \alpha, k}\right)\right) \\
    &= \sum_{n=1}^N \left(\frac{\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,\alpha,k}}{\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}}-\frac{\sum_{j=1}^J \nabla_\theta w_{n-1,j}^{A,F,\theta,\alpha, k}}{\sum_{j=1}^J w_{n-1,j}^{A,F,\theta,\alpha, k}}\right) \\
    &= \sum_{n=1}^N \left(\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n,j}^{F,\theta,\alpha,k}-\frac{1}{J}\sum_{j=1}^J \nabla_\theta w_{n-1,j}^{A,F,\theta,\alpha, k}\right)\\
    &= \frac{1}{J}\sum_{n=1}^N \sum_{j=1}^J \nabla_\theta \left(w_{n,j}^{F,\theta,\alpha,k}- w_{n-1,j}^{A,F,\theta,\alpha, k}\right).
\end{align}
It now follows that we need to bound the variance at a single timestep, namely
\begin{equation}
\Var\left(\frac{1}{J}\sum_{j=1}^J\nabla_\theta \left(w_{n,j}^{F,\theta,\alpha,k}- w_{n-1,j}^{A,F,\theta,\alpha, k}\right)\right).
\end{equation}
We use the derivative of the logarithm yet again to find, noting that $w_{n,j}^{F,\theta,\alpha,k}=1$ when $\theta=\phi$, that
\begin{align}
    \nabla_\theta w_{n,j}^{F,\theta,\alpha,k} = \frac{\nabla_\theta w_{n,j}^{F,\theta,\alpha,k}}{w_{n,j}^{F,\theta,\alpha,k}} = \nabla_\theta \log(w_{n,j}^{F,\theta,\alpha,k}) &= \sum_{i=n-k}^n \alpha^{n-i} \nabla_\theta \log\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right) 
    \\
    &= \sum_{i=n-k}^n \alpha^{n-i} \nabla_\theta \log\left(g_{i,j}^{A,F,\theta}\right).
\end{align}
Then,
\begin{align}
    &\Var\left(\nabla_\theta \big(w_{n,j}^{F,\theta,\alpha,k}- w_{n-1,j}^{A,F,\theta,\alpha, k}\big)\right) 
    \\
    &\qquad = \Var\left(\nabla_\theta \left(\sum_{i=n-k}^n \alpha^{n-i} \nabla_\theta \log\left(g_{i,j}^{A,F,\theta}\right)- \sum_{i=n-k-1}^{n-1} \alpha^{n-i} \nabla_\theta \log\left(g_{i,j}^{A,F,\theta}\right)\right)\right) \\
    &\qquad = \Var \left(\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1}\log\left(g_{i-1,j}^{A,F,\theta}\right)\right)\right).
\end{align}
Note that 
\begin{equation}
\nabla_\theta \hat s_{n,k}^\alpha(\theta) := \frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \nabla_\theta \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)
\end{equation} 
is a function bounded by $C\frac{G'(\theta)(1-\alpha^k)}{1-\alpha} =: CG'(\theta)r$ in each coordinate for some constant $C$ (by Assumption \ref{assump:bounded-measurement}) that depends only on $k$ timesteps, and not $n$. 
Subsequently, we suppress vector and matrix notation for the coordinates of $\theta$ and its derivatives, and their variances and covariances, with the variance of a vector interpreted as the sum of the variance of its components.
We invoke Lemma 2 of \cite{karjalainen23} to bound the $L_2$ error of each of the $k$ (from time $n-k$ to time $n$) functionals from its expectation under the posterior by $Cr\sqrt{p}G'(\theta)J^{-1/2}k$. 
That is, we have that
\begin{equation}\E\left[\|\nabla_\theta \hat s_{n,k}^\alpha(\theta) - \E_{\pi}\nabla_\theta \hat s_{n,k}^\alpha(\theta)\|_2^2\right]^{1/2} \leq \frac{Cr\sqrt{p}G'(\theta) k}{\sqrt{J}}.\end{equation}

By the bias-variance decomposition, this in turn bounds the variance at a single timestep by $C^2r^2p \, G'(\theta)^2J^{-1}k^2$. Concretely, we have that
\begin{equation}\Var\left(\frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)\right) \lesssim \frac{r^2k^2p2G'(\theta)^2}{J}.\end{equation}

It now remains to bound the variance of $\nabla_\theta \hat s_k^\alpha(\theta)$ by considering the covariance of each of the $N$ terms that comprise it. 
We decompose
\begin{align}
    \Var\big(\nabla_\theta \hat s_k^\alpha(\theta)\big) &= \Var\left(\sum_{n=1}^N \frac{1}{J}\sum_{j=1}^J\sum_{i=n-k}^n \nabla_\theta\left(\alpha^{n-i} \log\left(g_{i,j}^{A,F,\theta}\right)- \alpha^{n-i+1} \log\left(g_{i-1,j}^{A,F,\theta}\right)\right)\right) \\
    &= \sum_{n=1}^N\Var\big(\nabla_\theta \hat s_{n,k}^\alpha(\theta)\big)+ 2\sum_{m<n}\Cov\big(\nabla_\theta \hat s_{m,k}^\alpha(\theta), \nabla_\theta \hat s_{n,k}^\alpha(\theta)\big) \\
    &\lesssim \frac{r^2k^2p \, G'(\theta)^2N}{J} + 2\sum_{m<n}\Cov\big(\nabla_\theta \hat s_{m,k}^\alpha(\theta), \nabla_\theta \hat s_{n,k}^\alpha(\theta)\big).
\end{align}

Similarly to the proof of the MOP-$0$ case, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen23} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
\begin{equation}
\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},
\end{equation}
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}. By Lemma \ref{lem:dobrushin-implies-alpha-mixing}, the particle filter itself is strongly mixing, with $\alpha$-mixing coefficients $a(l) \leq (1-\epsilon)^{\floor{l/(c\log(J))}}$. Therefore, functions of particles are strongly mixing as well, with $\alpha$-mixing coefficients bounded by the original (to see this, observe that the $\sigma$-algebra of the functionals is contained within the original $\sigma$-algebra).

We now derive the $\alpha$-mixing coefficients of $(\nabla_\theta \hat s_{n,k}^\alpha(\theta))_{n=1}^N$. Observe that $\nabla_\theta \hat s_{n,k}^\alpha$ is strong mixing at lag $k+1$, as all weights beyond $k$ timesteps are truncated. We therefore have that the mixing time of $\nabla_\theta \hat s_{n,k}^\alpha(\theta)$ is $O(1+k+\log(J))$, and that the $\alpha$-mixing coefficients for $(\nabla_\theta \hat s_{n,k}^\alpha(\theta))_{n=1}^N$ are given by $a(l) \leq (1-\epsilon)^{\floor{l/(c(1+k+\log(J)))}}$.
Therefore, by Davydov's inequality and Lemma 2 of \cite{karjalainen23}, and noting that $\nabla_\theta \hat s_{n,k}^\alpha(\theta) \leq CrG'(\theta)$ by Assumption \ref{assump:bounded-measurement}, by a similar argument to the MOP-0 case, we find that 
\begin{equation}
\Cov\big(\nabla_\theta \hat s_{m,k}^\alpha(\theta), \nabla_\theta \hat s_{n,k}^\alpha(\theta)\big) \lesssim (1-\epsilon)^{\frac{1}{2}\floor{\frac{|n-m|}{c(1+k+\log(J))}}}\frac{r^2p \, G'(\theta)^2}{J}.
\end{equation}
Concretely, noting that $\nabla_\theta\hat s_{n,k}^\alpha(\theta)\leq CrG'(\theta)$ by Assumption \ref{assump:bounded-measurement}, and, without loss of generality, assuming $\E[(\nabla_\theta\hat s_{m,k}^\alpha(\theta))^4]^{1/4}\leq\E[(\nabla_\theta\hat s_{n,k}^\alpha(\theta))^4]^{1/4}$, we apply Davydov's inequality to see that
\begin{align}
    \Cov\big(\nabla_\theta\hat s_{m,k}^\alpha(\theta), \nabla_\theta\hat s_{n,k}^\alpha(\theta)\big) 
    &\leq a(n-m)^{1/2} \, \E\big[(\nabla_\theta\hat s_{m,k}^\alpha(\theta))^4\big]^{1/4} \, \E\big[(\nabla_\theta\hat s_{n,k}^\alpha(\theta))^4\big]^{1/4}\\
    &\leq a(n-m)^{1/2} \, \E\big[(\nabla_\theta\hat s_{n,k}^\alpha(\theta))^4\big]^{1/2}.
\end{align}
To bound this, we use the fact that 
\begin{equation}
\E\big[(\nabla_\theta\hat s_{n,k}^\alpha(\theta))^4\big] = \E\Big[\big(\nabla_\theta\hat s_{n,k}^\alpha(\theta)-\E\big[\nabla_\theta\hat s_{n,k}^\alpha(\theta)]\big)^4\Big]+\E\big[(\nabla_\theta\hat s_{n,k}^\alpha(\theta))^2\big]^2
\end{equation}
alongside Lemma 2 of \cite{karjalainen23}, which alongside the fact that
\begin{eqnarray}
\E\Big[\big(\nabla_\theta\hat s_{n,k}^\alpha(\theta)-\E[\nabla_\theta\hat s_{n,k}^\alpha(\theta)]\big)^4\Big] &\lesssim& \frac{r^4p^2G'(\theta)^4}{J^2}, 
\\
\E\Big[\big(\nabla_\theta\hat s_{n,k}^\alpha(\theta)-\E[\nabla_\theta\hat s_{n,k}^\alpha(\theta)]\big)^2\Big] &\lesssim& \frac{r^2p \, G'(\theta)^2}{J},
\end{eqnarray}
allows us to find that 
\begin{equation}
\E\Big[\big(\nabla_\theta\hat s_{n,k}^\alpha(\theta)\big)^4\Big]^{1/2} \lesssim  \sqrt{\frac{r^4p^2G'(\theta)^4}{J^2}+\left(\frac{r^2p \, G'(\theta)^2}{J}\right)^2} = \frac{r^2p \, G'(\theta)^2}{J},
\end{equation}
and conclude that 
\begin{equation}\Cov\big(\nabla_\theta\hat s_{m,k}^\alpha(\theta), \nabla_\theta\hat s_{n,k}^\alpha(\theta)\big) \leq (1-\epsilon)^{\frac{1}{2}\floor{\frac{|n-m|}{c(1+k+\log(J))}}}\frac{r^2p \, G'(\theta)^2}{J}.\end{equation}
Putting it all together, we see that
\begin{align}
    \Var\big(\nabla_\theta \hat s_{k}(\theta)\big) &= \sum_{n=1}^N \Var\big(\nabla_\theta\hat s_{n,k}^\alpha(\theta)\big) + 2\sum_{m<n} \Cov\big(\nabla_\theta\hat s_{m,k}^\alpha(\theta),\nabla_\theta\hat s_{n,k}^\alpha(\theta)\big)
    \\
    &\leq \frac{Nr^2pk^2G'(\theta)^2}{J} + 2\sum_{n=1}^{N-1} (N-n)(1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c(1+k+\log(J))}}}\frac{r^2p \, G'(\theta)^2}{J} 
    \\
    &\leq \frac{r^2pk^2G'(\theta)^2}{J} \left(N + 2\sum_{n=1}^{N-1} (N-n) (1-\epsilon)^{\frac{1}{2}\floor{\frac{n}{c(1+k+\log(J))}}}\right) 
    \\
    &\lesssim \frac{k^2r^2p \, G'(\theta)^2N}{J}.
\end{align}

We will now use this result to bound $\Var\big(\nabla_\theta \hat s_k^\alpha(\theta)\big)$. 
For random variables $X,Y$ where $|X-Y|$ is bounded almost surely by some $M$, we have that 
\begin{equation}
X-\E X - 2M \leq Y - \E Y \leq X - \E X + 2M\end{equation}
so,
\begin{equation}
(Y-\E Y)^2 \leq (X-\E X - 2M)^2 + (X-\E X +2M)^2.\end{equation}
Taking expectations, we get
\begin{equation}
\Var(Y) \leq \Var(X) + 8M^2.
\end{equation}
It follows from this result and the result we proved earlier, namely that
\begin{equation}
\big\|\nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta)\big\|_2^2 \leq \frac{2\alpha^k}{1-\alpha}Np \, G'(\theta)^2,
\end{equation}
that the variance of MOP-$\alpha$ proper is bounded by, for any $k \leq N$,
that the variance of MOP-$\alpha$ proper is bounded by
\begin{equation}
\Var\big(\nabla_\theta \hat\ell^\alpha(\theta)\big) \lesssim \frac{k^2r^2G'(\theta)^2Np}{J} + \frac{\alpha^{k}}{1-\alpha}Np \, G'(\theta)^2.
\end{equation}
As the above holds for any $k \leq N$,
\begin{equation}
\Var\big(\nabla_\theta \hat\ell^\alpha(\theta)\big) \lesssim \min_{k\leq N} \left(\frac{k^2p \, G'(\theta)^2Np}{(1-\alpha)^2J} + \frac{\alpha^{k}}{1-\alpha}Np \, G'(\theta)^2 \right).\end{equation}
\end{proof}

\subsection{MOP-$\alpha$ MSE Bound}

\begin{thm}[MSE of MOP-$\alpha$]
    When $\alpha \in (0,1)$, the MSE of MOP-$\alpha$ is given by
    \begin{equation}
        \E\big\|\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)\big\|_2^2 \; \lesssim \; \min_{k \leq N} Np \, G'(\theta)^2\left(\frac{k^2}{J}+(1-\epsilon)^{\floor{k/(c\log(J))}}+k+\frac{\alpha^k  + \alpha^{k+1} - \alpha}{1-\alpha}\right).
    \end{equation}
\end{thm}

\begin{proof}

The broad idea is to decompose the MSE into three terms as in Equation \ref{eq:mse-decomp}. The first term can be controlled by a mixing argument, the second term garners $O(Nk)$ error, and the third term is $O\left(\frac{N}{1-\alpha}\right)$. It is unclear whether the bias can be reduced further -- the gradient is a sum of $N$ terms, and is therefore $O(N)$ itself.

The first term is the problem term. We will first bound the ground truth of conditional scores from particle approximation conditional on correct filtering distribution at time $n-k$, and then bound that from the particle approximation resulting from an arbitrary filtering distribution at time $n-k$. That is, where $\hat\pi_n$ is the particle approximation at time $n$ of the posterior $\pi_n$ and we write $\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)$ for the truncated MOP-$1$ conditional score estimate given the correct filtering distribution at time $n-k$, we can first observe that
\begin{equation}
    \big\|\nabla_\theta\ell_\theta - \nabla_\theta \hat\ell^\alpha(\theta)\big\|_2^2 =  \left\lVert\sum_{n=1}^N \nabla_\theta \hat\ell_n^\alpha(\theta) - \sum_{n=1}^N \nabla_\theta \ell_n(\theta)\right\rVert_2^2 \leq \sum_{n=1}^N\left\lVert \nabla_\theta \hat\ell_n^\alpha(\theta) -  \nabla_\theta \ell_n(\theta)\right\rVert_2^2,
\end{equation}
and decompose
\begin{align} \nonumber
    & \sum_{n=1}^N\left\lVert \nabla_\theta \hat\ell_n^\alpha(\theta) -  \nabla_\theta \ell_n(\theta)\right\rVert_2^2 \; \leq  \; \sum_{n=1}^N\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\|_2^2  \; +
     \\ & \hspace{20mm}
     \sum_{n=1}^N\|\nabla_\theta\hat s_{n,k}^1(\theta) - \nabla_\theta\hat s_{n,k}^\alpha(\theta)\|_2^2 + \sum_{n=1}^N\|\nabla_\theta\hat s_{n,k}^\alpha(\theta) -  \nabla_\theta\hat\ell_n^\alpha(\theta)\|_2^2. \label{eq:mse-decomp}
\end{align}
We bound the first term of Equation \ref{eq:mse-decomp} , $\sum_{n=1}^N\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\|_2^2$, by decomposing it into two terms, 
\begin{eqnarray}
\nonumber
\hspace{-10mm} \sum_{n=1}^N\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\|_2^2 &\leq&
\\
&& \hspace{-40mm}
\sum_{n=1}^N \|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)\|_2^2 + \sum_{n=1}^N \|\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\|_2^2.
\end{eqnarray}
The first term of Equation \ref{eq:mse-decomp} is a particle approximation dependent on $k$ timesteps, so by Lemma 2 of \cite{karjalainen23}, this is bounded by
\begin{equation}\E\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)\|_2^2 \leq \frac{Cp \, G'(\theta)^2k^2}{J}.\end{equation}

Bounding the second term of Equation \ref{eq:mse-decomp} amounts to bounding the difference between functionals of two different particle measures that mix under the same Markov kernel. Here, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen23} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
\begin{equation}\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},\end{equation}
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}.

Then, we can bound 
$\E\|\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\|_2^2$ by
\begin{equation}\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu \textbf{M}_{n,n+k}-\nu \textbf{M}_{n,n+k}\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}} = \beta_{TV}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},\end{equation}
implying that
\begin{eqnarray}
 && \hspace{-30mm}
    \E\big\|\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\big\|_2^2 
\\
    &\lesssim& \sup_{\mu, \nu} \sup_{\|\psi\|_\infty \leq 1/2} p \, G'(\theta)^2\big|(\mu \textbf{M}_{n,n+k})(\psi)-(\nu \textbf{M}_{n,n+k})(\psi)\big| 
\\
    &\leq& \sup_{\mu, \nu} p \, G'(\theta)^2\big\|\mu \textbf{M}_{n,n+k}-\nu \textbf{M}_{n,n+k}\big\|_{\mathrm{TV}} 
\\
    &\leq& p \, G'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}} \|\hat\pi_{n-k} - \pi_{n-k}\|_{\text{TV}} 
\\
    &\leq& p \, G'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}} \sup_{\|\psi\|_{\infty} \leq 1/2} \big|\hat\pi_{n-k}(\psi) - \pi_{n-k}(\psi)\big| 
\\
    &\lesssim& p \, G'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}}.
\end{eqnarray}
Therefore, we have that
\begin{align}
    &\E\big\|\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)\big\|_2^2 \\
    &\leq \sum_{n=1}^N \E\big\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta)\big\|_2^2 + \sum_{n=1}^N \E\big\|\nabla_\theta\hat s_{n,k|\hat\pi_{n-k}=\pi_{n-k}}^1(\theta) - \nabla_\theta\hat s_{n,k}^1(\theta)\big\|_2^2\\
    &\lesssim N\frac{Cp \, G'(\theta)^2k^2}{J} + Np \, G'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}}.
\end{align}
Now that the first term, $\E\|\nabla_\theta\ell(\theta)\|_2^2$, is taken care of, it remains to bound 
$\E\|\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)\|_2^2$ 
and
$\E\|\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)\|_2^2$.
We see, by a similar argument to the variance bound, that
\begin{align}
    & \hspace{-20mm} \nonumber
    \E\big\|\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)\big\|_2^2
    \\ \nonumber
    &\leq \; \sum_{n=1}^N \E\left\lVert\nabla_\theta \Bigg(\left|\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,1,k}\right)- \log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right|
    + \right.
    \\ & \hspace{30mm} \left.
    \left|\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,1,k}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha,k}\right)\right|\Bigg)\right\rVert_2^2.
\end{align}
Each of these terms can be bounded using the derivative of the logarithm and that $w_{n,j}^{F,\theta,1,k} = w_{n,j}^{F,\theta,\alpha,k} = 1$ when $\theta=\phi$,
\begin{align}
    & \hspace{-30mm}
    \left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,1,k}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_{2}^2
    \\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2
    \\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=n-k}^{n}(1-\alpha^{(n-i)})p \, G'(\theta)^2
    \\
    &\leq p \, G'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right),
\end{align}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}.
So, 
\begin{align}
    \E\big\|\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)\big\|_\infty
    \leq 2Np \, G'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right) 
    \leq 2Np \, G'(\theta)^2k.
\end{align}
Now, we address the third term of Equation \ref{eq:mse-decomp}, $\E\|\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)\|_\infty$. From the variance argument, we know that we can bound $\|\nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta)\|$:
\begin{align}
    &\left\lVert \nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta) \right\rVert_2^2
    \\
    &= \left\lVert\sum_{n=1}^N \! \nabla_\theta \log \! \left( \! \frac{\sum_{j=1}^J\prod_{i=1}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=1}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) 
    \!-\! 
\sum_{n=1}^N\!\nabla_\theta\log\!\left(\!\frac{\sum_{j=1}^J\prod_{i=n-k}^n\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}{\sum_{j=1}^J\prod_{i=n-k}^{n-1}\left(\frac{g_{i,j}^{A,F,\theta}}{g_{i,j}^{A,F,\phi}} \right)^{\alpha^{(n-i)}}}\right) \! \right\rVert_2^2
    \\ \nonumber
    &\leq \sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha}\right)- \log\!\left(\sum_{j=1}^Jw_{n,j}^{F,\theta,\alpha,k}\right)\right)\right\lVert_2^2
    \\ & \hspace{40mm}
    +\sum_{n=1}^N \left\lVert\nabla_\theta \left(\log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta, \alpha}\right) + \log\left(\sum_{j=1}^Jw_{n-1,j}^{A, F,\theta,\alpha, k}\right)\right)\right\rVert_2^2.
\end{align}
Using the derivative of the logarithm and that $w_{n,j}^{F,\theta,\alpha} = w_{n,j}^{F,\theta,\alpha,k} = 1$ when $\theta=\phi$,
\begin{align}
    &\left\lVert\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha}\right)-\nabla_\theta\log\left(\sum_{j=1}^J w_{n,j}^{F,\theta,\alpha,k}\right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}\left\lVert\nabla_\theta\log\left(g_{i,j}^{A,F,\theta} \right)\right\rVert_2^2\\
    &\leq \frac{1}{J}\sum_{j=1}^J \sum_{i=1}^{n-k}\alpha^{(n-i)}p \, G'(\theta)^2\\
    &\leq p \, G'(\theta)^2\frac{\alpha^k-\alpha^n}{1-\alpha},
\end{align}
where the second-last line follows from Assumption \ref{assump:bounded-measurement}. Putting it together and taking expectations on both sides, we obtain
\begin{equation}
\E\left\lVert\nabla_\theta\hat\ell^\alpha(\theta) - \nabla_\theta \hat s_k^\alpha(\theta) \right\rVert_2^2 \leq  \frac{2\alpha^k}{1-\alpha}Np \, G'(\theta)^2,
\end{equation}
which is our desired error bound. 
Therefore, our decomposition yields the MSE bound
\begin{align}
    & \hspace{-5mm}
    \E\big\|\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^\alpha(\theta)\big\|_2^2 
    \\
    &\leq \E\big\|\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_k^1(\theta)\big\|_2^2 + \E\big\|\nabla_\theta\hat s_k^1(\theta) - \nabla_\theta\hat s_k^\alpha(\theta)\big\|_2^2 + \E\big\|\nabla_\theta\hat s_k^\alpha(\theta) -  \nabla_\theta\hat\ell^\alpha(\theta)\big\|_2^2 
    \\ \nonumber
    &\lesssim \min_{k \leq N} \left(N\frac{Cp \, G'(\theta)^2k^2}{J} + Np \, G'(\theta)^2(1-\epsilon)^{\floor{k/(c\log(J))}} \right.
    \\
    & \hspace{40mm} \left.
    + 2Np \, G'(\theta)^2\left(k-\frac{\alpha(1-\alpha^k)}{1-\alpha}\right) + 2\frac{\alpha^k}{1-\alpha}Np \, G'(\theta)^2\right) \\
    &\lesssim \min_{k \leq N} Np \, G'(\theta)^2\left(\frac{k^2}{J}+(1-\epsilon)^{\floor{k/(c\log(J))}}+k+\frac{\alpha^k  + \alpha^{k+1} - \alpha}{1-\alpha}\right).
\end{align}

\end{proof}

\begin{cor}[MSE of MOP-$0$]
    The MSE of MOP-$0$, i.e. the estimator of \cite{naesseth18}, is
    \begin{equation}
        \E\big\|\nabla_\theta\ell(\theta) - \nabla_\theta \hat\ell^0(\theta)\big\|_2^2 \; \lesssim 
        \; Np \, G'(\theta)^2\left(\frac{1}{J}+(1-\epsilon)^{\floor{1/(c\log(J))}}\right).
    \end{equation}
\end{cor}
\begin{proof}
    Observe that MOP-$0$ is equivalent to MOP-$(1,1)$. Noting that $k=1$ and that we only need to bound the difference between $\nabla_\theta \ell(\theta)$ and $\nabla_\theta \hat s_1^1(\theta)$, repeating the above arguments almost verbatim allows us to obtain this result.

As before, we bound the first term, $\sum_{n=1}^N\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,1}^1(\theta)\|_2^2$, by decomposing it into two terms, 
\begin{eqnarray} \nonumber
\hspace{-10mm} \sum_{n=1}^N\big\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,1}^1(\theta)\big\|_2^2 &\leq 
\\ &&\hspace{-45mm} 
\sum_{n=1}^N \big\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta)\big\|_2^2 + \sum_{n=1}^N \big\|\nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta) - \nabla_\theta\hat s_{n,1}^1(\theta)\big\|_2^2.
\end{eqnarray}
The first term is a particle approximation dependent on $1$ timesteps, so by Lemma 2 of \cite{karjalainen23}, this is bounded by
\begin{equation}\E\big\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta)\big\|_2^2 \leq \frac{Cp \, G'(\theta)^2}{J}.\end{equation}

The second term amounts to bounding the difference between functionals of two different particle measures that mix under the same Markov kernel. Here, we use Assumptions \ref{assump:bounded-process} and \ref{assump:bounded-measurement} that ensure strong mixing. We know from Theorem 3 of \cite{karjalainen23} that when $\textbf{M}_{n,n+k}$ is the $k$-step Markov operator from timestep $n$ and $\beta_{\text{TV}}(M) = \sup _{x, y \in E}\|M(x, \cdot)-M(y, \cdot)\|_{\mathrm{TV}}=\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu M-\nu M\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}}$ is the Dobrushin contraction coefficient of a Markov operator, 
\begin{equation}
\beta_{\text{TV}}(\textbf{M}_{n,n+k}) \leq (1-\epsilon)^{\floor{k/(c\log(J))}},
\end{equation}
i.e. the mixing time of the particle filter is $O(\log(J))$, where $\epsilon$ and $c$ depend on $\bar{M}, \underbar{M}, \bar{G}, \underbar{G}$ in \ref{assump:bounded-process} and \ref{assump:bounded-measurement}.

Then, we can bound 
$\E\|\nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta) - \nabla_\theta\hat s_{n,1}^1(\theta)\|_2^2$ by
\begin{equation}
\sup _{\mu, \nu \in \mathcal{P}, \mu \neq \nu} \frac{\|\mu \textbf{M}_{n,n+1}-\nu \textbf{M}_{n,n+1}\|_{\mathrm{TV}}}{\|\mu-\nu\|_{\mathrm{TV}}} = \beta_{TV}(\textbf{M}_{n,n+1}) \leq (1-\epsilon)^{\floor{1/(c\log(J))}},
\end{equation}
implying that
\begin{align}
    & \hspace{-10mm}
    \E\big\|\nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta) - \nabla_\theta\hat s_{n,1}^1(\theta)\big\|_2^2 
    \\ & \hspace{30mm}
    \lesssim \sup_{\mu, \nu} \sup_{\|\psi\|_\infty \leq 1/2} p \, G'(\theta)^2\big|(\mu \textbf{M}_{n,n+1})(\psi)-(\nu \textbf{M}_{n,n+1})(\psi)\big| 
    \\ & \hspace{30mm}
    \leq \sup_{\mu, \nu} p \, G'(\theta)^2\big\|\mu \textbf{M}_{n,n+1}-\nu \textbf{M}_{n,n+1}\big\|_{\mathrm{TV}} 
    \\ & \hspace{30mm}
    \leq p \, G'(\theta)^2(1-\epsilon)^{\floor{1/(c\log(J))}} \|\hat\pi_{n-1} - \pi_{n-1}\|_{\text{TV}} 
    \\ & \hspace{30mm}
    \leq p \, G'(\theta)^2(1-\epsilon)^{\floor{1/(c\log(J))}} \sup_{\|\psi\|_{\infty} \leq 1/2} \big|\hat\pi_{n-1}(\psi) - \pi_{n-1}(\psi)\big| 
    \\ & \hspace{30mm}
    \lesssim p \, G'(\theta)^2(1-\epsilon)^{\floor{1/(c\log(J))}}.
\end{align}

Therefore, we have that
\begin{align}
    &\E\big\|\nabla_\theta\ell(\theta) - \nabla_\theta\hat s_1^1(\theta)\big\|_2^2 \\
    &\leq \sum_{n=1}^N \E\big\|\nabla_\theta\ell_n(\theta) - \nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta)\big\|_2^2 + \sum_{n=1}^N \E\big\|\nabla_\theta\hat s_{n,1|\hat\pi_{n-1}=\pi_{n-1}}^1(\theta) - \nabla_\theta\hat s_{n,1}^1(\theta)\big\|_2^2\\
    &\lesssim N\frac{Cp \, G'(\theta)^2}{J} + Np \, G'(\theta)^2(1-\epsilon)^{\floor{1/(c\log(J))}}.
\end{align}
\end{proof}