### 2024-07-14 paper/si.tex

\documentclass[9pt,twoside,lineno]{pnas-new-si}

% Use the lineno option to display guide line numbers if required.
\usepackage{xr}
\makeatletter
\newcommand*{\addFileDependency}[1]{% argument=file name and extension
  \typeout{(#1)}
  \@addtofilelist{#1}
  \IfFileExists{#1}{}{\typeout{No file #1.}}
}
\makeatother

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\externaldocument{ms}

\templatetype{pnassupportinginfo}
\input{macros}
\usepackage{refcount}

\setcounter{thm}{\getrefnumber{finalthm}}
\addtocounter{thm}{-1}

\setcounter{lem}{\getrefnumber{finallem}}
\addtocounter{lem}{-1}

\setcounter{defn}{\getrefnumber{finaldefn}}
\addtocounter{defn}{-1}

\setcounter{prop}{\getrefnumber{finalprop}}
\addtocounter{prop}{-1}

\setcounter{aspt}{\getrefnumber{finalaspt}}
\addtocounter{aspt}{-1}

\begin{document}

\title{Accelerated Inference for Partially Observed Stochastic Processes with Automatic Differentiation}
\author{Kevin Tan, Giles Hooker, Edward L. Ionides}
\correspondingauthor{Corresponding Author name.\\E-mail: kevtan@wharton.upenn.edu}

%% Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found".
%\instructionspage

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\section{MOP-$\alpha$ Functional Forms}

\label{appendix:functional}
\input{appendix/functional}

\section{Optimization Convergence Analysis}
\label{appendix:convergence}
\input{appendix/convergence}

\section{Feynman-Kac Models and Monte Carlo Approximations}
\label{appendix:feynman}
\input{appendix/feynman}

\section{A Strong Law of Large Numbers for Triangular Arrays of Particles With Off-Parameter Resampling}
\label{appendix:targeting}
\input{appendix/targeting}

\section{Consistency of Off-Parameter Resampled Gradient Estimates}
\label{appendix:consistency}
\input{appendix/consistency}

\section{Bias-Variance Analysis}
\label{appendix:biasvar}
\input{appendix/biasvar}

\section{Figures for Bayesian Inference}
\label{appendix:bayes}
\input{appendix/bayes}

%%% Each figure should be on its own page

\bibliography{paper/bib-ifad}

\end{document}